---
title: "SSA revenue instruments: pooled data and initial descriptives"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook builds the pooled de jure Census and statutory-allocation datasets from the final country-level XLSX workbooks. It does not measure actual collections, transfers, disbursements, or expenditure execution. It does not create plots.

The analytical keys are always country-specific: `country + instrument_id` for Census instruments and `country + pair_id` for allocation channels. No reference is ever matched across countries.

```{r setup, include=FALSE}
input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.path = file.path(project_root, "figures", "01_initial_descriptives-")
)

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)
library(tibble)
library(jsonlite)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

country_root <- file.path(project_root, "countries")
derived_dir <- file.path(project_root, "data", "derived")
tables_dir <- file.path(project_root, "analysis", "tables")
dir.create(derived_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)

census_output_path <- file.path(derived_dir, "SSA_CENSUS_STACKED.csv")
allocation_output_path <- file.path(
  derived_dir,
  "SSA_EARMARK_ALLOCATIONS_STACKED.csv"
)
workbook_output_path <- file.path(
  tables_dir,
  "01_initial_descriptives.xlsx"
)

manifest_path <- Sys.getenv(
  "INITIAL_DESC_MANIFEST",
  unset = file.path(tempdir(), "01_initial_descriptives_tables.json")
)
dir.create(dirname(manifest_path), recursive = TRUE, showWarnings = FALSE)

census_columns <- c(
  "evidence_id", "row_type", "country", "instrument_id", "document_id",
  "document_type", "document_year", "legal_article", "page_start", "page_end",
  "official_name", "official_name_underspecified", "in_force", "tax_instrument",
  "tax_base_detail", "instrument_label_verbatim", "instrument_nature",
  "base_sector", "base_division", "is_penalty", "rate_value",
  "rate_is_schedule", "rate_schedule_detail", "rate_type", "rate_basis",
  "rate_base_quantum", "rate_reference_instrument", "change_type",
  "enabling_reference", "evidence_status"
)

allocation_columns <- c(
  "evidence_id", "row_type", "country", "instrument_id", "pair_id",
  "document_id", "document_type", "document_year", "legal_article",
  "page_start", "page_end", "official_name", "destination",
  "beneficiary_type", "allocation_nature", "share_value",
  "share_is_schedule", "share_schedule_detail", "share_type", "share_basis",
  "share_level", "share_pool", "is_purpose_restricted", "assignment_type",
  "census_ref", "destination_function", "destination_function_basis",
  "destination_function_multi", "destination_function_detail"
)

required_columns <- list(
  Census = census_columns,
  Allocation = allocation_columns
)

required_sheets <- c(
  Census = "census_rows",
  Allocation = "evidence_table"
)

numeric_fields <- c("rate_value", "rate_base_quantum", "share_value")
integer_fields <- c("page_start", "page_end", "share_level")
binary_fields <- c(
  "official_name_underspecified", "in_force", "is_penalty",
  "rate_is_schedule", "share_is_schedule", "is_purpose_restricted",
  "destination_function_multi"
)
flexible_text_fields <- c(
  "document_year", "base_division", "destination_function",
  "destination_function_detail"
)

target_type_for <- function(field) {
  case_when(
    field %in% numeric_fields ~ "numeric",
    field %in% integer_fields ~ "integer",
    field %in% binary_fields ~ "binary_0_1",
    TRUE ~ "text"
  )
}

is_blank_cell <- function(x) {
  length(x) == 0L || all(is.na(x)) ||
    (is.character(x) && length(x) == 1L && str_trim(x) == "")
}

cell_storage_type <- function(x) {
  if (is_blank_cell(x)) {
    return(NA_character_)
  }
  class(x)[1]
}

cell_to_text <- function(x) {
  if (is_blank_cell(x)) {
    return(NA_character_)
  }
  if (length(x) != 1L) {
    return(NA_character_)
  }
  if (inherits(x, "POSIXt")) {
    return(format(x, "%Y-%m-%d %H:%M:%S", tz = "UTC"))
  }
  if (is.character(x)) {
    return(x)
  }
  if (is.numeric(x)) {
    return(format(x, scientific = FALSE, trim = TRUE, digits = 15))
  }
  if (is.logical(x)) {
    return(if_else(x, "TRUE", "FALSE"))
  }
  as.character(x)
}

is_numeric_text <- function(x) {
  !is.na(x) && str_detect(
    x,
    "^[+-]?(?:[0-9]+\\.?[0-9]*|\\.[0-9]+)(?:[eE][+-]?[0-9]+)?$"
  ) && is.finite(suppressWarnings(as.numeric(x)))
}

is_integer_text <- function(x) {
  if (!is_numeric_text(x)) {
    return(FALSE)
  }
  value <- suppressWarnings(as.numeric(x))
  abs(value - round(value)) < 1e-9
}

is_binary_text <- function(x) {
  !is.na(x) && tolower(x) %in% c("0", "1", "true", "false")
}

cell_is_compatible <- function(x, field, target_type) {
  if (is_blank_cell(x)) {
    return(TRUE)
  }
  storage_type <- cell_storage_type(x)
  text_value <- cell_to_text(x)

  if (target_type == "text") {
    if (field %in% flexible_text_fields) {
      return(storage_type %in% c("character", "numeric", "integer"))
    }
    return(storage_type == "character")
  }
  if (target_type == "numeric") {
    return(is_numeric_text(text_value))
  }
  if (target_type == "integer") {
    return(is_integer_text(text_value))
  }
  if (target_type == "binary_0_1") {
    return(is_binary_text(text_value))
  }
  FALSE
}

convert_list_column <- function(column, field, target_type) {
  text_values <- map_chr(column, cell_to_text)

  if (target_type == "text") {
    return(text_values)
  }
  if (target_type == "numeric") {
    return(if_else(
      is.na(text_values),
      NA_real_,
      suppressWarnings(as.numeric(text_values))
    ))
  }
  if (target_type == "integer") {
    return(if_else(
      is.na(text_values),
      NA_integer_,
      as.integer(round(suppressWarnings(as.numeric(text_values))))
    ))
  }
  if (target_type == "binary_0_1") {
    return(case_when(
      is.na(text_values) ~ NA_integer_,
      tolower(text_values) %in% c("1", "true") ~ 1L,
      tolower(text_values) %in% c("0", "false") ~ 0L,
      TRUE ~ NA_integer_
    ))
  }
  stop("Unknown target type: ", target_type)
}

is_missing_value <- function(x) {
  if (is.character(x)) {
    return(is.na(x) | str_trim(x) == "")
  }
  is.na(x)
}

display_table <- function(x, digits = 4) {
  if (nrow(x) == 0L) {
    cat("No rows to report.\n\n")
  }
  knitr::kable(x, digits = digits, row.names = FALSE)
}
```

# 1. Source discovery and schema validation

The recursive search is anchored to the two canonical country output locations specified in the methodology. Generated pooled outputs, temporary Excel files, reconciled/intermediate workbooks, and the COFOG CSV are not candidate country inputs.

```{r discover-files}
all_country_xlsx <- list.files(
  country_root,
  pattern = "\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE,
  ignore.case = TRUE
)

all_country_paths <- str_replace_all(
  normalizePath(all_country_xlsx, winslash = "/", mustWork = TRUE),
  fixed(project_root),
  ""
) %>%
  str_remove("^/")

candidate_inventory <- tibble(
  relative_path = all_country_paths,
  file_name = basename(all_country_paths),
  country = str_match(all_country_paths, "^countries/([^/]+)/")[, 2],
  dataset = case_when(
    str_detect(
      all_country_paths,
      regex(
        "^countries/[^/]+/outputs/extract/[^/]+_CENSUS\\.xlsx$",
        ignore_case = TRUE
      )
    ) ~ "Census",
    str_detect(
      all_country_paths,
      regex(
        paste0(
          "^countries/[^/]+/outputs/allocations/",
          "[^/]+_EARMARKS?_ALLOCATIONS?\\.xlsx$"
        ),
        ignore_case = TRUE
      )
    ) ~ "Allocation",
    TRUE ~ NA_character_
  ),
  is_temporary = str_starts(file_name, "~$")
) %>%
  filter(!is.na(dataset), !is_temporary) %>%
  mutate(
    absolute_path = file.path(project_root, relative_path),
    required_sheet = unname(required_sheets[dataset])
  ) %>%
  arrange(country, dataset, relative_path)

duplicate_candidates <- candidate_inventory %>%
  count(country, dataset, name = "candidate_count") %>%
  filter(candidate_count > 1L)

if (nrow(duplicate_candidates) > 0L) {
  duplicate_detail <- candidate_inventory %>%
    semi_join(duplicate_candidates, by = c("country", "dataset")) %>%
    select(country, dataset, relative_path)
  print(duplicate_detail)
  stop(
    "Several plausible final files exist for at least one country and dataset. ",
    "No file was selected."
  )
}

inspect_schema <- function(dataset, country, absolute_path, relative_path,
                           required_sheet) {
  expected <- required_columns[[dataset]]

  tryCatch({
    sheets <- excel_sheets(absolute_path)
    if (!required_sheet %in% sheets) {
      return(tibble(
        dataset = dataset,
        country = country,
        relative_path = relative_path,
        sheet = NA_character_,
        source_rows = NA_integer_,
        source_columns = NA_integer_,
        missing_required = paste(expected, collapse = "; "),
        additional_columns = NA_character_,
        duplicate_headers = NA_character_,
        required_order_matches = FALSE,
        schema_status = "failed_missing_sheet",
        schema_reason = paste0("Required sheet not found: ", required_sheet)
      ))
    }

    header_only <- read_excel(
      absolute_path,
      sheet = required_sheet,
      n_max = 0,
      .name_repair = "minimal"
    )
    headers <- names(header_only)
    duplicate_headers <- unique(headers[duplicated(headers)])
    missing_required <- setdiff(expected, headers)
    additional_columns <- setdiff(headers, expected)
    source_rows <- nrow(read_excel(
      absolute_path,
      sheet = required_sheet,
      col_types = "list",
      trim_ws = FALSE,
      .name_repair = "minimal"
    ))
    present_required <- headers[headers %in% expected]
    expected_present <- expected[expected %in% headers]

    schema_ok <- length(missing_required) == 0L &&
      length(duplicate_headers) == 0L

    schema_reason_value <- if (length(missing_required) > 0L) {
      paste0(
        "Missing required columns: ",
        paste(missing_required, collapse = "; ")
      )
    } else if (length(duplicate_headers) > 0L) {
      paste0(
        "Duplicated headers: ",
        paste(duplicate_headers, collapse = "; ")
      )
    } else {
      NA_character_
    }

    tibble(
      dataset = dataset,
      country = country,
      relative_path = relative_path,
      sheet = required_sheet,
      source_rows = source_rows,
      source_columns = length(headers),
      missing_required = if_else(
        length(missing_required) == 0L,
        NA_character_,
        paste(missing_required, collapse = "; ")
      ),
      additional_columns = if_else(
        length(additional_columns) == 0L,
        NA_character_,
        paste(additional_columns, collapse = "; ")
      ),
      duplicate_headers = if_else(
        length(duplicate_headers) == 0L,
        NA_character_,
        paste(duplicate_headers, collapse = "; ")
      ),
      required_order_matches = identical(present_required, expected_present),
      schema_status = if_else(schema_ok, "schema_ok", "failed_schema"),
      schema_reason = schema_reason_value
    )
  }, error = function(e) {
    tibble(
      dataset = dataset,
      country = country,
      relative_path = relative_path,
      sheet = required_sheet,
      source_rows = NA_integer_,
      source_columns = NA_integer_,
      missing_required = NA_character_,
      additional_columns = NA_character_,
      duplicate_headers = NA_character_,
      required_order_matches = FALSE,
      schema_status = "failed_read",
      schema_reason = conditionMessage(e)
    )
  })
}

schema_audit <- pmap_dfr(
  candidate_inventory %>%
    select(dataset, country, absolute_path, relative_path, required_sheet),
  inspect_schema
)
```

## File discovery result

```{r print-schema-audit}
display_table(schema_audit)
```

# 2. Type validation and pooled stacks

Required columns are read as typed Excel cells first. Storage types and semantic compatibility are audited before any common conversion. Code fields whose Excel storage legitimately varies between numeric and text are converted explicitly to text without padding or recoding their values. Any incompatible non-missing value fails the entire source file.

```{r load-and-stack}
schema_valid_files <- candidate_inventory %>%
  inner_join(
    schema_audit %>%
      filter(schema_status == "schema_ok") %>%
      select(dataset, country, relative_path, sheet),
    by = c("dataset", "country", "relative_path")
  )

load_validated_file <- function(dataset, country, absolute_path, relative_path,
                                required_sheet, sheet) {
  expected <- required_columns[[dataset]]

  tryCatch({
    raw_data <- read_excel(
      absolute_path,
      sheet = sheet,
      col_types = "list",
      trim_ws = FALSE,
      .name_repair = "minimal"
    ) %>%
      select(all_of(expected))

    field_audit <- map_dfr(expected, function(field) {
      target_type <- target_type_for(field)
      column <- raw_data[[field]]
      storage_types <- map_chr(column, cell_storage_type)
      compatible <- map_lgl(
        column,
        cell_is_compatible,
        field = field,
        target_type = target_type
      )

      tibble(
        dataset = dataset,
        country = country,
        relative_path = relative_path,
        field = field,
        target_type = target_type,
        observed_storage_types = {
          observed <- sort(unique(na.omit(storage_types)))
          if (length(observed) == 0L) "all_missing" else paste(observed, collapse = "; ")
        },
        n_nonmissing = sum(!is.na(storage_types)),
        n_incompatible = sum(!compatible),
        conversion_rule = case_when(
          target_type == "text" && field %in% flexible_text_fields ~
            "explicit character conversion; no padding or recoding",
          target_type == "text" ~ "preserve character value exactly",
          target_type == "numeric" ~ "strict numeric parse",
          target_type == "integer" ~ "strict whole-number parse",
          target_type == "binary_0_1" ~ "normalize 0/1 or TRUE/FALSE to integer 0/1",
          TRUE ~ NA_character_
        ),
        type_status = if_else(
          sum(!compatible) == 0L,
          "compatible",
          "incompatible"
        )
      )
    })

    if (any(field_audit$n_incompatible > 0L)) {
      return(list(
        status = "failed_type",
        reason = "One or more required fields contained incompatible values",
        n_rows = nrow(raw_data),
        data = NULL,
        type_audit = field_audit
      ))
    }

    converted_columns <- map(expected, function(field) {
      convert_list_column(
        raw_data[[field]],
        field,
        target_type_for(field)
      )
    })
    names(converted_columns) <- expected
    converted_data <- as_tibble(converted_columns)

    observed_countries <- converted_data$country[
      !is_missing_value(converted_data$country)
    ] %>%
      unique()

    country_ok <- length(observed_countries) == 1L &&
      identical(observed_countries, country)

    if (!country_ok) {
      return(list(
        status = "failed_country",
        reason = paste0(
          "Workbook country values do not uniquely equal folder country ",
          country,
          ": ",
          paste(observed_countries, collapse = "; ")
        ),
        n_rows = nrow(raw_data),
        data = NULL,
        type_audit = field_audit
      ))
    }

    list(
      status = "loaded",
      reason = NA_character_,
      n_rows = nrow(converted_data),
      data = converted_data,
      type_audit = field_audit
    )
  }, error = function(e) {
    list(
      status = "failed_read",
      reason = conditionMessage(e),
      n_rows = NA_integer_,
      data = NULL,
      type_audit = tibble(
        dataset = character(), country = character(), relative_path = character(),
        field = character(), target_type = character(),
        observed_storage_types = character(), n_nonmissing = integer(),
        n_incompatible = integer(), conversion_rule = character(),
        type_status = character()
      )
    )
  })
}

load_results <- pmap(
  schema_valid_files %>%
    select(
      dataset, country, absolute_path, relative_path,
      required_sheet, sheet
    ),
  load_validated_file
)

load_status <- schema_valid_files %>%
  select(dataset, country, relative_path) %>%
  mutate(
    load_status = map_chr(load_results, "status"),
    load_reason = map_chr(load_results, function(x) {
      if (is.null(x$reason) || length(x$reason) == 0L) NA_character_ else x$reason
    }),
    loaded_rows = map_int(load_results, function(x) {
      if (is.null(x$n_rows) || length(x$n_rows) == 0L) NA_integer_ else x$n_rows
    })
  )

type_audit <- map_dfr(load_results, "type_audit")

loaded_data_index <- which(load_status$load_status == "loaded")
loaded_data <- load_results[loaded_data_index] %>% map("data")
loaded_metadata <- schema_valid_files[loaded_data_index, ]

census <- loaded_data[
  loaded_metadata$dataset == "Census"
] %>%
  bind_rows()

allocation <- loaded_data[
  loaded_metadata$dataset == "Allocation"
] %>%
  bind_rows()

if (nrow(census) == 0L || nrow(allocation) == 0L) {
  stop("At least one pooled dataset has no successfully loaded country files.")
}

stopifnot(identical(names(census), census_columns))
stopifnot(identical(names(allocation), allocation_columns))

file_inventory <- schema_audit %>%
  left_join(
    load_status,
    by = c("dataset", "country", "relative_path")
  ) %>%
  mutate(
    candidate_status = "found",
    final_status = case_when(
      schema_status != "schema_ok" ~ schema_status,
      !is.na(load_status) ~ load_status,
      TRUE ~ "failed_unknown"
    ),
    final_reason = coalesce(schema_reason, load_reason),
    loaded_rows = if_else(final_status == "loaded", loaded_rows, NA_integer_)
  ) %>%
  select(
    dataset, country, relative_path, candidate_status, required_sheet = sheet,
    source_rows, loaded_rows, final_status, final_reason
  ) %>%
  arrange(dataset, country)

file_status_summary <- expand_grid(
  dataset = c("Census", "Allocation"),
  status = c("found", "loaded", "failed", "duplicated", "skipped")
) %>%
  left_join(
    bind_rows(
      file_inventory %>%
        count(dataset, status = candidate_status, name = "n_files"),
      file_inventory %>%
        transmute(
          dataset,
          status = case_when(
            final_status == "loaded" ~ "loaded",
            str_starts(final_status, "failed") ~ "failed",
            TRUE ~ final_status
          )
        ) %>%
        count(dataset, status, name = "n_files")
    ) %>%
      group_by(dataset, status) %>%
      summarise(n_files = sum(n_files), .groups = "drop"),
    by = c("dataset", "status")
  ) %>%
  mutate(n_files = replace_na(n_files, 0L))

stack_reconciliation <- bind_rows(
  tibble(
    dataset = "Census",
    loaded_source_rows = sum(file_inventory$loaded_rows[
      file_inventory$dataset == "Census"
    ], na.rm = TRUE),
    stacked_rows = nrow(census),
    rows_reconcile = loaded_source_rows == stacked_rows
  ),
  tibble(
    dataset = "Allocation",
    loaded_source_rows = sum(file_inventory$loaded_rows[
      file_inventory$dataset == "Allocation"
    ], na.rm = TRUE),
    stacked_rows = nrow(allocation),
    rows_reconcile = loaded_source_rows == stacked_rows
  )
)

stopifnot(all(stack_reconciliation$rows_reconcile))
```

## File and type status

```{r print-file-status}
display_table(file_status_summary)
```

```{r print-file-inventory}
display_table(file_inventory)
```

```{r print-type-audit}
display_table(type_audit)
```

```{r print-stack-reconciliation}
display_table(stack_reconciliation)
```

# 3. Flatten and resolve `census_ref`

Each non-missing `census_ref` is expected to be a JSON array. Tokens are trimmed only in a derived lookup field; the raw allocation variable remains unchanged. Same-country Census `evidence_id` is tried first, followed explicitly by same-country Census `instrument_id`. Matching both namespaces to the same unique Census record is reported as namespace overlap, not ambiguity.

```{r parse-census-ref}
parse_reference_cell <- function(country, allocation_evidence_id, pair_id,
                                 census_ref_raw, allocation_row) {
  base_row <- tibble(
    country = country,
    allocation_row = allocation_row,
    allocation_evidence_id = allocation_evidence_id,
    pair_id = pair_id,
    census_ref_raw = census_ref_raw
  )

  if (is.na(census_ref_raw) || str_trim(census_ref_raw) == "") {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "missing_raw",
        parse_detail = "census_ref is missing or blank"
      ))
  }

  parsed <- tryCatch(
    jsonlite::fromJSON(census_ref_raw, simplifyVector = FALSE),
    error = identity
  )

  if (inherits(parsed, "error")) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "malformed_json",
        parse_detail = conditionMessage(parsed)
      ))
  }

  if (!is.list(parsed)) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = cell_to_text(parsed),
        ref_token_trimmed = str_trim(cell_to_text(parsed)),
        parse_status = "malformed_non_array",
        parse_detail = "JSON value is not an array"
      ))
  }

  if (length(parsed) == 0L) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "malformed_empty_array",
        parse_detail = "JSON array contains no reference values"
      ))
  }

  map_dfr(seq_along(parsed), function(position) {
    element <- parsed[[position]]
    element_is_scalar <- !is.list(element) && length(element) == 1L
    token <- if (element_is_scalar) cell_to_text(element) else NA_character_
    token_trimmed <- if (!is.na(token)) str_trim(token) else NA_character_

    base_row %>%
      mutate(
        element_position = as.integer(position),
        ref_token = token,
        ref_token_trimmed = token_trimmed,
        parse_status = case_when(
          !element_is_scalar ~ "malformed_non_scalar",
          is.na(token_trimmed) || token_trimmed == "" ~ "malformed_empty_token",
          TRUE ~ "parsed"
        ),
        parse_detail = case_when(
          !element_is_scalar ~ "Array element is not a scalar",
          is.na(token_trimmed) || token_trimmed == "" ~
            "Array element is missing or blank",
          TRUE ~ NA_character_
        )
      )
  })
}

reference_tokens <- pmap_dfr(
  allocation %>%
    mutate(allocation_row = row_number()) %>%
    select(
      country, allocation_evidence_id = evidence_id, pair_id,
      census_ref_raw = census_ref, allocation_row
    ),
  parse_reference_cell
)

census_evidence_map <- census %>%
  mutate(
    evidence_lookup = str_trim(evidence_id),
    target_instrument = if_else(
      is_missing_value(instrument_id),
      NA_character_,
      str_trim(instrument_id)
    )
  ) %>%
  filter(!is_missing_value(evidence_lookup)) %>%
  group_by(country, evidence_lookup) %>%
  summarise(
    evidence_match_rows = n(),
    evidence_target_count = n_distinct(target_instrument, na.rm = TRUE),
    evidence_target = {
      targets <- unique(na.omit(target_instrument))
      if (length(targets) == 1L) targets[[1]] else NA_character_
    },
    .groups = "drop"
  )

census_instrument_map <- census %>%
  mutate(instrument_lookup = str_trim(instrument_id)) %>%
  filter(!is_missing_value(instrument_lookup)) %>%
  group_by(country, instrument_lookup) %>%
  summarise(
    instrument_match_rows = n(),
    instrument_target_count = n_distinct(instrument_lookup),
    instrument_target = {
      targets <- unique(na.omit(instrument_lookup))
      if (length(targets) == 1L) targets[[1]] else NA_character_
    },
    .groups = "drop"
  )

global_key_locations <- bind_rows(
  census %>%
    transmute(country, ref_token_trimmed = str_trim(evidence_id)),
  census %>%
    transmute(country, ref_token_trimmed = str_trim(instrument_id))
) %>%
  filter(!is_missing_value(ref_token_trimmed)) %>%
  distinct(country, ref_token_trimmed) %>%
  group_by(ref_token_trimmed) %>%
  summarise(
    matching_countries = paste(sort(unique(country)), collapse = "; "),
    .groups = "drop"
  )

reference_resolution <- reference_tokens %>%
  left_join(
    census_evidence_map,
    by = c("country", "ref_token_trimmed" = "evidence_lookup")
  ) %>%
  left_join(
    census_instrument_map,
    by = c("country", "ref_token_trimmed" = "instrument_lookup")
  ) %>%
  left_join(global_key_locations, by = "ref_token_trimmed") %>%
  mutate(
    evidence_match_rows = replace_na(evidence_match_rows, 0L),
    evidence_target_count = replace_na(evidence_target_count, 0L),
    instrument_match_rows = replace_na(instrument_match_rows, 0L),
    instrument_target_count = replace_na(instrument_target_count, 0L),
    namespace_overlap = evidence_match_rows > 0L & instrument_match_rows > 0L,
    namespace_same_target = namespace_overlap &
      evidence_target_count == 1L & instrument_target_count == 1L &
      coalesce(evidence_target == instrument_target, FALSE),
    token_country_prefix = str_match(
      ref_token_trimmed,
      "^([A-Z]{3})_"
    )[, 2],
    expected_identifier_form = str_detect(
      ref_token_trimmed,
      "^(?:[A-Z]{3}_.+_CEN_[0-9]+|[A-Z]{3}_INS_[0-9]+)$"
    ),
    other_country_match = !is.na(matching_countries) &
      !str_detect(
        paste0(";", matching_countries, ";"),
        fixed(paste0(";", country, ";"))
      ),
    resolution_status = case_when(
      parse_status != "parsed" ~ parse_status,
      evidence_match_rows > 1L ~ "ambiguous_evidence_id",
      evidence_match_rows == 1L & evidence_target_count == 0L ~
        "missing_target_instrument_id",
      evidence_match_rows == 1L & evidence_target_count > 1L ~
        "ambiguous_evidence_id",
      evidence_match_rows == 1L & namespace_overlap &
        !namespace_same_target ~ "id_namespace_collision",
      evidence_match_rows == 1L & evidence_target_count == 1L ~
        "resolved_evidence_id",
      evidence_match_rows == 0L & instrument_match_rows > 1L ~
        "ambiguous_instrument_id",
      evidence_match_rows == 0L & instrument_match_rows == 1L &
        instrument_target_count == 1L ~ "resolved_instrument_id",
      evidence_match_rows == 0L & instrument_match_rows == 1L &
        instrument_target_count != 1L ~ "ambiguous_instrument_id",
      !is.na(token_country_prefix) & token_country_prefix != country ~
        "cross_country_prefix",
      other_country_match ~ "cross_country_match",
      expected_identifier_form ~ "unmatched",
      TRUE ~ "malformed_identifier"
    ),
    resolved_instrument_id = case_when(
      resolution_status == "resolved_evidence_id" ~ evidence_target,
      resolution_status == "resolved_instrument_id" ~ instrument_target,
      TRUE ~ NA_character_
    ),
    successfully_resolved = resolution_status %in% c(
      "resolved_evidence_id", "resolved_instrument_id"
    ),
    resolution_detail = case_when(
      resolution_status == "resolved_evidence_id" & namespace_same_target ~
        "Evidence and instrument namespaces overlap on the same unique Census record; evidence-ID precedence used",
      resolution_status == "resolved_evidence_id" ~
        "Resolved by same-country Census evidence_id",
      resolution_status == "resolved_instrument_id" ~
        "Resolved by explicit same-country Census instrument_id fallback",
      resolution_status == "unmatched" ~
        "Well-formed identifier did not match a loaded Census record in any country",
      resolution_status == "malformed_identifier" ~
        "Parsed scalar does not follow the documented Census evidence/instrument ID form",
      str_starts(resolution_status, "cross_country") ~
        "Reference was not used because it points outside the allocation country",
      parse_status != "parsed" ~ parse_detail,
      TRUE ~ resolution_status
    )
  )

reference_denominator <- reference_resolution %>%
  count(country, name = "n_reference_values")

reference_resolution_summary <- reference_resolution %>%
  group_by(country) %>%
  summarise(
    n_reference_values = n(),
    n_resolved = sum(successfully_resolved),
    pct_resolved = n_resolved / n_reference_values,
    n_resolved_evidence_id = sum(resolution_status == "resolved_evidence_id"),
    n_resolved_instrument_id = sum(resolution_status == "resolved_instrument_id"),
    n_namespace_overlap_same_record = sum(
      successfully_resolved & namespace_same_target
    ),
    n_unmatched = sum(resolution_status == "unmatched"),
    n_malformed = sum(str_starts(resolution_status, "malformed")),
    n_cross_country = sum(str_starts(resolution_status, "cross_country")),
    n_ambiguous = sum(
      str_starts(resolution_status, "ambiguous") |
        resolution_status == "id_namespace_collision"
    ),
    .groups = "drop"
  ) %>%
  bind_rows(
    reference_resolution %>%
      summarise(
        country = "Overall",
        n_reference_values = n(),
        n_resolved = sum(successfully_resolved),
        pct_resolved = n_resolved / n_reference_values,
        n_resolved_evidence_id = sum(
          resolution_status == "resolved_evidence_id"
        ),
        n_resolved_instrument_id = sum(
          resolution_status == "resolved_instrument_id"
        ),
        n_namespace_overlap_same_record = sum(
          successfully_resolved & namespace_same_target
        ),
        n_unmatched = sum(resolution_status == "unmatched"),
        n_malformed = sum(str_starts(resolution_status, "malformed")),
        n_cross_country = sum(str_starts(resolution_status, "cross_country")),
        n_ambiguous = sum(
          str_starts(resolution_status, "ambiguous") |
            resolution_status == "id_namespace_collision"
        )
      )
  )

reference_resolution_issues <- reference_resolution %>%
  filter(!successfully_resolved) %>%
  select(
    country, allocation_row, allocation_evidence_id, pair_id,
    element_position, census_ref_raw, ref_token, ref_token_trimmed,
    resolution_status, resolution_detail, matching_countries
  ) %>%
  arrange(country, allocation_row, element_position)

resolved_links <- reference_resolution %>%
  filter(successfully_resolved, !is.na(resolved_instrument_id)) %>%
  distinct(country, instrument_id = resolved_instrument_id)

census <- census %>%
  left_join(
    resolved_links %>% mutate(is_earmarked = 1L),
    by = c("country", "instrument_id")
  ) %>%
  mutate(is_earmarked = replace_na(is_earmarked, 0L)) %>%
  select(all_of(census_columns), is_earmarked)
```

## Census-reference resolution

```{r print-ref-summary}
display_table(reference_resolution_summary)
```

```{r print-ref-issues}
display_table(reference_resolution_issues)
```

# 4. Sanity checks

```{r prepare-sanity-checks}
country_overlap <- full_join(
  census %>% distinct(country) %>% mutate(in_census = TRUE),
  allocation %>% distinct(country) %>% mutate(in_allocation = TRUE),
  by = "country"
) %>%
  mutate(
    in_census = replace_na(in_census, FALSE),
    in_allocation = replace_na(in_allocation, FALSE),
    dataset_presence = case_when(
      in_census & in_allocation ~ "both",
      in_census ~ "census_only",
      in_allocation ~ "allocation_only"
    )
  ) %>%
  arrange(country)

country_key_counts <- bind_rows(
  census %>%
    group_by(country) %>%
    summarise(
      dataset = "Census",
      n_rows = n(),
      n_distinct_evidence_id = n_distinct(
        evidence_id[!is_missing_value(evidence_id)]
      ),
      n_distinct_instrument_id = n_distinct(
        instrument_id[!is_missing_value(instrument_id)]
      ),
      n_distinct_pair_id = NA_integer_,
      n_distinct_document_id = n_distinct(
        document_id[!is_missing_value(document_id)]
      ),
      .groups = "drop"
    ),
  allocation %>%
    group_by(country) %>%
    summarise(
      dataset = "Allocation",
      n_rows = n(),
      n_distinct_evidence_id = n_distinct(
        evidence_id[!is_missing_value(evidence_id)]
      ),
      n_distinct_instrument_id = n_distinct(
        instrument_id[!is_missing_value(instrument_id)]
      ),
      n_distinct_pair_id = n_distinct(pair_id[!is_missing_value(pair_id)]),
      n_distinct_document_id = n_distinct(
        document_id[!is_missing_value(document_id)]
      ),
      .groups = "drop"
    )
) %>%
  select(dataset, country, everything()) %>%
  arrange(dataset, country)

missing_by_column <- bind_rows(
  map_dfr(names(census), function(field) {
    missing <- is_missing_value(census[[field]])
    tibble(
      dataset = "Census",
      field = field,
      n_rows = nrow(census),
      n_missing = sum(missing),
      pct_missing = sum(missing) / nrow(census)
    )
  }),
  map_dfr(names(allocation), function(field) {
    missing <- is_missing_value(allocation[[field]])
    tibble(
      dataset = "Allocation",
      field = field,
      n_rows = nrow(allocation),
      n_missing = sum(missing),
      pct_missing = sum(missing) / nrow(allocation)
    )
  })
)

missing_ids <- full_join(
  census %>%
    group_by(country) %>%
    summarise(
      missing_census_evidence_id_rows = sum(is_missing_value(evidence_id)),
      missing_census_instrument_id_rows = sum(is_missing_value(instrument_id)),
      .groups = "drop"
    ),
  allocation %>%
    group_by(country) %>%
    summarise(
      missing_allocation_evidence_id_rows = sum(is_missing_value(evidence_id)),
      missing_allocation_instrument_id_rows = sum(is_missing_value(instrument_id)),
      missing_allocation_pair_id_rows = sum(is_missing_value(pair_id)),
      missing_census_ref_rows = sum(is_missing_value(census_ref)),
      .groups = "drop"
    ),
  by = "country"
) %>%
  mutate(across(-country, ~ replace_na(.x, 0L))) %>%
  arrange(country)

duplicate_evidence_id <- bind_rows(
  census %>%
    filter(!is_missing_value(evidence_id)) %>%
    count(country, evidence_id, name = "n_rows") %>%
    filter(n_rows > 1L) %>%
    mutate(dataset = "Census", .before = 1),
  allocation %>%
    filter(!is_missing_value(evidence_id)) %>%
    count(country, evidence_id, name = "n_rows") %>%
    filter(n_rows > 1L) %>%
    mutate(dataset = "Allocation", .before = 1)
)

duplicate_census_instrument <- census %>%
  filter(!is_missing_value(instrument_id)) %>%
  count(country, instrument_id, name = "n_rows") %>%
  filter(n_rows > 1L) %>%
  arrange(country, instrument_id)

duplicate_allocation_pair <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  group_by(country, pair_id) %>%
  summarise(
    n_rows = n(),
    n_evidence_ids = n_distinct(evidence_id[!is_missing_value(evidence_id)]),
    evidence_ids = paste(sort(unique(evidence_id)), collapse = "; "),
    .groups = "drop"
  ) %>%
  filter(n_rows > 1L) %>%
  arrange(country, pair_id)

make_row_signature <- function(data) {
  vapply(seq_len(nrow(data)), function(i) {
    paste(
      vapply(data, function(column) {
        value <- column[[i]]
        if (is.na(value)) "<NA>" else as.character(value)
      }, character(1)),
      collapse = "\u241E"
    )
  }, character(1))
}

find_full_row_duplicates <- function(data, dataset) {
  signatures <- make_row_signature(data)
  signature_counts <- table(signatures)
  duplicated_signatures <- names(signature_counts[signature_counts > 1L])

  if (length(duplicated_signatures) == 0L) {
    return(tibble(
      dataset = character(), country = character(), source_row = integer(),
      evidence_id = character(), analytical_key = character(),
      n_identical_rows = integer()
    ))
  }

  tibble(
    dataset = dataset,
    country = data$country,
    source_row = seq_len(nrow(data)),
    evidence_id = data$evidence_id,
    analytical_key = if (dataset == "Census") data$instrument_id else data$pair_id,
    signature = signatures
  ) %>%
    filter(signature %in% duplicated_signatures) %>%
    mutate(n_identical_rows = as.integer(signature_counts[signature])) %>%
    select(-signature)
}

duplicate_full_rows <- bind_rows(
  find_full_row_duplicates(census, "Census"),
  find_full_row_duplicates(allocation, "Allocation")
)

duplicated_pair_rows <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  add_count(country, pair_id, name = "pair_row_count") %>%
  filter(pair_row_count > 1L) %>%
  select(-pair_row_count)

if (nrow(duplicated_pair_rows) == 0L) {
  pair_conflicts <- tibble(
    country = character(), pair_id = character(), field = character(),
    n_distinct_values = integer(), values = character(), n_rows = integer()
  )
} else {
  pair_conflicts <- duplicated_pair_rows %>%
    mutate(across(everything(), as.character)) %>%
    pivot_longer(
      cols = -c(country, pair_id, evidence_id),
      names_to = "field",
      values_to = "value"
    ) %>%
    mutate(value = if_else(is_missing_value(value), "Missing", value)) %>%
    group_by(country, pair_id, field) %>%
    summarise(
      n_distinct_values = n_distinct(value),
      values = paste(sort(unique(value)), collapse = " || "),
      n_rows = n(),
      .groups = "drop"
    ) %>%
    filter(n_distinct_values > 1L) %>%
    arrange(country, pair_id, field)
}
```

## Countries present

```{r print-country-overlap}
display_table(country_overlap)
```

## Rows and distinct keys by country

```{r print-country-key-counts}
display_table(country_key_counts)
```

## Missing values by column

```{r print-missing-by-column}
display_table(missing_by_column)
```

## Missing identifiers

```{r print-missing-ids}
display_table(missing_ids)
```

## Duplicate evidence IDs

```{r print-duplicate-evidence}
display_table(duplicate_evidence_id)
```

## Duplicate Census instrument IDs

```{r print-duplicate-census-instrument}
display_table(duplicate_census_instrument)
```

## Duplicate allocation pair IDs

```{r print-duplicate-allocation-pair}
display_table(duplicate_allocation_pair)
```

## Exact duplicate rows

```{r print-duplicate-full-rows}
display_table(duplicate_full_rows)
```

## Conflicting fields within `country + pair_id`

```{r print-pair-conflicts}
display_table(pair_conflicts)
```

## Category and document-year audits

```{r prepare-category-audit}
census_category_spec <- list(
  row_type = c("source"),
  official_name_underspecified = c("0", "1"),
  in_force = c("0", "1"),
  tax_instrument = c(
    "VAT", "excise", "fuel_levy", "telecom_levy", "environmental_tax",
    "payroll_tax", "resource_revenue", "redevance",
    "parafiscal_contribution", "fine", "other"
  ),
  base_sector = c(
    "A_agriculture_forestry_fishing", "B_mining_quarrying",
    "C_manufacturing", "D_electricity_gas", "E_water_waste",
    "F_construction", "G_trade", "H_transport_storage",
    "I_accommodation_food", "J_information_communication",
    "K_financial_insurance", "L_real_estate",
    "M_professional_technical", "N_admin_support",
    "O_public_administration", "P_education", "Q_health_social",
    "R_arts_entertainment", "S_other_services",
    "T_households_employers", "U_extraterritorial",
    "general_economy_wide", "not_applicable"
  ),
  base_division = "REGEX_01_TO_99",
  is_penalty = c("0", "1"),
  rate_is_schedule = c("0", "1"),
  rate_type = c(
    "ad_valorem_pct", "per_unit_amount", "fixed_amount",
    "amount_per_unit_of_base", "rate_point_increment"
  ),
  rate_basis = c(
    "stated", "banded", "deferred_subordinate_act", "cross_reference",
    "ceiling_delegated", "floor_delegated",
    "increment_on_other_instrument", "not_applicable"
  ),
  change_type = c(
    "new", "amendment_rate", "amendment_base", "amendment_key",
    "merger", "split", "replacement", "repeal", "none"
  ),
  evidence_status = c("attested", "referenced_not_enacted_here")
)

allocation_category_spec <- list(
  row_type = c("allocation"),
  beneficiary_type = c(
    "fund", "agency", "ministry", "collectivite_territoriale", "sector",
    "public_employees", "program", "supranational", "general_budget"
  ),
  allocation_nature = c("proceeds_share", "cost_recovery_component"),
  share_is_schedule = c("0", "1"),
  share_type = c("pct", "per_unit_amount", "fixed_amount"),
  share_basis = c("stated", "whole_proceeds_implied", "deferred_arrete"),
  is_purpose_restricted = c("0", "1"),
  assignment_type = c(
    "earmark_functional", "hybrid_devolved_earmark",
    "equalization_transfer", "community_levy_external",
    "tax_sharing_specific", "tax_sharing_general", "derivation_transfer"
  ),
  destination_function_basis = c("stated", "named_body", "inferred"),
  destination_function_multi = c("0", "1")
)

audit_category_field <- function(data, dataset, field, expected) {
  values <- data[[field]]
  displayed <- if_else(is_missing_value(values), "Missing", as.character(values))

  tibble(value = displayed) %>%
    count(value, name = "n_rows") %>%
    mutate(
      dataset = dataset,
      field = field,
      expected_status = case_when(
        value == "Missing" ~ "missing",
        identical(expected, "REGEX_01_TO_99") &
          str_detect(value, "^(?:0[1-9]|[1-9][0-9])$") ~ "expected",
        identical(expected, "REGEX_01_TO_99") ~ "unexpected",
        value %in% expected ~ "expected",
        TRUE ~ "unexpected"
      ),
      .before = 1
    )
}

category_audit <- bind_rows(
  imap_dfr(census_category_spec, ~ audit_category_field(
    census, "Census", .y, .x
  )),
  imap_dfr(allocation_category_spec, ~ audit_category_field(
    allocation, "Allocation", .y, .x
  ))
) %>%
  arrange(dataset, field, expected_status, value)

sort_year_values <- function(values) {
  values <- values[!is_missing_value(values)] %>% unique()
  if (length(values) == 0L) {
    return(NA_character_)
  }
  numeric_year <- suppressWarnings(as.integer(values))
  order_index <- order(is.na(numeric_year), numeric_year, values)
  paste(values[order_index], collapse = "; ")
}

document_years <- bind_rows(
  census %>% mutate(dataset = "Census", .before = 1),
  allocation %>% mutate(dataset = "Allocation", .before = 1)
) %>%
  group_by(dataset, country) %>%
  summarise(
    n_rows = n(),
    document_years = sort_year_values(document_year),
    min_document_year = {
      years <- suppressWarnings(as.integer(document_year))
      if (all(is.na(years))) NA_integer_ else min(years, na.rm = TRUE)
    },
    max_document_year = {
      years <- suppressWarnings(as.integer(document_year))
      if (all(is.na(years))) NA_integer_ else max(years, na.rm = TRUE)
    },
    n_missing_document_year = sum(is_missing_value(document_year)),
    n_unparseable_document_year = sum(
      !is_missing_value(document_year) &
        !str_detect(document_year, "^[0-9]{4}$")
    ),
    unparseable_values = {
      bad <- unique(document_year[
        !is_missing_value(document_year) &
          !str_detect(document_year, "^[0-9]{4}$")
      ])
      if (length(bad) == 0L) NA_character_ else paste(sort(bad), collapse = "; ")
    },
    .groups = "drop"
  ) %>%
  arrange(dataset, country)
```

```{r print-category-audit}
display_table(category_audit)
```

```{r print-document-years}
display_table(document_years)
```

# 5. Country and legal-corpus coverage

```{r prepare-corpus-coverage}
make_corpus_coverage <- function(data) {
  data %>%
    group_by(country) %>%
    summarise(
      n_documents = n_distinct(document_id[!is_missing_value(document_id)]),
      document_years = sort_year_values(document_year),
      corpus_type = case_when(
        n_documents == 1L ~ "single_document",
        n_documents > 1L ~ "multi_document",
        TRUE ~ NA_character_
      ),
      .groups = "drop"
    ) %>%
    arrange(country)
}

census_coverage <- make_corpus_coverage(census)
allocation_coverage <- make_corpus_coverage(allocation)
```

## Census corpus

```{r print-census-coverage}
display_table(census_coverage)
```

## Allocation corpus

```{r print-allocation-coverage}
display_table(allocation_coverage)
```

# 6. Database at a glance

```{r prepare-database-glance}
country_universe <- sort(unique(c(census$country, allocation$country)))

database_glance <- tibble(country = country_universe) %>%
  left_join(
    census %>%
      filter(!is_missing_value(instrument_id)) %>%
      distinct(country, instrument_id) %>%
      count(country, name = "D"),
    by = "country"
  ) %>%
  left_join(
    resolved_links %>% count(country, name = "E"),
    by = "country"
  ) %>%
  left_join(
    allocation %>%
      filter(!is_missing_value(pair_id)) %>%
      distinct(country, pair_id) %>%
      count(country, name = "P"),
    by = "country"
  ) %>%
  mutate(
    across(c(D, E, P), ~ replace_na(.x, 0L)),
    earmark_incidence = if_else(D > 0L, E / D, NA_real_),
    allocation_density = if_else(E > 0L, P / E, NA_real_)
  )
```

```{r print-database-glance}
display_table(database_glance)
```

# 7. Global composition of the allocation layer

The unit is a distinct `country + pair_id`. No strict-earmark, in-force, penalty, or other analytical filter is imposed. Missing category values are represented as `Missing`.

For each non-destination category, duplicate allocation rows must agree on the category. The notebook stops rather than selecting a value if one pair has conflicting values in that field.

```{r prepare-composition}
allocation_country_totals <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  distinct(country, pair_id) %>%
  count(country, name = "country_total_allocations")

allocation_countries <- allocation %>%
  distinct(country) %>%
  arrange(country)

make_composition <- function(field) {
  pair_categories <- allocation %>%
    filter(!is_missing_value(pair_id)) %>%
    transmute(
      country,
      pair_id,
      category = if_else(
        is_missing_value(.data[[field]]),
        "Missing",
        as.character(.data[[field]])
      )
    ) %>%
    distinct()

  category_conflicts <- pair_categories %>%
    count(country, pair_id, name = "n_categories") %>%
    filter(n_categories > 1L)

  if (nrow(category_conflicts) > 0L) {
    print(category_conflicts)
    stop(
      "Conflicting values within country + pair_id for composition field: ",
      field
    )
  }

  country_category_counts <- pair_categories %>%
    count(country, category, name = "n_allocations")

  complete_counts <- expand_grid(
    country = allocation_countries$country,
    category = sort(unique(pair_categories$category))
  ) %>%
    left_join(country_category_counts, by = c("country", "category")) %>%
    mutate(n_allocations = replace_na(n_allocations, 0L)) %>%
    left_join(allocation_country_totals, by = "country") %>%
    mutate(
      within_country_share = if_else(
        country_total_allocations > 0L,
        n_allocations / country_total_allocations,
        NA_real_
      )
    )

  complete_counts %>%
    group_by(category) %>%
    summarise(
      pooled_distinct_allocations = sum(n_allocations),
      countries_with_category = sum(n_allocations > 0L),
      median_allocations_among_users = if_else(
        any(n_allocations > 0L),
        median(n_allocations[n_allocations > 0L]),
        NA_real_
      ),
      mean_within_country_share = if_else(
        any(is.na(within_country_share)),
        NA_real_,
        mean(within_country_share)
      ),
      .groups = "drop"
    ) %>%
    arrange(desc(pooled_distinct_allocations), category)
}

composition_assignment <- make_composition("assignment_type")
composition_nature <- make_composition("allocation_nature")
composition_purpose <- make_composition("is_purpose_restricted")
composition_beneficiary <- make_composition("beneficiary_type")
```

## Assignment type

```{r print-composition-assignment}
display_table(composition_assignment)
```

## Allocation nature

```{r print-composition-nature}
display_table(composition_nature)
```

## Purpose restriction

```{r print-composition-purpose}
display_table(composition_purpose)
```

## Beneficiary type

```{r print-composition-beneficiary}
display_table(composition_beneficiary)
```

## Destination functions

The long destination-function table combines the primary function and every semicolon-delimited additional function, trims lookup values, removes empty values, and counts each `country + pair_id + destination_function` combination once. One allocation may finance several functions; therefore destination-function counts can exceed the total number of allocations and their mean within-country shares can sum to more than 100 percent.

```{r prepare-destination-functions}
cofog_path <- file.path(project_root, "data", "COFOG_GFSM2014_mapping.csv")

cofog_mapping <- read_delim(
  cofog_path,
  delim = ";",
  col_types = cols(.default = col_character()),
  trim_ws = FALSE,
  show_col_types = FALSE
) %>%
  select(
    destination_function = cofog_code,
    mapping_level = level,
    destination_function_name = cofog_name
  )

stopifnot(
  all(c("destination_function", "destination_function_name") %in%
        names(cofog_mapping)),
  nrow(cofog_mapping) == n_distinct(cofog_mapping$destination_function)
)

special_function_labels <- tribble(
  ~destination_function, ~destination_function_name, ~mapping_level,
  "general_budget", "General budget", "non_GFSM",
  "local_budget", "Local budget", "non_GFSM",
  "unallocated", "Unallocated", "non_GFSM",
  "Missing", "Missing", "missing"
)

primary_functions <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  transmute(
    country,
    pair_id,
    destination_function = if_else(
      is_missing_value(destination_function),
      NA_character_,
      str_trim(as.character(destination_function))
    ),
    source = "primary"
  ) %>%
  filter(!is.na(destination_function), destination_function != "")

detail_functions <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  transmute(
    country,
    pair_id,
    destination_function_detail = if_else(
      is_missing_value(destination_function_detail),
      NA_character_,
      as.character(destination_function_detail)
    )
  ) %>%
  separate_longer_delim(destination_function_detail, delim = ";") %>%
  transmute(
    country,
    pair_id,
    destination_function = str_trim(destination_function_detail),
    source = "detail"
  ) %>%
  filter(!is.na(destination_function), destination_function != "")

function_links_present <- bind_rows(primary_functions, detail_functions) %>%
  group_by(country, pair_id, destination_function) %>%
  summarise(
    source = paste(sort(unique(source)), collapse = "; "),
    .groups = "drop"
  )

pairs_without_function <- allocation %>%
  filter(!is_missing_value(pair_id)) %>%
  distinct(country, pair_id) %>%
  anti_join(
    function_links_present %>% distinct(country, pair_id),
    by = c("country", "pair_id")
  ) %>%
  mutate(
    destination_function = "Missing",
    source = "missing"
  )

destination_function_long <- bind_rows(
  function_links_present,
  pairs_without_function
) %>%
  left_join(cofog_mapping, by = "destination_function") %>%
  left_join(
    special_function_labels,
    by = "destination_function",
    suffix = c("", "_special")
  ) %>%
  mutate(
    destination_function_name = coalesce(
      destination_function_name,
      destination_function_name_special
    ),
    mapping_level = coalesce(mapping_level, mapping_level_special),
    destination_function_label = case_when(
      destination_function == "Missing" ~ "Missing",
      is.na(destination_function_name) ~ destination_function,
      TRUE ~ paste0(
        destination_function,
        " - ",
        destination_function_name
      )
    )
  ) %>%
  select(
    country, pair_id, destination_function, destination_function_name,
    destination_function_label, mapping_level, source
  ) %>%
  arrange(country, pair_id, destination_function)

function_map_audit <- destination_function_long %>%
  count(
    destination_function, destination_function_name, mapping_level,
    name = "n_distinct_pair_function_links"
  ) %>%
  mutate(
    mapping_status = case_when(
      destination_function == "Missing" ~ "missing",
      mapping_level == "non_GFSM" ~ "preserved_non_GFSM",
      mapping_level == "group" ~ "mapped_GFSM_group",
      !is.na(mapping_level) ~ "mapped_non_group_code",
      TRUE ~ "unmapped"
    )
  ) %>%
  arrange(mapping_status, destination_function)

function_country_counts <- destination_function_long %>%
  count(
    country, destination_function, destination_function_name,
    destination_function_label,
    name = "n_allocations"
  )

function_categories <- destination_function_long %>%
  distinct(
    destination_function, destination_function_name,
    destination_function_label
  )

function_complete_counts <- crossing(
  allocation_countries,
  function_categories
) %>%
  left_join(
    function_country_counts,
    by = c(
      "country", "destination_function", "destination_function_name",
      "destination_function_label"
    )
  ) %>%
  mutate(n_allocations = replace_na(n_allocations, 0L)) %>%
  left_join(allocation_country_totals, by = "country") %>%
  mutate(
    within_country_share = if_else(
      country_total_allocations > 0L,
      n_allocations / country_total_allocations,
      NA_real_
    )
  )

composition_destination <- function_complete_counts %>%
  group_by(
    destination_function, destination_function_name,
    destination_function_label
  ) %>%
  summarise(
    pooled_distinct_allocations = sum(n_allocations),
    countries_with_category = sum(n_allocations > 0L),
    median_allocations_among_users = if_else(
      any(n_allocations > 0L),
      median(n_allocations[n_allocations > 0L]),
      NA_real_
    ),
    mean_within_country_share = if_else(
      any(is.na(within_country_share)),
      NA_real_,
      mean(within_country_share)
    ),
    .groups = "drop"
  ) %>%
  arrange(desc(pooled_distinct_allocations), destination_function)
```

```{r print-destination-long}
display_table(destination_function_long)
```

```{r print-function-map-audit}
display_table(function_map_audit)
```

```{r print-composition-destination}
display_table(composition_destination)
```

# 8. Filter and transformation audit

The final pooled CSVs retain every row from every successfully loaded file. The rows excluded from analytical key universes are reported below; no source observation is silently removed.

```{r prepare-filter-audit}
filter_audit <- bind_rows(
  tibble(
    step = "Census source-file validation",
    dataset = "Census",
    condition = "Keep only files with every required column, compatible values, and matching folder/workbook country",
    rows_before = sum(schema_audit$source_rows[schema_audit$dataset == "Census"], na.rm = TRUE),
    rows_after = nrow(census),
    relevant_keys_before = NA_integer_,
    relevant_keys_after = n_distinct(
      paste(census$country, census$instrument_id),
      na.rm = TRUE
    )
  ),
  tibble(
    step = "Allocation source-file validation",
    dataset = "Allocation",
    condition = "Keep only files with every required column, compatible values, and matching folder/workbook country",
    rows_before = sum(schema_audit$source_rows[schema_audit$dataset == "Allocation"], na.rm = TRUE),
    rows_after = nrow(allocation),
    relevant_keys_before = NA_integer_,
    relevant_keys_after = n_distinct(
      paste(allocation$country, allocation$pair_id),
      na.rm = TRUE
    )
  ),
  tibble(
    step = "Census analytical key universe",
    dataset = "Census",
    condition = "Exclude rows with missing country or instrument_id from distinct-key counts only",
    rows_before = nrow(census),
    rows_after = sum(
      !is_missing_value(census$country) &
        !is_missing_value(census$instrument_id)
    ),
    relevant_keys_before = n_distinct(
      paste(census$country, census$instrument_id),
      na.rm = TRUE
    ),
    relevant_keys_after = n_distinct(
      paste(
        census$country[
          !is_missing_value(census$country) &
            !is_missing_value(census$instrument_id)
        ],
        census$instrument_id[
          !is_missing_value(census$country) &
            !is_missing_value(census$instrument_id)
        ]
      )
    )
  ),
  tibble(
    step = "Allocation analytical key universe",
    dataset = "Allocation",
    condition = "Exclude rows with missing country or pair_id from distinct allocation counts only",
    rows_before = nrow(allocation),
    rows_after = sum(
      !is_missing_value(allocation$country) &
        !is_missing_value(allocation$pair_id)
    ),
    relevant_keys_before = n_distinct(
      paste(allocation$country, allocation$pair_id),
      na.rm = TRUE
    ),
    relevant_keys_after = n_distinct(
      paste(
        allocation$country[
          !is_missing_value(allocation$country) &
            !is_missing_value(allocation$pair_id)
        ],
        allocation$pair_id[
          !is_missing_value(allocation$country) &
            !is_missing_value(allocation$pair_id)
        ]
      )
    )
  ),
  tibble(
    step = "Successful census_ref links",
    dataset = "Allocation to Census",
    condition = "Retain only unambiguous same-country resolved reference values for E and is_earmarked",
    rows_before = nrow(reference_resolution),
    rows_after = sum(reference_resolution$successfully_resolved),
    relevant_keys_before = n_distinct(
      paste(
        reference_resolution$country,
        reference_resolution$ref_token_trimmed
      ),
      na.rm = TRUE
    ),
    relevant_keys_after = nrow(resolved_links)
  )
)
```

```{r print-filter-audit}
display_table(filter_audit)
```

# 9. Save pooled data and analytical tables

```{r save-and-verify-csv}
write_csv(census, census_output_path, na = "")
write_csv(allocation, allocation_output_path, na = "")

census_saved_check <- read_csv(
  census_output_path,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)
allocation_saved_check <- read_csv(
  allocation_output_path,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

stopifnot(
  nrow(census_saved_check) == nrow(census),
  nrow(allocation_saved_check) == nrow(allocation),
  identical(names(census_saved_check), c(census_columns, "is_earmarked")),
  identical(names(allocation_saved_check), allocation_columns)
)

stack_reconciliation <- stack_reconciliation %>%
  mutate(
    saved_csv_rows = if_else(
      dataset == "Census",
      nrow(census_saved_check),
      nrow(allocation_saved_check)
    ),
    csv_rows_reconcile = saved_csv_rows == stacked_rows,
    exact_output_columns = if_else(
      dataset == "Census",
      identical(names(census_saved_check), c(census_columns, "is_earmarked")),
      identical(names(allocation_saved_check), allocation_columns)
    )
  )

stopifnot(
  all(stack_reconciliation$csv_rows_reconcile),
  all(stack_reconciliation$exact_output_columns)
)
```

## Final row-count reconciliation

```{r print-final-reconciliation}
display_table(stack_reconciliation)
```

```{r prepare-run-summary}
overall_resolution <- reference_resolution_summary %>%
  filter(country == "Overall")

unexpected_category_count <- category_audit %>%
  filter(expected_status == "unexpected") %>%
  summarise(n = sum(n_rows)) %>%
  pull(n)

unmapped_function_count <- function_map_audit %>%
  filter(mapping_status == "unmapped") %>%
  summarise(n = sum(n_distinct_pair_function_links)) %>%
  pull(n)

if (length(unmapped_function_count) == 0L) unmapped_function_count <- 0L

run_summary <- tribble(
  ~section, ~metric, ~value, ~note,
  "Scope", "Database interpretation", "de jure legal inventory", "Not collections, transfers, disbursements, or expenditure execution",
  "Scope", "Census analytical key", "country + instrument_id", "Cross-country IDs are never pooled as a single key",
  "Scope", "Allocation analytical key", "country + pair_id", "One canonical statutory allocation channel",
  "Files", "Census files found", as.character(sum(file_inventory$dataset == "Census")), "Canonical final XLSX candidates",
  "Files", "Census files loaded", as.character(sum(file_inventory$dataset == "Census" & file_inventory$final_status == "loaded")), "Schema- and type-valid files",
  "Files", "Allocation files found", as.character(sum(file_inventory$dataset == "Allocation")), "Canonical final XLSX candidates",
  "Files", "Allocation files loaded", as.character(sum(file_inventory$dataset == "Allocation" & file_inventory$final_status == "loaded")), "Schema- and type-valid files",
  "Rows", "Census pooled rows", as.character(nrow(census)), "Final CSV includes is_earmarked",
  "Rows", "Allocation pooled rows", as.character(nrow(allocation)), "Broad canonical allocation layer",
  "Countries", "Census countries", as.character(n_distinct(census$country)), "COD Census is excluded because its final XLSX lacks base_division",
  "Countries", "Allocation countries", as.character(n_distinct(allocation$country)), "All valid final allocation files",
  "References", "Flattened census_ref values", as.character(overall_resolution$n_reference_values), "Occurrence-level denominator",
  "References", "Successfully resolved values", as.character(overall_resolution$n_resolved), "Unambiguous same-country links only",
  "References", "Percent resolved", format(overall_resolution$pct_resolved, digits = 8), "Stored as a proportion",
  "References", "Direct instrument_id fallback resolutions", as.character(overall_resolution$n_resolved_instrument_id), "Explicitly supported after evidence-ID matching",
  "References", "Namespace-overlap resolutions", as.character(overall_resolution$n_namespace_overlap_same_record), "Evidence_id and instrument_id identify the same unique row",
  "Warnings", "Unresolved reference values", as.character(overall_resolution$n_reference_values - overall_resolution$n_resolved), "See ref_resolution_issues",
  "Warnings", "Duplicated allocation pair groups", as.character(nrow(duplicate_allocation_pair)), "See dup_allocation_pair and pair_conflicts",
  "Warnings", "Unexpected category observations", as.character(unexpected_category_count), "Values are reported, not recoded",
  "Warnings", "Unmapped destination-function links", as.character(unmapped_function_count), "GFSM codes are joined to COFOG mapping; specials preserved",
  "Destination functions", "Non-additivity note", "Counts and shares can exceed allocation totals or 100%", "One allocation may finance several functions",
  "Sources", "Source-version rule", "Final XLSX files are authoritative", "Intermediate or same-stem JSON files are not substituted"
)
```

## Run summary

```{r print-run-summary}
display_table(run_summary)
```

```{r stage-workbook-tables, include=FALSE}
workbook_tables <- list(
  run_summary = run_summary,
  file_status_summary = file_status_summary,
  file_inventory = file_inventory,
  schema_audit = schema_audit,
  type_audit = type_audit,
  stack_reconciliation = stack_reconciliation,
  filter_audit = filter_audit,
  country_overlap = country_overlap,
  country_key_counts = country_key_counts,
  missing_by_column = missing_by_column,
  missing_ids = missing_ids,
  dup_evidence_id = duplicate_evidence_id,
  dup_census_instrument = duplicate_census_instrument,
  dup_allocation_pair = duplicate_allocation_pair,
  dup_full_rows = duplicate_full_rows,
  pair_conflicts = pair_conflicts,
  category_audit = category_audit,
  document_years = document_years,
  ref_resolution_summary = reference_resolution_summary,
  ref_resolution_issues = reference_resolution_issues,
  census_coverage = census_coverage,
  allocation_coverage = allocation_coverage,
  database_glance = database_glance,
  destination_long = destination_function_long,
  function_map_audit = function_map_audit,
  comp_assignment = composition_assignment,
  comp_nature = composition_nature,
  comp_purpose = composition_purpose,
  comp_beneficiary = composition_beneficiary,
  comp_destination = composition_destination
)

serialize_table <- function(data, name) {
  rows <- map(seq_len(nrow(data)), function(i) {
    set_names(
      map(data, function(column) {
        value <- column[[i]]
        if (length(value) == 0L || is.na(value)) NA else unname(value)
      }),
      names(data)
    )
  })

  list(
    name = name,
    columns = names(data),
    rows = rows
  )
}

workbook_manifest <- list(
  title = "SSA revenue instruments: initial descriptives",
  generated_at = format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"),
  output_path = normalizePath(
    workbook_output_path,
    winslash = "/",
    mustWork = FALSE
  ),
  sheets = imap(workbook_tables, serialize_table)
)

write_json(
  workbook_manifest,
  manifest_path,
  auto_unbox = TRUE,
  pretty = TRUE,
  na = "null",
  digits = NA
)
```

The workbook is authored from the staged tables with the project’s spreadsheet tooling and is verified separately after this notebook completes.
