---
title: "SSA revenue instruments: stack and prepare the data"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook stacks the final country-level XLSX workbooks, applies the requested analytical filters, expands Census links, performs essential sanity checks, and reports legal-corpus coverage. The data describe the **de jure** legal system; they do not measure collections, transfers, disbursements, or expenditure execution. No plots or broader descriptive statistics are produced.

All identifiers are country-specific. Census instruments are keyed by `country + instrument_id`, allocation channels by `country + pair_id`, and expanded links by `country + pair_id + census_instrument_id`. References are never matched across countries.

```{r setup, include=FALSE}
input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(readxl)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(readr)
library(tibble)
library(jsonlite)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

country_root <- file.path(project_root, "countries")
derived_dir <- file.path(project_root, "data", "derived")
tables_dir <- file.path(project_root, "analysis", "tables")
dir.create(derived_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(tables_dir, recursive = TRUE, showWarnings = FALSE)

output_paths <- list(
  census_stacked = file.path(derived_dir, "SSA_CENSUS_STACKED.csv"),
  allocation_stacked = file.path(
    derived_dir, "SSA_EARMARK_ALLOCATIONS_STACKED.csv"
  ),
  census = file.path(derived_dir, "SSA_CENSUS.csv"),
  allocation = file.path(derived_dir, "SSA_EARMARK_ALLOCATIONS.csv"),
  allocation_expanded = file.path(
    derived_dir, "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  ),
  corpus_coverage = file.path(
    tables_dir, "01_country_corpus_coverage.csv"
  ),
  sanity_workbook = file.path(
    tables_dir, "01_stack_sanity_checks.xlsx"
  )
)

manifest_path <- Sys.getenv(
  "STACK_SANITY_MANIFEST",
  unset = file.path(tempdir(), "01_stack_sanity_checks_manifest.json")
)
dir.create(dirname(manifest_path), recursive = TRUE, showWarnings = FALSE)

census_columns <- c(
  "evidence_id", "row_type", "country", "instrument_id", "document_id",
  "document_type", "document_year", "legal_article", "page_start", "page_end",
  "official_name", "official_name_underspecified", "in_force",
  "tax_instrument", "tax_base_detail", "instrument_label_verbatim",
  "instrument_nature", "base_sector", "is_penalty", "rate_value",
  "rate_is_schedule", "rate_schedule_detail", "rate_type", "rate_basis",
  "rate_base_quantum", "rate_reference_instrument", "change_type",
  "enabling_reference", "evidence_status"
)

allocation_columns <- c(
  "evidence_id", "row_type", "country", "instrument_id", "pair_id",
  "document_id", "document_type", "document_year", "legal_article",
  "page_start", "page_end", "official_name", "destination",
  "beneficiary_type", "allocation_nature", "share_value",
  "share_is_schedule", "share_schedule_detail", "share_type", "share_basis",
  "share_level", "share_pool", "is_purpose_restricted", "assignment_type",
  "census_ref", "destination_function", "destination_function_basis",
  "destination_function_multi", "destination_function_detail"
)

required_columns <- list(
  Census = census_columns,
  Allocation = allocation_columns
)

required_sheets <- c(
  Census = "census_rows",
  Allocation = "evidence_table"
)

numeric_fields <- c("rate_value", "rate_base_quantum", "share_value")
integer_fields <- c("page_start", "page_end", "share_level")
binary_fields <- c(
  "official_name_underspecified", "in_force", "is_penalty",
  "rate_is_schedule", "share_is_schedule", "is_purpose_restricted",
  "destination_function_multi"
)
flexible_text_fields <- c(
  "document_year", "destination_function", "destination_function_detail"
)

collapse_or_na <- function(values, separator = "; ") {
  values <- unique(values[!is.na(values) & values != ""])
  if (length(values) == 0L) NA_character_ else paste(values, collapse = separator)
}

is_blank_cell <- function(x) {
  length(x) == 0L || all(is.na(x)) ||
    (is.character(x) && length(x) == 1L && str_trim(x) == "")
}

cell_storage_type <- function(x) {
  if (is_blank_cell(x)) return(NA_character_)
  class(x)[1]
}

cell_to_text <- function(x) {
  if (is_blank_cell(x) || length(x) != 1L) return(NA_character_)
  if (inherits(x, "POSIXt")) {
    return(format(x, "%Y-%m-%d %H:%M:%S", tz = "UTC"))
  }
  if (is.character(x)) return(x)
  if (is.numeric(x)) {
    return(format(x, scientific = FALSE, trim = TRUE, digits = 15))
  }
  if (is.logical(x)) return(ifelse(x, "TRUE", "FALSE"))
  as.character(x)
}

is_missing_value <- function(x) {
  if (is.character(x)) is.na(x) | str_trim(x) == "" else is.na(x)
}

is_numeric_text <- function(x) {
  !is.na(x) && str_detect(
    x,
    "^[+-]?(?:[0-9]+\\.?[0-9]*|\\.[0-9]+)(?:[eE][+-]?[0-9]+)?$"
  ) && is.finite(suppressWarnings(as.numeric(x)))
}

is_integer_text <- function(x) {
  if (!is_numeric_text(x)) return(FALSE)
  value <- suppressWarnings(as.numeric(x))
  abs(value - round(value)) < 1e-9
}

is_binary_text <- function(x) {
  !is.na(x) && tolower(x) %in% c("0", "1", "true", "false")
}

target_type_for <- function(field) {
  case_when(
    field %in% numeric_fields ~ "numeric",
    field %in% integer_fields ~ "integer",
    field %in% binary_fields ~ "binary_0_1",
    TRUE ~ "text"
  )
}

cell_is_compatible <- function(x, field, target_type) {
  if (is_blank_cell(x)) return(TRUE)
  storage_type <- cell_storage_type(x)
  text_value <- cell_to_text(x)

  if (target_type == "text") {
    if (field %in% flexible_text_fields) {
      return(storage_type %in% c("character", "numeric", "integer"))
    }
    return(storage_type == "character")
  }
  if (target_type == "numeric") return(is_numeric_text(text_value))
  if (target_type == "integer") return(is_integer_text(text_value))
  if (target_type == "binary_0_1") return(is_binary_text(text_value))
  FALSE
}

convert_list_column <- function(column, field, target_type) {
  text_values <- map_chr(column, cell_to_text)

  if (target_type == "text") return(text_values)
  if (target_type == "numeric") {
    return(ifelse(
      is.na(text_values), NA_real_, suppressWarnings(as.numeric(text_values))
    ))
  }
  if (target_type == "integer") {
    return(ifelse(
      is.na(text_values),
      NA_integer_,
      as.integer(round(suppressWarnings(as.numeric(text_values))))
    ))
  }
  if (target_type == "binary_0_1") {
    return(case_when(
      is.na(text_values) ~ NA_integer_,
      tolower(text_values) %in% c("1", "true") ~ 1L,
      tolower(text_values) %in% c("0", "false") ~ 0L,
      TRUE ~ NA_integer_
    ))
  }
  stop("Unknown target type: ", target_type)
}

n_distinct_country_key <- function(data, key_columns) {
  required <- c("country", key_columns)
  valid <- rep(TRUE, nrow(data))
  for (field in required) {
    valid <- valid & !is_missing_value(data[[field]])
  }
  if (!any(valid)) return(0L)
  nrow(distinct(data[valid, required, drop = FALSE]))
}

display_table <- function(data, digits = 4) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    knitr::kable(data, digits = digits, row.names = FALSE)
  }
}
```

# 1. Discover and validate final country files

The recursive search is restricted to the canonical country-final locations identified in `AGENTS.md`, plus the consolidated Anglophone Census location used by Ghana and South Africa: `countries/<ISO>/outputs/extract/` for existing Census files, `countries/<ISO>/outputs/census/*_ALL_CENSUS.xlsx` for consolidated Census files, and `countries/<ISO>/outputs/allocations/` for Allocation files. Other output, derived, intermediate, backup, and Excel lock files are not candidates.

```{r discover-files}
all_country_xlsx <- list.files(
  country_root,
  pattern = "\\.xlsx$",
  recursive = TRUE,
  full.names = TRUE,
  ignore.case = TRUE
)

project_root_normalized <- normalizePath(
  project_root, winslash = "/", mustWork = TRUE
)
all_country_paths <- normalizePath(
  all_country_xlsx, winslash = "/", mustWork = TRUE
)
relative_paths <- str_remove(
  all_country_paths,
  fixed(paste0(project_root_normalized, "/"))
)

candidate_inventory <- tibble(
  absolute_path = all_country_paths,
  relative_path = relative_paths,
  file_name = basename(relative_paths),
  country = str_match(relative_paths, "^countries/([^/]+)/")[, 2],
  dataset = case_when(
    str_detect(
      relative_paths,
      regex(
        paste0(
          "^countries/[^/]+/outputs/",
          "(?:extract/[^/]+_CENSUS|census/[^/]+_ALL_CENSUS)\\.xlsx$"
        ),
        ignore_case = TRUE
      )
    ) ~ "Census",
    str_detect(
      relative_paths,
      regex(
        paste0(
          "^countries/[^/]+/outputs/allocations/",
          "[^/]+_EARMARKS?_ALLOCATIONS?\\.xlsx$"
        ),
        ignore_case = TRUE
      )
    ) ~ "Allocation",
    TRUE ~ NA_character_
  ),
  is_temporary = str_starts(file_name, fixed("~$"))
) %>%
  filter(!is.na(dataset), !is_temporary) %>%
  mutate(required_sheet = unname(required_sheets[dataset])) %>%
  arrange(country, dataset, relative_path)

if (nrow(candidate_inventory) == 0L) {
  stop("No canonical final country XLSX files were found.")
}

duplicate_candidates <- candidate_inventory %>%
  count(country, dataset, name = "candidate_count") %>%
  filter(candidate_count > 1L)

if (nrow(duplicate_candidates) > 0L) {
  duplicate_detail <- candidate_inventory %>%
    semi_join(duplicate_candidates, by = c("country", "dataset")) %>%
    select(country, dataset, relative_path)
  print(duplicate_detail)
  stop(
    "Several plausible final files exist for at least one country and dataset. ",
    "No outputs were written."
  )
}

candidate_countries <- sort(unique(candidate_inventory$country))
country_dataset_presence_preflight <- expand_grid(
  country = candidate_countries,
  dataset = c("Census", "Allocation")
) %>%
  left_join(
    candidate_inventory %>% count(country, dataset, name = "n_files"),
    by = c("country", "dataset")
  ) %>%
  mutate(n_files = replace_na(n_files, 0L))

missing_country_dataset_preflight <- country_dataset_presence_preflight %>%
  filter(n_files == 0L)

if (nrow(missing_country_dataset_preflight) > 0L) {
  print(missing_country_dataset_preflight)
  stop(
    "At least one country has only one of the two required datasets. ",
    "No outputs were written."
  )
}

if (!"COD" %in% candidate_countries) {
  stop("COD is missing from the final-file inventory. No outputs were written.")
}

inspect_schema <- function(dataset, country, absolute_path, relative_path,
                           required_sheet) {
  expected <- required_columns[[dataset]]

  tryCatch({
    sheets <- excel_sheets(absolute_path)
    if (!required_sheet %in% sheets) {
      return(tibble(
        dataset = dataset,
        country = country,
        relative_path = relative_path,
        sheet = required_sheet,
        source_rows = NA_integer_,
        source_columns = NA_integer_,
        missing_required = paste(expected, collapse = "; "),
        additional_columns = NA_character_,
        duplicate_headers = NA_character_,
        required_order_matches = FALSE,
        schema_status = "failed_missing_sheet",
        schema_reason = paste0("Required sheet not found: ", required_sheet)
      ))
    }

    header_only <- read_excel(
      absolute_path,
      sheet = required_sheet,
      n_max = 0,
      .name_repair = "minimal"
    )
    headers <- names(header_only)
    duplicate_headers <- unique(headers[duplicated(headers)])
    missing_required <- setdiff(expected, headers)
    additional_columns <- setdiff(headers, expected)
    source_rows <- nrow(read_excel(
      absolute_path,
      sheet = required_sheet,
      col_types = "list",
      trim_ws = FALSE,
      .name_repair = "minimal"
    ))

    schema_ok <- length(missing_required) == 0L &&
      length(duplicate_headers) == 0L

    schema_reason_value <- if (length(missing_required) > 0L) {
      paste0(
        "Missing required columns: ", paste(missing_required, collapse = "; ")
      )
    } else if (length(duplicate_headers) > 0L) {
      paste0(
        "Duplicated headers: ", paste(duplicate_headers, collapse = "; ")
      )
    } else {
      NA_character_
    }

    tibble(
      dataset = dataset,
      country = country,
      relative_path = relative_path,
      sheet = required_sheet,
      source_rows = source_rows,
      source_columns = length(headers),
      missing_required = collapse_or_na(missing_required),
      additional_columns = collapse_or_na(additional_columns),
      duplicate_headers = collapse_or_na(duplicate_headers),
      required_order_matches = identical(
        headers[headers %in% expected], expected
      ),
      schema_status = ifelse(schema_ok, "schema_ok", "failed_schema"),
      schema_reason = schema_reason_value
    )
  }, error = function(e) {
    tibble(
      dataset = dataset,
      country = country,
      relative_path = relative_path,
      sheet = required_sheet,
      source_rows = NA_integer_,
      source_columns = NA_integer_,
      missing_required = NA_character_,
      additional_columns = NA_character_,
      duplicate_headers = NA_character_,
      required_order_matches = FALSE,
      schema_status = "failed_read",
      schema_reason = conditionMessage(e)
    )
  })
}

schema_audit <- pmap_dfr(
  candidate_inventory %>%
    select(dataset, country, absolute_path, relative_path, required_sheet),
  inspect_schema
)

if (any(schema_audit$schema_status != "schema_ok")) {
  print(schema_audit)
  stop(
    "At least one final workbook failed schema validation. ",
    "Missing required columns were not created and no outputs were written."
  )
}
```

Required source columns are read first as typed Excel cells. Storage types and semantic compatibility are audited before explicit conversion to a common type. Mixed numeric/text storage is permitted only for fields documented as flexible text or numeric-like values; incompatible non-missing values stop the run.

```{r load-and-stack}
load_validated_file <- function(dataset, country, absolute_path, relative_path,
                                required_sheet) {
  expected <- required_columns[[dataset]]

  tryCatch({
    raw_data <- read_excel(
      absolute_path,
      sheet = required_sheet,
      col_types = "list",
      trim_ws = FALSE,
      .name_repair = "minimal"
    ) %>%
      select(all_of(expected))

    field_audit <- map_dfr(expected, function(field) {
      target_type <- target_type_for(field)
      column <- raw_data[[field]]
      storage_types <- map_chr(column, cell_storage_type)
      compatible <- map_lgl(
        column,
        cell_is_compatible,
        field = field,
        target_type = target_type
      )
      text_values <- map_chr(column, cell_to_text)
      incompatible_examples <- unique(text_values[!compatible])

      tibble(
        dataset = dataset,
        country = country,
        relative_path = relative_path,
        field = field,
        target_type = target_type,
        observed_storage_types = {
          observed <- sort(unique(na.omit(storage_types)))
          if (length(observed) == 0L) {
            "all_missing"
          } else {
            paste(observed, collapse = "; ")
          }
        },
        n_nonmissing = sum(!is.na(storage_types)),
        n_incompatible = sum(!compatible),
        incompatible_examples = collapse_or_na(head(incompatible_examples, 5)),
        conversion_rule = case_when(
          target_type == "text" && field %in% flexible_text_fields ~
            "explicit character conversion without padding or recoding",
          target_type == "text" ~ "preserve character value exactly",
          target_type == "numeric" ~ "strict numeric parse",
          target_type == "integer" ~ "strict whole-number parse",
          target_type == "binary_0_1" ~
            "normalize 0/1 or TRUE/FALSE to integer 0/1",
          TRUE ~ NA_character_
        ),
        type_status = ifelse(
          sum(!compatible) == 0L, "compatible", "incompatible"
        )
      )
    })

    if (any(field_audit$n_incompatible > 0L)) {
      return(list(
        status = "failed_type",
        reason = "One or more required fields contain incompatible values",
        n_rows = nrow(raw_data),
        data = NULL,
        type_audit = field_audit
      ))
    }

    converted_columns <- map(expected, function(field) {
      convert_list_column(
        raw_data[[field]], field, target_type_for(field)
      )
    })
    names(converted_columns) <- expected
    converted_data <- as_tibble(converted_columns)

    observed_countries <- unique(converted_data$country[
      !is_missing_value(converted_data$country)
    ])
    country_ok <- length(observed_countries) == 1L &&
      observed_countries[[1]] == country

    if (!country_ok) {
      return(list(
        status = "failed_country",
        reason = paste0(
          "Workbook country values do not uniquely equal folder country ",
          country, ": ", paste(observed_countries, collapse = "; ")
        ),
        n_rows = nrow(converted_data),
        data = NULL,
        type_audit = field_audit
      ))
    }

    list(
      status = "loaded",
      reason = NA_character_,
      n_rows = nrow(converted_data),
      data = converted_data,
      type_audit = field_audit
    )
  }, error = function(e) {
    list(
      status = "failed_read",
      reason = conditionMessage(e),
      n_rows = NA_integer_,
      data = NULL,
      type_audit = tibble(
        dataset = character(), country = character(),
        relative_path = character(), field = character(),
        target_type = character(), observed_storage_types = character(),
        n_nonmissing = integer(), n_incompatible = integer(),
        incompatible_examples = character(), conversion_rule = character(),
        type_status = character()
      )
    )
  })
}

load_results <- pmap(
  candidate_inventory %>%
    select(dataset, country, absolute_path, relative_path, required_sheet),
  load_validated_file
)

load_status <- candidate_inventory %>%
  select(dataset, country, relative_path) %>%
  mutate(
    load_status = map_chr(load_results, "status"),
    load_reason = map_chr(load_results, function(result) {
      if (is.null(result$reason) || length(result$reason) == 0L) {
        NA_character_
      } else {
        result$reason
      }
    }),
    loaded_rows = map_int(load_results, function(result) {
      if (is.null(result$n_rows) || length(result$n_rows) == 0L) {
        NA_integer_
      } else {
        as.integer(result$n_rows)
      }
    })
  )

type_audit <- map_dfr(load_results, "type_audit")

type_file_summary <- type_audit %>%
  group_by(dataset, country, relative_path) %>%
  summarise(
    type_status = ifelse(
      any(n_incompatible > 0L), "incompatible", "compatible"
    ),
    type_variation_fields = collapse_or_na(
      field[str_detect(observed_storage_types, fixed(";"))]
    ),
    incompatible_fields = collapse_or_na(field[n_incompatible > 0L]),
    incompatible_examples = collapse_or_na(
      incompatible_examples[n_incompatible > 0L]
    ),
    .groups = "drop"
  )

if (any(load_status$load_status != "loaded")) {
  print(load_status)
  print(type_audit %>% filter(n_incompatible > 0L))
  stop(
    "At least one final workbook failed type or country validation. ",
    "No incompatible values were coerced and no outputs were written."
  )
}

SSA_CENSUS_STACKED <- load_results[
  candidate_inventory$dataset == "Census"
] %>%
  map("data") %>%
  bind_rows()

SSA_EARMARK_ALLOCATIONS_STACKED <- load_results[
  candidate_inventory$dataset == "Allocation"
] %>%
  map("data") %>%
  bind_rows()

stopifnot(
  identical(names(SSA_CENSUS_STACKED), census_columns),
  identical(names(SSA_EARMARK_ALLOCATIONS_STACKED), allocation_columns),
  nrow(SSA_CENSUS_STACKED) == sum(
    load_status$loaded_rows[load_status$dataset == "Census"]
  ),
  nrow(SSA_EARMARK_ALLOCATIONS_STACKED) == sum(
    load_status$loaded_rows[load_status$dataset == "Allocation"]
  )
)

loaded_census_countries <- sort(unique(SSA_CENSUS_STACKED$country))
loaded_allocation_countries <- sort(
  unique(SSA_EARMARK_ALLOCATIONS_STACKED$country)
)

if (!identical(loaded_census_countries, loaded_allocation_countries)) {
  stop(
    "Loaded Census and Allocation country sets differ. No outputs were written."
  )
}
if (!"COD" %in% loaded_census_countries) {
  stop("COD did not load successfully. No outputs were written.")
}
```

# 2. Create the analytical Allocation table

`SSA_EARMARK_ALLOCATIONS` retains only direct shares of gross instrument proceeds (`share_level == 1`). Missing or other share levels are excluded and counted in the audit.

```{r filter-allocation}
allocation_stacked_with_row_id <- SSA_EARMARK_ALLOCATIONS_STACKED %>%
  mutate(.allocation_row_id = row_number())

allocation_analytical_with_row_id <- allocation_stacked_with_row_id %>%
  filter(share_level == 1L)

SSA_EARMARK_ALLOCATIONS <- allocation_analytical_with_row_id %>%
  select(all_of(allocation_columns))

make_filter_audit <- function(before, after, dataset, condition, key_column,
                              filter_fields) {
  countries <- sort(unique(before$country[!is_missing_value(before$country)]))
  scopes <- c(countries, "Overall")

  map_dfr(scopes, function(scope_country) {
    before_scope <- if (scope_country == "Overall") {
      before
    } else {
      before %>% filter(country == scope_country)
    }
    after_scope <- if (scope_country == "Overall") {
      after
    } else {
      after %>% filter(country == scope_country)
    }
    missing_filter_row <- rep(FALSE, nrow(before_scope))
    for (field in filter_fields) {
      missing_filter_row <- missing_filter_row |
        is_missing_value(before_scope[[field]])
    }

    tibble(
      audit_type = "filter",
      dataset = dataset,
      country = scope_country,
      condition_or_status = condition,
      rows_before = nrow(before_scope),
      rows_after = nrow(after_scope),
      keys_before = n_distinct_country_key(before_scope, key_column),
      keys_after = n_distinct_country_key(after_scope, key_column),
      missing_filter_value_rows = sum(missing_filter_row)
    )
  })
}

allocation_filter_audit <- make_filter_audit(
  SSA_EARMARK_ALLOCATIONS_STACKED,
  SSA_EARMARK_ALLOCATIONS,
  dataset = "Allocation",
  condition = "share_level == 1",
  key_column = "pair_id",
  filter_fields = "share_level"
)

cat(
  "Allocation filter: ",
  nrow(SSA_EARMARK_ALLOCATIONS_STACKED), " rows and ",
  n_distinct_country_key(SSA_EARMARK_ALLOCATIONS_STACKED, "pair_id"),
  " country-pair keys before; ",
  nrow(SSA_EARMARK_ALLOCATIONS), " rows and ",
  n_distinct_country_key(SSA_EARMARK_ALLOCATIONS, "pair_id"),
  " keys after.\n",
  sep = ""
)
```

# 3. Expand and resolve Census references

The expansion starts from `SSA_EARMARK_ALLOCATIONS`, not the unfiltered stack. `census_ref` must be a JSON array of Census row IDs. Each scalar token is matched case-sensitively to Census `evidence_id` within the same country, then resolved to that Census row's `instrument_id`. No `instrument_id` fallback and no cross-country match are used.

```{r parse-and-resolve-census-ref}
parse_reference_cell <- function(country, allocation_evidence_id, pair_id,
                                 census_ref_raw, allocation_row_id) {
  base_row <- tibble(
    country = country,
    .allocation_row_id = allocation_row_id,
    allocation_evidence_id = allocation_evidence_id,
    pair_id = pair_id,
    census_ref_raw = census_ref_raw
  )

  if (is.na(census_ref_raw) || str_trim(census_ref_raw) == "") {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "null_missing_raw",
        parse_detail = "census_ref is missing or blank"
      ))
  }

  parsed <- tryCatch(
    jsonlite::fromJSON(census_ref_raw, simplifyVector = FALSE),
    error = identity
  )

  if (inherits(parsed, "error")) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "malformed_json",
        parse_detail = conditionMessage(parsed)
      ))
  }

  if (is.null(parsed)) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "null_json",
        parse_detail = "census_ref is JSON null"
      ))
  }

  if (!is.list(parsed) || !is.null(names(parsed))) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = if (!is.list(parsed)) cell_to_text(parsed) else NA_character_,
        ref_token_trimmed = if (!is.list(parsed)) {
          str_trim(cell_to_text(parsed))
        } else {
          NA_character_
        },
        parse_status = "malformed_non_array",
        parse_detail = "JSON value is not an array"
      ))
  }

  if (length(parsed) == 0L) {
    return(base_row %>%
      mutate(
        element_position = NA_integer_,
        ref_token = NA_character_,
        ref_token_trimmed = NA_character_,
        parse_status = "null_empty_array",
        parse_detail = "census_ref is an empty JSON array"
      ))
  }

  map_dfr(seq_along(parsed), function(position) {
    element <- parsed[[position]]
    if (is.null(element) || length(element) == 0L) {
      return(base_row %>%
        mutate(
          element_position = as.integer(position),
          ref_token = NA_character_,
          ref_token_trimmed = NA_character_,
          parse_status = "null_array_element",
          parse_detail = "JSON array element is null"
        ))
    }

    element_is_scalar <- !is.list(element) && length(element) == 1L
    token <- if (element_is_scalar) cell_to_text(element) else NA_character_
    token_trimmed <- if (!is.na(token)) str_trim(token) else NA_character_

    base_row %>%
      mutate(
        element_position = as.integer(position),
        ref_token = token,
        ref_token_trimmed = token_trimmed,
        parse_status = case_when(
          !element_is_scalar ~ "malformed_non_scalar",
          is.na(token_trimmed) || token_trimmed == "" ~
            "malformed_empty_token",
          TRUE ~ "parsed"
        ),
        parse_detail = case_when(
          !element_is_scalar ~ "JSON array element is not a scalar",
          is.na(token_trimmed) || token_trimmed == "" ~
            "JSON array element is missing or blank",
          TRUE ~ NA_character_
        )
      )
  })
}

reference_tokens <- pmap_dfr(
  allocation_analytical_with_row_id %>%
    select(
      country,
      allocation_evidence_id = evidence_id,
      pair_id,
      census_ref_raw = census_ref,
      allocation_row_id = .allocation_row_id
    ),
  parse_reference_cell
)

census_evidence_map <- SSA_CENSUS_STACKED %>%
  mutate(
    evidence_lookup = str_trim(evidence_id),
    instrument_lookup = str_trim(instrument_id)
  ) %>%
  filter(!is_missing_value(evidence_lookup)) %>%
  group_by(country, evidence_lookup) %>%
  summarise(
    same_country_match_rows = n(),
    target_instrument_count = n_distinct(
      instrument_lookup[!is_missing_value(instrument_lookup)]
    ),
    census_evidence_id_match = ifelse(
      n() == 1L, first(evidence_lookup), NA_character_
    ),
    census_instrument_id_match = {
      targets <- unique(
        instrument_lookup[!is_missing_value(instrument_lookup)]
      )
      if (length(targets) == 1L) targets[[1]] else NA_character_
    },
    .groups = "drop"
  )

global_evidence_locations <- SSA_CENSUS_STACKED %>%
  transmute(country, evidence_lookup = str_trim(evidence_id)) %>%
  filter(!is_missing_value(evidence_lookup)) %>%
  distinct(country, evidence_lookup) %>%
  group_by(evidence_lookup) %>%
  summarise(
    matching_countries = paste(sort(unique(country)), collapse = "; "),
    .groups = "drop"
  )

reference_resolution <- reference_tokens %>%
  left_join(
    census_evidence_map,
    by = c("country", "ref_token_trimmed" = "evidence_lookup")
  ) %>%
  left_join(
    global_evidence_locations,
    by = c("ref_token_trimmed" = "evidence_lookup")
  ) %>%
  mutate(
    same_country_match_rows = replace_na(same_country_match_rows, 0L),
    target_instrument_count = replace_na(target_instrument_count, 0L),
    token_country_prefix = str_match(
      ref_token_trimmed, "^([A-Z]{3})_"
    )[, 2],
    expected_evidence_id_form = str_detect(
      ref_token_trimmed,
      "^[A-Z]{3}_.+_CEN_[0-9]+$"
    ),
    exact_other_country_match = same_country_match_rows == 0L &
      !is.na(matching_countries),
    outcome = case_when(
      str_starts(parse_status, "null_") ~ "null",
      str_starts(parse_status, "malformed_") ~ "malformed",
      same_country_match_rows > 1L ~ "ambiguous",
      same_country_match_rows == 1L & target_instrument_count > 1L ~
        "ambiguous",
      same_country_match_rows == 1L & target_instrument_count == 0L ~
        "unmatched",
      same_country_match_rows == 1L & target_instrument_count == 1L ~
        "matched",
      !is.na(token_country_prefix) & token_country_prefix != country ~
        "cross_country",
      exact_other_country_match ~ "cross_country",
      expected_evidence_id_form ~ "unmatched",
      TRUE ~ "malformed"
    ),
    resolution_status = case_when(
      outcome == "matched" ~ "matched_same_country_evidence_id",
      outcome == "ambiguous" ~ "ambiguous_same_country_evidence_id",
      outcome == "cross_country" ~ "cross_country_reference",
      outcome == "unmatched" & same_country_match_rows == 1L ~
        "matched_evidence_missing_instrument_id",
      outcome == "unmatched" ~ "unmatched_evidence_id",
      outcome == "malformed" & parse_status == "parsed" ~
        "malformed_identifier",
      TRUE ~ parse_status
    ),
    census_evidence_id = ifelse(
      outcome == "matched", census_evidence_id_match, NA_character_
    ),
    census_instrument_id = ifelse(
      outcome == "matched", census_instrument_id_match, NA_character_
    ),
    successfully_resolved = outcome == "matched"
  )

all_reference_outcomes <- c(
  "matched", "null", "malformed", "unmatched", "cross_country", "ambiguous"
)

reference_resolution_summary <- reference_resolution %>%
  count(country, outcome, name = "n") %>%
  complete(
    country = sort(unique(reference_resolution$country)),
    outcome = all_reference_outcomes,
    fill = list(n = 0L)
  ) %>%
  pivot_wider(names_from = outcome, values_from = n, values_fill = 0L) %>%
  bind_rows(
    reference_resolution %>%
      count(outcome, name = "n") %>%
      complete(outcome = all_reference_outcomes, fill = list(n = 0L)) %>%
      pivot_wider(names_from = outcome, values_from = n, values_fill = 0L) %>%
      mutate(country = "Overall", .before = 1)
  ) %>%
  mutate(
    n_reference_values = matched + null + malformed + unmatched +
      cross_country + ambiguous,
    pct_matched = ifelse(n_reference_values == 0L, NA_real_, matched /
      n_reference_values)
  ) %>%
  arrange(country != "Overall", country)
```

Only successfully resolved references can enter the expanded table. When repeated source rows produce the same `country + pair_id + census_instrument_id` link, the notebook audits every differing retained column, keeps the first source-row occurrence deterministically, and removes later duplicate links as requested.

```{r create-expanded-allocation}
matched_reference_rows <- reference_resolution %>%
  filter(successfully_resolved) %>%
  select(
    .allocation_row_id, element_position,
    census_evidence_id, census_instrument_id
  )

expanded_link_candidates_all <- allocation_analytical_with_row_id %>%
  inner_join(matched_reference_rows, by = ".allocation_row_id") %>%
  arrange(.allocation_row_id, element_position)

expanded_link_candidates <- expanded_link_candidates_all %>%
  filter(
    !is_missing_value(country),
    !is_missing_value(pair_id),
    !is_missing_value(census_instrument_id)
  )

matched_links_missing_key <- expanded_link_candidates_all %>%
  filter(
    is_missing_value(country) |
      is_missing_value(pair_id) |
      is_missing_value(census_instrument_id)
  )

duplicate_link_keys <- expanded_link_candidates %>%
  count(country, pair_id, census_instrument_id, name = "n_source_rows") %>%
  filter(n_source_rows > 1L)

different_nonmissing_values <- function(values) {
  if (is.character(values)) {
    values <- values[!is.na(values) & values != ""]
  } else {
    values <- values[!is.na(values)]
  }
  n_distinct(values) > 1L
}

duplicate_link_detail <- expanded_link_candidates %>%
  semi_join(
    duplicate_link_keys,
    by = c("country", "pair_id", "census_instrument_id")
  ) %>%
  group_by(country, pair_id, census_instrument_id) %>%
  group_modify(function(group_data, group_keys) {
    fields_to_check <- intersect(allocation_columns, names(group_data))
    differing_fields <- fields_to_check[vapply(
      group_data[fields_to_check],
      different_nonmissing_values,
      logical(1)
    )]
    conflicting_fields <- setdiff(differing_fields, "evidence_id")

    tibble(
      n_source_rows = nrow(group_data),
      differing_retained_columns = collapse_or_na(differing_fields),
      conflicting_retained_columns = collapse_or_na(conflicting_fields),
      has_conflicting_retained_values = length(conflicting_fields) > 0L
    )
  }) %>%
  ungroup()

SSA_EARMARK_ALLOCATIONS_EXPANDED <- expanded_link_candidates %>%
  distinct(
    country, pair_id, census_instrument_id,
    .keep_all = TRUE
  ) %>%
  select(
    all_of(allocation_columns),
    census_evidence_id,
    census_instrument_id
  )

stopifnot(
  identical(
    names(SSA_EARMARK_ALLOCATIONS_EXPANDED),
    c(allocation_columns, "census_evidence_id", "census_instrument_id")
  ),
  !any(duplicated(
    SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
      select(country, pair_id, census_instrument_id)
  ))
)
```

# 4. Add `is_earmarked` and create the analytical Census table

`is_earmarked` is based only on successfully resolved, level-1 Allocation links. It is added to the unfiltered Census stack before applying the requested current, non-penalty Census filter.

```{r create-census-tables}
earmarked_census_keys <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  transmute(
    country,
    instrument_id = census_instrument_id,
    is_earmarked = 1L
  ) %>%
  filter(
    !is_missing_value(country),
    !is_missing_value(instrument_id)
  ) %>%
  distinct(country, instrument_id, .keep_all = TRUE)

SSA_CENSUS_STACKED <- SSA_CENSUS_STACKED %>%
  left_join(
    earmarked_census_keys,
    by = c("country", "instrument_id")
  ) %>%
  mutate(is_earmarked = replace_na(is_earmarked, 0L)) %>%
  select(all_of(census_columns), is_earmarked)

SSA_CENSUS <- SSA_CENSUS_STACKED %>%
  filter(in_force == 1L, is_penalty == 0L) %>%
  select(all_of(census_columns), is_earmarked)

census_filter_audit <- make_filter_audit(
  SSA_CENSUS_STACKED,
  SSA_CENSUS,
  dataset = "Census",
  condition = "in_force == 1 and is_penalty == 0",
  key_column = "instrument_id",
  filter_fields = c("in_force", "is_penalty")
)

cat(
  "Census filter: ", nrow(SSA_CENSUS_STACKED), " rows and ",
  n_distinct_country_key(SSA_CENSUS_STACKED, "instrument_id"),
  " country-instrument keys before; ", nrow(SSA_CENSUS), " rows and ",
  n_distinct_country_key(SSA_CENSUS, "instrument_id"),
  " keys after.\n",
  sep = ""
)
```

# 5. Essential sanity checks

Exactly three combined sanity tables are produced: Coverage, Key integrity, and Link/filter audit.

```{r prepare-coverage-sanity}
file_load_audit <- schema_audit %>%
  left_join(
    load_status,
    by = c("dataset", "country", "relative_path")
  ) %>%
  left_join(
    type_file_summary,
    by = c("dataset", "country", "relative_path")
  ) %>%
  mutate(
    final_status = case_when(
      schema_status != "schema_ok" ~ schema_status,
      load_status != "loaded" ~ load_status,
      TRUE ~ "loaded"
    )
  )

country_presence <- full_join(
  SSA_CENSUS_STACKED %>% distinct(country) %>% mutate(in_census = TRUE),
  SSA_EARMARK_ALLOCATIONS_STACKED %>%
    distinct(country) %>%
    mutate(in_allocation = TRUE),
  by = "country"
) %>%
  mutate(
    in_census = replace_na(in_census, FALSE),
    in_allocation = replace_na(in_allocation, FALSE),
    missing_dataset = case_when(
      in_census & in_allocation ~ NA_character_,
      in_census ~ "Allocation",
      in_allocation ~ "Census",
      TRUE ~ "Census; Allocation"
    ),
    status = ifelse(is.na(missing_dataset), "both", "missing_dataset")
  ) %>%
  arrange(country)

coverage_file_summary <- file_load_audit %>%
  group_by(dataset) %>%
  summarise(
    n_files_found = n(),
    n_files_loaded = sum(final_status == "loaded"),
    n_files_failed = sum(str_starts(final_status, "failed")),
    n_files_duplicated = 0L,
    n_countries = n_distinct(country[final_status == "loaded"]),
    n_rows = sum(loaded_rows[final_status == "loaded"], na.rm = TRUE),
    record_type = "file_summary",
    country = "Overall",
    status = ifelse(all(final_status == "loaded"), "complete", "incomplete"),
    missing_dataset = NA_character_,
    relative_path = NA_character_,
    required_sheet = NA_character_,
    missing_required = collapse_or_na(missing_required),
    additional_columns = collapse_or_na(additional_columns),
    duplicate_headers = collapse_or_na(duplicate_headers),
    required_order_matches = all(required_order_matches),
    type_status = ifelse(
      all(type_status == "compatible"), "compatible", "incompatible"
    ),
    type_variation_fields = collapse_or_na(type_variation_fields),
    detail = "Canonical final country workbooks",
    .groups = "drop"
  ) %>%
  select(
    record_type, dataset, country, status, n_files_found, n_files_loaded,
    n_files_failed, n_files_duplicated, n_countries, n_rows,
    missing_dataset, relative_path, required_sheet, missing_required,
    additional_columns, duplicate_headers, required_order_matches,
    type_status, type_variation_fields, detail
  )

coverage_country_files <- file_load_audit %>%
  transmute(
    record_type = "country_file",
    dataset,
    country,
    status = final_status,
    n_files_found = 1L,
    n_files_loaded = as.integer(final_status == "loaded"),
    n_files_failed = as.integer(str_starts(final_status, "failed")),
    n_files_duplicated = 0L,
    n_countries = 1L,
    n_rows = loaded_rows,
    missing_dataset = NA_character_,
    relative_path,
    required_sheet = sheet,
    missing_required,
    additional_columns,
    duplicate_headers,
    required_order_matches,
    type_status,
    type_variation_fields,
    detail = coalesce(schema_reason, load_reason, "Loaded successfully")
  )

coverage_country_presence <- country_presence %>%
  transmute(
    record_type = "country_presence",
    dataset = "Both",
    country,
    status,
    n_files_found = NA_integer_,
    n_files_loaded = NA_integer_,
    n_files_failed = NA_integer_,
    n_files_duplicated = NA_integer_,
    n_countries = 1L,
    n_rows = NA_integer_,
    missing_dataset,
    relative_path = NA_character_,
    required_sheet = NA_character_,
    missing_required = NA_character_,
    additional_columns = NA_character_,
    duplicate_headers = NA_character_,
    required_order_matches = NA,
    type_status = NA_character_,
    type_variation_fields = NA_character_,
    detail = paste0(
      "in_census=", in_census, "; in_allocation=", in_allocation
    )
  )

coverage_missing_summary <- tibble(
  record_type = "missing_dataset_summary",
  dataset = "Both",
  country = "Overall",
  status = ifelse(
    any(!is.na(country_presence$missing_dataset)), "incomplete", "complete"
  ),
  n_files_found = NA_integer_,
  n_files_loaded = NA_integer_,
  n_files_failed = NA_integer_,
  n_files_duplicated = NA_integer_,
  n_countries = sum(!is.na(country_presence$missing_dataset)),
  n_rows = NA_integer_,
  missing_dataset = collapse_or_na(country_presence$missing_dataset),
  relative_path = NA_character_,
  required_sheet = NA_character_,
  missing_required = NA_character_,
  additional_columns = NA_character_,
  duplicate_headers = NA_character_,
  required_order_matches = NA,
  type_status = NA_character_,
  type_variation_fields = NA_character_,
  detail = collapse_or_na(
    country_presence$country[!is.na(country_presence$missing_dataset)]
  )
)

coverage_sanity <- bind_rows(
  coverage_file_summary,
  coverage_country_files,
  coverage_country_presence,
  coverage_missing_summary
) %>%
  arrange(factor(record_type, levels = c(
    "file_summary", "country_file", "country_presence",
    "missing_dataset_summary"
  )), dataset, country)
```

```{r prepare-key-integrity}
make_key_integrity <- function(data, dataset_scope, dataset, key_name,
                               key_columns, interpretation) {
  countries <- sort(unique(data$country[!is_missing_value(data$country)]))
  scopes <- c(countries, "Overall")

  map_dfr(scopes, function(scope_country) {
    scoped <- if (scope_country == "Overall") {
      data
    } else {
      data %>% filter(country == scope_country)
    }
    missing_country <- is_missing_value(scoped$country)
    missing_id <- rep(FALSE, nrow(scoped))
    for (field in key_columns) {
      missing_id <- missing_id | is_missing_value(scoped[[field]])
    }
    valid <- !missing_country & !missing_id
    valid_keys <- scoped[valid, c("country", key_columns), drop = FALSE] %>%
      count(across(everything()), name = "key_rows")
    duplicate_keys <- valid_keys %>% filter(key_rows > 1L)

    tibble(
      dataset_scope = dataset_scope,
      dataset = dataset,
      country = scope_country,
      key_name = key_name,
      n_rows = nrow(scoped),
      n_missing_country_rows = sum(missing_country),
      n_missing_id_rows = sum(missing_id),
      n_missing_key_rows = sum(missing_country | missing_id),
      n_distinct_nonmissing_keys = nrow(valid_keys),
      n_duplicate_key_groups = nrow(duplicate_keys),
      n_duplicate_key_rows = sum(duplicate_keys$key_rows),
      n_excess_duplicate_rows = sum(duplicate_keys$key_rows - 1L),
      interpretation = interpretation
    )
  })
}

key_integrity <- bind_rows(
  make_key_integrity(
    SSA_CENSUS_STACKED, "SSA_CENSUS_STACKED", "Census",
    "country + evidence_id", "evidence_id",
    "Census evidence-row identifier"
  ),
  make_key_integrity(
    SSA_CENSUS_STACKED, "SSA_CENSUS_STACKED", "Census",
    "country + instrument_id", "instrument_id",
    "Census instrument key"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS_STACKED,
    "SSA_EARMARK_ALLOCATIONS_STACKED", "Allocation",
    "country + evidence_id", "evidence_id",
    "Allocation evidence-row identifier"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS_STACKED,
    "SSA_EARMARK_ALLOCATIONS_STACKED", "Allocation",
    "country + instrument_id", "instrument_id",
    "Repeated values are expected when one instrument has several channels"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS_STACKED,
    "SSA_EARMARK_ALLOCATIONS_STACKED", "Allocation",
    "country + pair_id", "pair_id",
    "Allocation channel key"
  ),
  make_key_integrity(
    SSA_CENSUS, "SSA_CENSUS", "Census",
    "country + evidence_id", "evidence_id",
    "Analytical Census evidence-row identifier"
  ),
  make_key_integrity(
    SSA_CENSUS, "SSA_CENSUS", "Census",
    "country + instrument_id", "instrument_id",
    "Analytical Census instrument key"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS,
    "SSA_EARMARK_ALLOCATIONS", "Allocation",
    "country + evidence_id", "evidence_id",
    "Analytical Allocation evidence-row identifier"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS,
    "SSA_EARMARK_ALLOCATIONS", "Allocation",
    "country + instrument_id", "instrument_id",
    "Repeated values are expected when one instrument has several channels"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS,
    "SSA_EARMARK_ALLOCATIONS", "Allocation",
    "country + pair_id", "pair_id",
    "Analytical Allocation channel key"
  ),
  make_key_integrity(
    SSA_EARMARK_ALLOCATIONS_EXPANDED,
    "SSA_EARMARK_ALLOCATIONS_EXPANDED", "Allocation",
    "country + pair_id + census_instrument_id",
    c("pair_id", "census_instrument_id"),
    "Required expanded-link grain"
  )
) %>%
  arrange(dataset_scope, key_name, country != "Overall", country)
```

```{r prepare-link-filter-audit}
reference_summary_audit <- reference_resolution_summary %>%
  transmute(
    audit_type = "reference_summary",
    dataset = "Allocation",
    country,
    condition_or_status = "census_ref resolution",
    n_reference_values,
    matched,
    null,
    malformed,
    unmatched,
    cross_country,
    ambiguous,
    pct_matched,
    rows_before = NA_integer_,
    rows_after = NA_integer_,
    keys_before = NA_integer_,
    keys_after = NA_integer_,
    missing_filter_value_rows = NA_integer_,
    matched_link_candidate_rows = NA_integer_,
    expanded_link_rows = NA_integer_,
    duplicate_link_rows_removed = NA_integer_,
    duplicate_link_groups = NA_integer_,
    conflicting_duplicate_link_groups = NA_integer_,
    reference_examples = NA_character_,
    pair_examples = NA_character_,
    conflicting_fields = NA_character_,
    detail = "Occurrence-level reference counts; matched only within country"
  )

reference_issue_audit <- reference_resolution %>%
  filter(outcome != "matched") %>%
  group_by(country, outcome) %>%
  summarise(
    n_reference_values = n(),
    reference_examples = collapse_or_na(
      unique(coalesce(ref_token_trimmed, census_ref_raw))
    ),
    pair_examples = collapse_or_na(unique(pair_id)),
    detail = collapse_or_na(unique(resolution_status)),
    .groups = "drop"
  ) %>%
  transmute(
    audit_type = "reference_issue_detail",
    dataset = "Allocation",
    country,
    condition_or_status = outcome,
    n_reference_values,
    matched = NA_integer_,
    null = NA_integer_,
    malformed = NA_integer_,
    unmatched = NA_integer_,
    cross_country = NA_integer_,
    ambiguous = NA_integer_,
    pct_matched = NA_real_,
    rows_before = NA_integer_,
    rows_after = NA_integer_,
    keys_before = NA_integer_,
    keys_after = NA_integer_,
    missing_filter_value_rows = NA_integer_,
    matched_link_candidate_rows = NA_integer_,
    expanded_link_rows = NA_integer_,
    duplicate_link_rows_removed = NA_integer_,
    duplicate_link_groups = NA_integer_,
    conflicting_duplicate_link_groups = NA_integer_,
    reference_examples,
    pair_examples,
    conflicting_fields = NA_character_,
    detail
  )

filter_audit_rows <- bind_rows(
  allocation_filter_audit,
  census_filter_audit
) %>%
  transmute(
    audit_type,
    dataset,
    country,
    condition_or_status,
    n_reference_values = NA_integer_,
    matched = NA_integer_,
    null = NA_integer_,
    malformed = NA_integer_,
    unmatched = NA_integer_,
    cross_country = NA_integer_,
    ambiguous = NA_integer_,
    pct_matched = NA_real_,
    rows_before,
    rows_after,
    keys_before,
    keys_after,
    missing_filter_value_rows,
    matched_link_candidate_rows = NA_integer_,
    expanded_link_rows = NA_integer_,
    duplicate_link_rows_removed = NA_integer_,
    duplicate_link_groups = NA_integer_,
    conflicting_duplicate_link_groups = NA_integer_,
    reference_examples = NA_character_,
    pair_examples = NA_character_,
    conflicting_fields = NA_character_,
    detail = "Distinct keys exclude missing country or ID values"
  )

link_audit_countries <- sort(unique(SSA_EARMARK_ALLOCATIONS$country))
link_dedup_audit <- map_dfr(
  c(link_audit_countries, "Overall"),
  function(scope_country) {
    candidate_scope <- if (scope_country == "Overall") {
      expanded_link_candidates
    } else {
      expanded_link_candidates %>% filter(country == scope_country)
    }
    expanded_scope <- if (scope_country == "Overall") {
      SSA_EARMARK_ALLOCATIONS_EXPANDED
    } else {
      SSA_EARMARK_ALLOCATIONS_EXPANDED %>% filter(country == scope_country)
    }
    duplicate_scope <- if (scope_country == "Overall") {
      duplicate_link_detail
    } else {
      duplicate_link_detail %>% filter(country == scope_country)
    }
    missing_key_scope <- if (scope_country == "Overall") {
      matched_links_missing_key
    } else {
      matched_links_missing_key %>% filter(country == scope_country)
    }

    tibble(
      audit_type = "expanded_link_deduplication",
      dataset = "Allocation",
      country = scope_country,
      condition_or_status = "one country + pair_id + census_instrument_id link",
      n_reference_values = NA_integer_,
      matched = NA_integer_,
      null = NA_integer_,
      malformed = NA_integer_,
      unmatched = NA_integer_,
      cross_country = NA_integer_,
      ambiguous = NA_integer_,
      pct_matched = NA_real_,
      rows_before = NA_integer_,
      rows_after = NA_integer_,
      keys_before = NA_integer_,
      keys_after = NA_integer_,
      missing_filter_value_rows = NA_integer_,
      matched_link_candidate_rows = nrow(candidate_scope),
      expanded_link_rows = nrow(expanded_scope),
      duplicate_link_rows_removed = nrow(candidate_scope) - nrow(expanded_scope),
      duplicate_link_groups = nrow(duplicate_scope),
      conflicting_duplicate_link_groups = sum(
        duplicate_scope$has_conflicting_retained_values
      ),
      reference_examples = NA_character_,
      pair_examples = collapse_or_na(unique(duplicate_scope$pair_id)),
      conflicting_fields = if (nrow(duplicate_scope) == 0L) {
        NA_character_
      } else {
        paste(
          paste0(
            duplicate_scope$pair_id, " -> ",
            duplicate_scope$conflicting_retained_columns
          ),
          collapse = " | "
        )
      },
      detail = paste0(
        "First source-row occurrence retained after audit; matched candidates ",
        "missing a complete link key=", nrow(missing_key_scope)
      )
    )
  }
)

link_filter_audit <- bind_rows(
  reference_summary_audit,
  reference_issue_audit,
  filter_audit_rows,
  link_dedup_audit
) %>%
  arrange(
    factor(audit_type, levels = c(
      "filter", "reference_summary", "reference_issue_detail",
      "expanded_link_deduplication"
    )),
    dataset,
    country != "Overall",
    country,
    condition_or_status
  )
```

## Coverage

```{r display-coverage-sanity}
display_table(coverage_sanity)
```

## Key integrity

```{r display-key-integrity}
display_table(key_integrity)
```

## Link and filter audit

```{r display-link-filter-audit}
display_table(link_filter_audit)
```

# 6. Country and legal-corpus coverage

Coverage uses the unfiltered stacked datasets. Distinct blank `document_id` values are not counted. Source year values are preserved, deduplicated, sorted, and joined with `"; "`.

```{r corpus-coverage}
sort_year_values <- function(values) {
  values <- unique(values[!is_missing_value(values)])
  if (length(values) == 0L) return(NA_character_)
  paste(sort(values), collapse = "; ")
}

make_corpus_coverage <- function(data, dataset_name) {
  data %>%
    group_by(country) %>%
    summarise(
      n_documents = n_distinct(
        document_id[!is_missing_value(document_id)]
      ),
      document_years = sort_year_values(document_year),
      corpus_type = case_when(
        n_documents == 1L ~ "single_document",
        n_documents > 1L ~ "multi_document",
        TRUE ~ NA_character_
      ),
      .groups = "drop"
    ) %>%
    mutate(dataset = dataset_name, .before = 1) %>%
    select(dataset, country, n_documents, document_years, corpus_type)
}

country_corpus_coverage <- bind_rows(
  make_corpus_coverage(SSA_CENSUS_STACKED, "Census"),
  make_corpus_coverage(
    SSA_EARMARK_ALLOCATIONS_STACKED, "Allocation"
  )
) %>%
  arrange(dataset, country)

display_table(country_corpus_coverage)
```

# 7. Export and verify the five datasets

```{r export-csv-files}
write_csv(SSA_CENSUS_STACKED, output_paths$census_stacked, na = "")
write_csv(
  SSA_EARMARK_ALLOCATIONS_STACKED,
  output_paths$allocation_stacked,
  na = ""
)
write_csv(SSA_CENSUS, output_paths$census, na = "")
write_csv(SSA_EARMARK_ALLOCATIONS, output_paths$allocation, na = "")
write_csv(
  SSA_EARMARK_ALLOCATIONS_EXPANDED,
  output_paths$allocation_expanded,
  na = ""
)
write_csv(
  country_corpus_coverage,
  output_paths$corpus_coverage,
  na = ""
)

verify_csv <- function(path, expected_rows, expected_columns) {
  saved <- read_csv(
    path,
    col_types = cols(.default = col_character()),
    show_col_types = FALSE,
    progress = FALSE
  )
  stopifnot(
    nrow(saved) == expected_rows,
    identical(names(saved), expected_columns)
  )
  invisible(TRUE)
}

verify_csv(
  output_paths$census_stacked,
  nrow(SSA_CENSUS_STACKED),
  c(census_columns, "is_earmarked")
)
verify_csv(
  output_paths$allocation_stacked,
  nrow(SSA_EARMARK_ALLOCATIONS_STACKED),
  allocation_columns
)
verify_csv(
  output_paths$census,
  nrow(SSA_CENSUS),
  c(census_columns, "is_earmarked")
)
verify_csv(
  output_paths$allocation,
  nrow(SSA_EARMARK_ALLOCATIONS),
  allocation_columns
)
verify_csv(
  output_paths$allocation_expanded,
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED),
  c(allocation_columns, "census_evidence_id", "census_instrument_id")
)
verify_csv(
  output_paths$corpus_coverage,
  nrow(country_corpus_coverage),
  c("dataset", "country", "n_documents", "document_years", "corpus_type")
)
```

```{r stage-sanity-workbook, include=FALSE}
serialize_table <- function(data, name) {
  rows <- map(seq_len(nrow(data)), function(row_number) {
    set_names(
      map(data, function(column) {
        value <- column[[row_number]]
        if (length(value) == 0L || is.na(value)) NA else unname(value)
      }),
      names(data)
    )
  })

  list(name = name, columns = names(data), rows = rows)
}

workbook_tables <- list(
  coverage = coverage_sanity,
  key_integrity = key_integrity,
  link_filter_audit = link_filter_audit
)

workbook_manifest <- list(
  title = "SSA stack sanity checks",
  generated_at = format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"),
  output_path = normalizePath(
    output_paths$sanity_workbook, winslash = "/", mustWork = FALSE
  ),
  sheets = imap(workbook_tables, serialize_table)
)

write_json(
  workbook_manifest,
  manifest_path,
  auto_unbox = TRUE,
  pretty = TRUE,
  na = "null",
  digits = NA
)
```

# 8. Run summary

```{r run-summary, results='asis'}
overall_reference_summary <- reference_resolution_summary %>%
  filter(country == "Overall")

cat(
  "- Included countries (", length(loaded_census_countries), "): ",
  paste(loaded_census_countries, collapse = ", "), ".\n",
  "- Stacked rows: Census ", nrow(SSA_CENSUS_STACKED),
  "; Allocation ", nrow(SSA_EARMARK_ALLOCATIONS_STACKED), ".\n",
  "- Analytical rows: Census ", nrow(SSA_CENSUS),
  "; Allocation ", nrow(SSA_EARMARK_ALLOCATIONS),
  "; expanded links ", nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED), ".\n",
  "- Reference outcomes: matched ", overall_reference_summary$matched,
  "; null ", overall_reference_summary$null,
  "; malformed ", overall_reference_summary$malformed,
  "; unmatched ", overall_reference_summary$unmatched,
  "; cross-country ", overall_reference_summary$cross_country,
  "; ambiguous ", overall_reference_summary$ambiguous, ".\n",
  "- Duplicate expanded-link rows removed: ",
  nrow(expanded_link_candidates) - nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED),
  "; duplicate-link groups with conflicting retained values: ",
  sum(duplicate_link_detail$has_conflicting_retained_values), ".\n",
  sep = ""
)
```

The three sanity tables were staged for `analysis/tables/01_stack_sanity_checks.xlsx`. Workbook authoring and visual verification use the project's spreadsheet tooling after the notebook render completes.
