---
title: "SSA revenue instruments: country-level summary indicators"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
params:
  analysis_variant: "full"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook computes country-level count indicators from the four prepared derived datasets specified for this task. The database is a **de jure** legal inventory; the indicators do not measure collections, transfers, disbursements, or expenditure execution. No source workbook is opened, no input file is modified, and no plot is created.

The analytical units remain distinct throughout: Census instrument (`country + instrument_id`), Allocation source group (`country + instrument_id`), Allocation pair (`country + pair_id`), evidence row (`country + evidence_id`), and atomic Allocation link (`country + pair_id + census_instrument_id`).

```{r setup, include=FALSE}
input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

analysis_variant <- params$analysis_variant
allowed_analysis_variants <- c("full", "partial_exclusive")
if (!analysis_variant %in% allowed_analysis_variants) {
  stop(
    "Unknown analysis_variant: ", analysis_variant,
    ". Expected one of: ", paste(allowed_analysis_variants, collapse = "; ")
  )
}
is_partial_exclusive <- identical(analysis_variant, "partial_exclusive")
partial_exclusive_statuses <- c("partial_earmark", "exclusive_earmark")

analysis_table_dir <- file.path(project_root, "analysis", "tables")
if (is_partial_exclusive) {
  analysis_table_dir <- file.path(analysis_table_dir, "partial_exclusive")
}

input_paths <- c(
  SSA_CENSUS_STACKED = file.path(
    project_root, "data", "derived", "SSA_CENSUS_STACKED.csv"
  ),
  SSA_CENSUS = file.path(
    project_root, "data", "derived", "SSA_CENSUS.csv"
  ),
  SSA_EARMARK_ALLOCATIONS = file.path(
    project_root, "data", "derived", "SSA_EARMARK_ALLOCATIONS.csv"
  ),
  SSA_EARMARK_ALLOCATIONS_EXPANDED = file.path(
    project_root,
    "data",
    "derived",
    "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  )
)

output_path <- file.path(
  analysis_table_dir, "02_country_summary.csv"
)
dir.create(dirname(output_path), recursive = TRUE, showWarnings = FALSE)

output_columns <- c(
  "country",
  "D_all_census_instruments",
  "D_census_instruments",
  "E_all_linked_census_instruments",
  "E_linked_census_instruments",
  "A_allocation_source_groups",
  "P_allocation_pairs",
  "L_atomic_allocation_links",
  "R_evidence_rows",
  "allocation_incidence_E_D",
  "allocation_link_density_L_E",
  "pairs_per_source_group_P_A",
  "unlinked_census_instruments_D_minus_E",
  "allocation_scope",
  "census_scope"
)

is_missing_id <- function(values) {
  if (is.character(values)) {
    is.na(values) | str_trim(values) == ""
  } else {
    is.na(values)
  }
}

safe_ratio <- function(numerator, denominator) {
  if_else(denominator == 0L, NA_real_, numerator / denominator)
}

display_table <- function(data, digits = 6) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    knitr::kable(data, digits = digits, row.names = FALSE)
  }
}
```

# 1. Load and validate the four derived datasets

Only the four declared CSV files are read. Identifier fields are imported as text; the already-derived binary and level flags used for scope checks are imported as integers.

```{r load-derived-data}
missing_input_files <- names(input_paths)[!file.exists(input_paths)]
if (length(missing_input_files) > 0L) {
  stop(
    "Missing required derived input files: ",
    paste(missing_input_files, collapse = "; ")
  )
}

SSA_CENSUS_STACKED <- read_csv(
  input_paths[["SSA_CENSUS_STACKED"]],
  col_types = cols(
    .default = col_character(),
    in_force = col_integer(),
    is_penalty = col_integer(),
    is_earmarked = col_integer()
  ),
  show_col_types = FALSE,
  progress = FALSE
)

SSA_CENSUS <- read_csv(
  input_paths[["SSA_CENSUS"]],
  col_types = cols(
    .default = col_character(),
    in_force = col_integer(),
    is_penalty = col_integer(),
    is_earmarked = col_integer()
  ),
  show_col_types = FALSE,
  progress = FALSE
)

SSA_EARMARK_ALLOCATIONS <- read_csv(
  input_paths[["SSA_EARMARK_ALLOCATIONS"]],
  col_types = cols(
    .default = col_character(),
    share_level = col_integer()
  ),
  show_col_types = FALSE,
  progress = FALSE
)

SSA_EARMARK_ALLOCATIONS_EXPANDED <- read_csv(
  input_paths[["SSA_EARMARK_ALLOCATIONS_EXPANDED"]],
  col_types = cols(
    .default = col_character(),
    share_level = col_integer()
  ),
  show_col_types = FALSE,
  progress = FALSE
)

required_fields <- list(
  SSA_CENSUS_STACKED = c("country", "instrument_id"),
  SSA_CENSUS = c(
    "country", "instrument_id", "in_force", "is_penalty", "is_earmarked"
  ),
  SSA_EARMARK_ALLOCATIONS = c(
    "country", "instrument_id", "pair_id", "evidence_id", "share_level",
    if (is_partial_exclusive) "earmark_status"
  ),
  SSA_EARMARK_ALLOCATIONS_EXPANDED = c(
    "country", "pair_id", "census_instrument_id", "census_evidence_id",
    "share_level", if (is_partial_exclusive) "earmark_status"
  )
)

input_objects <- list(
  SSA_CENSUS_STACKED = SSA_CENSUS_STACKED,
  SSA_CENSUS = SSA_CENSUS,
  SSA_EARMARK_ALLOCATIONS = SSA_EARMARK_ALLOCATIONS,
  SSA_EARMARK_ALLOCATIONS_EXPANDED = SSA_EARMARK_ALLOCATIONS_EXPANDED
)

input_inventory <- imap_dfr(input_objects, function(data, dataset) {
  missing_fields <- setdiff(required_fields[[dataset]], names(data))
  tibble(
    dataset = dataset,
    file = basename(input_paths[[dataset]]),
    n_rows = nrow(data),
    n_columns = ncol(data),
    n_countries = n_distinct(
      data$country[!is_missing_id(data$country)]
    ),
    missing_required_fields = if (length(missing_fields) == 0L) {
      NA_character_
    } else {
      paste(missing_fields, collapse = "; ")
    },
    schema_status = if_else(
      length(missing_fields) == 0L, "valid", "invalid"
    )
  )
})

if (any(input_inventory$schema_status == "invalid")) {
  print(input_inventory)
  stop("At least one derived input is missing required fields.")
}

SSA_EARMARK_ALLOCATIONS_UNFILTERED <- SSA_EARMARK_ALLOCATIONS
SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED <-
  SSA_EARMARK_ALLOCATIONS_EXPANDED

if (is_partial_exclusive) {
  allocation_status_problems <- bind_rows(
    SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
      filter(
        is_missing_id(earmark_status) |
          !earmark_status %in% c(
            "assigned_unrestricted", "cost_recovery_only",
            "partial_earmark", "exclusive_earmark"
          )
      ) %>%
      transmute(dataset = "SSA_EARMARK_ALLOCATIONS", country, instrument_id,
                earmark_status),
    SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
      filter(
        is_missing_id(earmark_status) |
          !earmark_status %in% c(
            "assigned_unrestricted", "cost_recovery_only",
            "partial_earmark", "exclusive_earmark"
          )
      ) %>%
      transmute(
        dataset = "SSA_EARMARK_ALLOCATIONS_EXPANDED",
        country, instrument_id, earmark_status
      )
  )

  allocation_status_map <- SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
    filter(!is_missing_id(country), !is_missing_id(instrument_id)) %>%
    distinct(country, instrument_id, earmark_status)
  expanded_status_map <- SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
    filter(!is_missing_id(country), !is_missing_id(instrument_id)) %>%
    distinct(country, instrument_id, earmark_status)

  status_key_conflicts <- bind_rows(
    allocation_status_map %>%
      count(country, instrument_id, name = "n_statuses") %>%
      filter(n_statuses != 1L) %>%
      mutate(dataset = "SSA_EARMARK_ALLOCATIONS", .before = 1),
    expanded_status_map %>%
      count(country, instrument_id, name = "n_statuses") %>%
      filter(n_statuses != 1L) %>%
      mutate(dataset = "SSA_EARMARK_ALLOCATIONS_EXPANDED", .before = 1)
  )
  expanded_status_mismatches <- expanded_status_map %>%
    anti_join(
      allocation_status_map,
      by = c("country", "instrument_id", "earmark_status")
    )

  if (nrow(allocation_status_problems) > 0L ||
      nrow(status_key_conflicts) > 0L ||
      nrow(expanded_status_mismatches) > 0L) {
    print(allocation_status_problems)
    print(status_key_conflicts)
    print(expanded_status_mismatches)
    stop("Allocation earmark_status validation failed before filtering.")
  }

  SSA_EARMARK_ALLOCATIONS <- SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
    filter(earmark_status %in% partial_exclusive_statuses)
  SSA_EARMARK_ALLOCATIONS_EXPANDED <-
    SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
    filter(earmark_status %in% partial_exclusive_statuses)
}

analysis_filter_audit <- tribble(
  ~dataset, ~measure, ~before, ~after,
  "SSA_EARMARK_ALLOCATIONS", "rows",
  nrow(SSA_EARMARK_ALLOCATIONS_UNFILTERED),
  nrow(SSA_EARMARK_ALLOCATIONS),
  "SSA_EARMARK_ALLOCATIONS", "distinct country + pair_id",
  nrow(distinct(
    SSA_EARMARK_ALLOCATIONS_UNFILTERED, country, pair_id
  )),
  nrow(distinct(SSA_EARMARK_ALLOCATIONS, country, pair_id)),
  "SSA_EARMARK_ALLOCATIONS_EXPANDED", "expanded links",
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED),
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED)
)

filtered_status_distribution <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country, instrument_id, earmark_status) %>%
  count(earmark_status, name = "n_allocation_source_instruments") %>%
  arrange(earmark_status)

filtered_countries <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country) %>%
  arrange(country)

if (is_partial_exclusive &&
    (!all(SSA_EARMARK_ALLOCATIONS$earmark_status %in%
          partial_exclusive_statuses) ||
     !all(SSA_EARMARK_ALLOCATIONS_EXPANDED$earmark_status %in%
          partial_exclusive_statuses))) {
  stop("A non-partial/non-exclusive status remains after filtering.")
}

# Subsequent inventory and coverage checks describe the active analysis objects.
input_objects <- list(
  SSA_CENSUS_STACKED = SSA_CENSUS_STACKED,
  SSA_CENSUS = SSA_CENSUS,
  SSA_EARMARK_ALLOCATIONS = SSA_EARMARK_ALLOCATIONS,
  SSA_EARMARK_ALLOCATIONS_EXPANDED = SSA_EARMARK_ALLOCATIONS_EXPANDED
)
```

## Allocation-status filter audit

```{r display-analysis-filter-audit}
display_table(analysis_filter_audit)
display_table(filtered_status_distribution)
cat(
  "Countries remaining in the allocation sample (", nrow(filtered_countries),
  "): ", paste(filtered_countries$country, collapse = ", "), ".\n",
  "Only partial_earmark and exclusive_earmark present: ",
  if_else(
    all(SSA_EARMARK_ALLOCATIONS$earmark_status %in%
        partial_exclusive_statuses) &&
      all(SSA_EARMARK_ALLOCATIONS_EXPANDED$earmark_status %in%
          partial_exclusive_statuses),
    "yes", "no"
  ), ".\n",
  sep = ""
)
```

The prepared analytical scopes are verified rather than re-filtered. A scope violation is a hard stop because silently filtering it here would impose an unreported additional transformation.

```{r verify-input-scopes}
scope_audit <- tribble(
  ~dataset, ~declared_scope, ~n_rows, ~n_scope_violations,
  "SSA_CENSUS",
  "in_force == 1 and is_penalty == 0",
  nrow(SSA_CENSUS),
  sum(
    is.na(SSA_CENSUS$in_force) |
      is.na(SSA_CENSUS$is_penalty) |
      SSA_CENSUS$in_force != 1L |
      SSA_CENSUS$is_penalty != 0L
  ),
  "SSA_EARMARK_ALLOCATIONS",
  "share_level == 1",
  nrow(SSA_EARMARK_ALLOCATIONS),
  sum(
    is.na(SSA_EARMARK_ALLOCATIONS$share_level) |
      SSA_EARMARK_ALLOCATIONS$share_level != 1L
  ),
  "SSA_EARMARK_ALLOCATIONS_EXPANDED",
  "share_level == 1",
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED),
  sum(
    is.na(SSA_EARMARK_ALLOCATIONS_EXPANDED$share_level) |
      SSA_EARMARK_ALLOCATIONS_EXPANDED$share_level != 1L
  )
) %>%
  mutate(scope_status = if_else(
    n_scope_violations == 0L, "valid", "violation"
  ))

if (any(scope_audit$n_scope_violations > 0L)) {
  print(scope_audit)
  stop(
    "A derived input violates its declared analytical scope; no additional ",
    "filter was applied and no output was written."
  )
}

country_coverage <- reduce(
  imap(input_objects, function(data, dataset) {
    data %>%
      filter(!is_missing_id(country)) %>%
      distinct(country) %>%
      mutate(!!dataset := TRUE)
  }),
  full_join,
  by = "country"
) %>%
  mutate(across(-country, ~ replace_na(.x, FALSE))) %>%
  arrange(country)
```

## Input inventory

```{r display-input-inventory}
display_table(input_inventory)
```

## Scope audit

```{r display-scope-audit}
display_table(scope_audit)
```

## Country coverage

```{r display-country-coverage}
display_table(country_coverage)
```

# 2. Missing identifiers

Missingness is reported for every identifier used in the requested counts or joins. Source values are not filled or corrected.

```{r missing-id-audit}
make_missing_id_report <- function(data, dataset, id_fields) {
  country_rows <- data %>%
    mutate(country_scope = if_else(
      is_missing_id(country), "Missing_country", country
    )) %>%
    select(country_scope, all_of(id_fields)) %>%
    pivot_longer(
      cols = all_of(id_fields),
      names_to = "id_field",
      values_to = "id_value",
      values_transform = list(id_value = as.character)
    ) %>%
    group_by(country = country_scope, id_field) %>%
    summarise(
      n_rows = n(),
      n_missing_id_rows = sum(is_missing_id(id_value)),
      pct_missing = n_missing_id_rows / n_rows,
      .groups = "drop"
    )

  overall_rows <- data %>%
    select(all_of(id_fields)) %>%
    pivot_longer(
      cols = everything(),
      names_to = "id_field",
      values_to = "id_value",
      values_transform = list(id_value = as.character)
    ) %>%
    group_by(id_field) %>%
    summarise(
      country = "Overall",
      n_rows = n(),
      n_missing_id_rows = sum(is_missing_id(id_value)),
      pct_missing = n_missing_id_rows / n_rows,
      .groups = "drop"
    ) %>%
    select(country, everything())

  bind_rows(overall_rows, country_rows) %>%
    mutate(dataset = dataset, .before = 1) %>%
    arrange(id_field, country != "Overall", country)
}

missing_id_report <- bind_rows(
  make_missing_id_report(
    SSA_CENSUS_STACKED,
    "SSA_CENSUS_STACKED",
    c("country", "instrument_id")
  ),
  make_missing_id_report(
    SSA_CENSUS,
    "SSA_CENSUS",
    c("country", "instrument_id")
  ),
  make_missing_id_report(
    SSA_EARMARK_ALLOCATIONS,
    "SSA_EARMARK_ALLOCATIONS",
    c("country", "instrument_id", "pair_id", "evidence_id")
  ),
  make_missing_id_report(
    SSA_EARMARK_ALLOCATIONS_EXPANDED,
    "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    c("country", "pair_id", "census_instrument_id", "census_evidence_id")
  )
)
```

```{r display-missing-id-audit}
display_table(missing_id_report)
```

# 3. Country-level counts

No instrument-type restriction is applied. Allocation inputs are used at their prepared level-1 scope, and no allocation level is mixed with another.

```{r compute-count-components}
country_universe <- bind_rows(
  SSA_CENSUS_STACKED %>% select(country),
  SSA_CENSUS %>% select(country),
  SSA_EARMARK_ALLOCATIONS %>% select(country),
  SSA_EARMARK_ALLOCATIONS_EXPANDED %>% select(country)
) %>%
  filter(!is_missing_id(country)) %>%
  distinct(country) %>%
  arrange(country)

D_all_counts <- SSA_CENSUS_STACKED %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(instrument_id)
  ) %>%
  distinct(country, instrument_id) %>%
  count(country, name = "D_all_census_instruments")

D_counts <- SSA_CENSUS %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(instrument_id)
  ) %>%
  distinct(country, instrument_id) %>%
  count(country, name = "D_census_instruments")

E_all_counts <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(census_instrument_id)
  ) %>%
  distinct(country, census_instrument_id) %>%
  count(country, name = "E_all_linked_census_instruments")

E_counts <- if (is_partial_exclusive) {
  analytical_census_keys_for_e <- SSA_CENSUS %>%
    filter(
      !is_missing_id(country),
      !is_missing_id(instrument_id)
    ) %>%
    distinct(country, census_instrument_id = instrument_id)

  SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
    filter(
      !is_missing_id(country),
      !is_missing_id(census_instrument_id)
    ) %>%
    distinct(country, census_instrument_id) %>%
    semi_join(
      analytical_census_keys_for_e,
      by = c("country", "census_instrument_id")
    ) %>%
    count(country, name = "E_linked_census_instruments")
} else {
  SSA_CENSUS %>%
    filter(
      is_earmarked == 1L,
      !is_missing_id(country),
      !is_missing_id(instrument_id)
    ) %>%
    distinct(country, instrument_id) %>%
    count(country, name = "E_linked_census_instruments")
}

A_counts <- SSA_EARMARK_ALLOCATIONS %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(instrument_id)
  ) %>%
  distinct(country, instrument_id) %>%
  count(country, name = "A_allocation_source_groups")

P_counts <- SSA_EARMARK_ALLOCATIONS %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(pair_id)
  ) %>%
  distinct(country, pair_id) %>%
  count(country, name = "P_allocation_pairs")

R_counts <- SSA_EARMARK_ALLOCATIONS %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(evidence_id)
  ) %>%
  distinct(country, evidence_id) %>%
  count(country, name = "R_evidence_rows")
```

`L_atomic_allocation_links` is the only requested count that applies a join-time restriction. The expanded link table is restricted to `country + census_instrument_id` values that appear as `country + instrument_id` in the analytical Census. The link grain is made explicit with `distinct(country, pair_id, census_instrument_id)`.

The prior stacking audit removed six repeated link rows across four Allocation pairs (`BEN_PAIR_0001`, `CIV_PAIR_0090`, `CIV_PAIR_0091`, and `GIN_PAIR_0039`) and retained the first source occurrence. This notebook treats the resulting expanded dataset as authoritative and does not reconstruct those rows from `census_ref`. Re-applying the distinct atomic-link key prevents duplication, while the retained allocation attributes do not enter any indicator requested here.

```{r compute-atomic-links}
analytical_census_keys <- SSA_CENSUS %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(instrument_id)
  ) %>%
  distinct(country, census_instrument_id = instrument_id)

all_atomic_links <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  filter(
    !is_missing_id(country),
    !is_missing_id(pair_id),
    !is_missing_id(census_instrument_id)
  ) %>%
  distinct(country, pair_id, census_instrument_id)

analytical_atomic_links <- all_atomic_links %>%
  semi_join(
    analytical_census_keys,
    by = c("country", "census_instrument_id")
  ) %>%
  distinct(country, pair_id, census_instrument_id)

L_counts <- analytical_atomic_links %>%
  count(country, name = "L_atomic_allocation_links")

make_link_filter_audit <- function(country_value) {
  expanded_scope <- if (country_value == "Overall") {
    SSA_EARMARK_ALLOCATIONS_EXPANDED
  } else {
    SSA_EARMARK_ALLOCATIONS_EXPANDED %>% filter(country == country_value)
  }
  all_link_scope <- if (country_value == "Overall") {
    all_atomic_links
  } else {
    all_atomic_links %>% filter(country == country_value)
  }
  analytical_link_scope <- if (country_value == "Overall") {
    analytical_atomic_links
  } else {
    analytical_atomic_links %>% filter(country == country_value)
  }

  tibble(
    country = country_value,
    condition = paste0(
      "Keep links whose country + census_instrument_id appears in SSA_CENSUS"
    ),
    expanded_rows_before = nrow(expanded_scope),
    atomic_links_before = nrow(all_link_scope),
    atomic_links_after = nrow(analytical_link_scope),
    atomic_links_excluded = nrow(all_link_scope) - nrow(analytical_link_scope)
  )
}

link_filter_audit <- map_dfr(
  c(country_universe$country, "Overall"),
  make_link_filter_audit
) %>%
  arrange(country != "Overall", country)
```

```{r display-link-filter-audit}
display_table(link_filter_audit)
```

# 4. Ratios and final country table

```{r construct-country-summary}
country_summary <- country_universe %>%
  left_join(D_all_counts, by = "country") %>%
  left_join(D_counts, by = "country") %>%
  left_join(E_all_counts, by = "country") %>%
  left_join(E_counts, by = "country") %>%
  left_join(A_counts, by = "country") %>%
  left_join(P_counts, by = "country") %>%
  left_join(L_counts, by = "country") %>%
  left_join(R_counts, by = "country") %>%
  mutate(
    across(
      c(
        D_all_census_instruments,
        D_census_instruments,
        E_all_linked_census_instruments,
        E_linked_census_instruments,
        A_allocation_source_groups,
        P_allocation_pairs,
        L_atomic_allocation_links,
        R_evidence_rows
      ),
      ~ replace_na(as.integer(.x), 0L)
    ),
    allocation_incidence_E_D = safe_ratio(
      E_linked_census_instruments,
      D_census_instruments
    ),
    allocation_link_density_L_E = safe_ratio(
      L_atomic_allocation_links,
      E_linked_census_instruments
    ),
    pairs_per_source_group_P_A = safe_ratio(
      P_allocation_pairs,
      A_allocation_source_groups
    ),
    unlinked_census_instruments_D_minus_E =
      D_census_instruments - E_linked_census_instruments,
    allocation_scope = "share_level_1",
    census_scope = "in_force_1_non_penalty"
  ) %>%
  select(all_of(output_columns)) %>%
  arrange(country)

stopifnot(
  identical(names(country_summary), output_columns),
  !any(duplicated(country_summary$country)),
  !any(is.infinite(country_summary$allocation_incidence_E_D)),
  !any(is.infinite(country_summary$allocation_link_density_L_E)),
  !any(is.infinite(country_summary$pairs_per_source_group_P_A))
)
```

## Final country-level indicators

```{r display-country-summary}
display_table(country_summary)
```

# 5. Validation inequalities

The four requested inequalities are evaluated country by country. Violations are reported but the computed values are never corrected.

```{r validation-checks}
validation_checks <- bind_rows(
  country_summary %>%
    transmute(
      country,
      check = "E_all <= D_all",
      left_value = E_all_linked_census_instruments,
      operator = "<=",
      right_value = D_all_census_instruments,
      passed = left_value <= right_value
    ),
  country_summary %>%
    transmute(
      country,
      check = "E <= D",
      left_value = E_linked_census_instruments,
      operator = "<=",
      right_value = D_census_instruments,
      passed = left_value <= right_value
    ),
  country_summary %>%
    transmute(
      country,
      check = "L >= E",
      left_value = L_atomic_allocation_links,
      operator = ">=",
      right_value = E_linked_census_instruments,
      passed = left_value >= right_value
    ),
  country_summary %>%
    transmute(
      country,
      check = "P >= A",
      left_value = P_allocation_pairs,
      operator = ">=",
      right_value = A_allocation_source_groups,
      passed = left_value >= right_value
    )
) %>%
  mutate(validation_status = if_else(passed, "passed", "violation")) %>%
  arrange(country, check)

validation_violations <- validation_checks %>%
  filter(!passed)

validation_summary <- validation_checks %>%
  group_by(check) %>%
  summarise(
    n_countries_checked = n(),
    n_passed = sum(passed),
    n_violations = sum(!passed),
    violating_countries = {
      values <- country[!passed]
      if (length(values) == 0L) NA_character_ else paste(values, collapse = "; ")
    },
    .groups = "drop"
  )
```

## Validation summary

```{r display-validation-summary}
display_table(validation_summary)
```

## Validation violations

```{r display-validation-violations}
display_table(validation_violations)
```

# 6. Export and verify

```{r export-country-summary}
write_csv(country_summary, output_path, na = "")

saved_country_summary <- read_csv(
  output_path,
  col_types = cols(
    country = col_character(),
    D_all_census_instruments = col_integer(),
    D_census_instruments = col_integer(),
    E_all_linked_census_instruments = col_integer(),
    E_linked_census_instruments = col_integer(),
    A_allocation_source_groups = col_integer(),
    P_allocation_pairs = col_integer(),
    L_atomic_allocation_links = col_integer(),
    R_evidence_rows = col_integer(),
    allocation_incidence_E_D = col_double(),
    allocation_link_density_L_E = col_double(),
    pairs_per_source_group_P_A = col_double(),
    unlinked_census_instruments_D_minus_E = col_integer(),
    allocation_scope = col_character(),
    census_scope = col_character()
  ),
  show_col_types = FALSE,
  progress = FALSE
)

stopifnot(
  nrow(saved_country_summary) == nrow(country_summary),
  identical(names(saved_country_summary), output_columns),
  isTRUE(all.equal(
    saved_country_summary,
    country_summary,
    check.attributes = FALSE,
    tolerance = 1e-12
  ))
)
```

# 7. Run summary

```{r run-summary, results='asis'}
overall_missing_ids <- missing_id_report %>%
  filter(country == "Overall") %>%
  summarise(n = sum(n_missing_id_rows)) %>%
  pull(n)

cat(
  "- Analysis variant: ", analysis_variant, ".\n",
  "- Countries included (", nrow(country_summary), "): ",
  paste(country_summary$country, collapse = ", "), ".\n",
  "- Missing identifier rows across the requested key fields: ",
  overall_missing_ids, ".\n",
  "- Validation violations: ", nrow(validation_violations), ".\n",
  "- Atomic expanded links before the analytical Census restriction: ",
  nrow(all_atomic_links), "; after restriction: ",
  nrow(analytical_atomic_links), ".\n",
  "- Output: ",
  if_else(
    is_partial_exclusive,
    "analysis/tables/partial_exclusive/02_country_summary.csv",
    "analysis/tables/02_country_summary.csv"
  ), ".\n",
  sep = ""
)
```
