---
title: "SSA revenue instruments: basic allocation summary"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
params:
  analysis_variant: "full"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook summarizes the canonical level-1 statutory allocation layer. The database is a **de jure** legal inventory; it does not measure actual collections, transfers, disbursements, or expenditure execution. The analytical scope is selected once in setup and audited below; no additional downstream restriction is imposed, and no plot is created.

All allocation-composition statistics use one distinct `country + pair_id` from `SSA_EARMARK_ALLOCATIONS.csv`. The expanded link table is loaded only for input context and is never counted as allocations.

```{r setup, include=FALSE}
input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

analysis_variant <- params$analysis_variant
allowed_analysis_variants <- c("full", "partial_exclusive")
if (!analysis_variant %in% allowed_analysis_variants) {
  stop(
    "Unknown analysis_variant: ", analysis_variant,
    ". Expected one of: ", paste(allowed_analysis_variants, collapse = "; ")
  )
}
is_partial_exclusive <- identical(analysis_variant, "partial_exclusive")
partial_exclusive_statuses <- c("partial_earmark", "exclusive_earmark")

analysis_table_dir <- file.path(project_root, "analysis", "tables")
if (is_partial_exclusive) {
  analysis_table_dir <- file.path(analysis_table_dir, "partial_exclusive")
}

input_paths <- c(
  SSA_CENSUS = file.path(
    project_root, "data", "derived", "SSA_CENSUS.csv"
  ),
  SSA_EARMARK_ALLOCATIONS = file.path(
    project_root, "data", "derived", "SSA_EARMARK_ALLOCATIONS.csv"
  ),
  SSA_EARMARK_ALLOCATIONS_EXPANDED = file.path(
    project_root,
    "data",
    "derived",
    "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  ),
  COFOG_GFSM2014_mapping = file.path(
    project_root, "data", "COFOG_GFSM2014_mapping.csv"
  )
)

output_paths <- c(
  assignment_type = file.path(
    analysis_table_dir, "03_assignment_type_summary.csv"
  ),
  allocation_nature = file.path(
    analysis_table_dir, "03_allocation_nature_summary.csv"
  ),
  purpose_restriction = file.path(
    analysis_table_dir,
    "03_purpose_restriction_summary.csv"
  ),
  beneficiary_type = file.path(
    analysis_table_dir, "03_beneficiary_type_summary.csv"
  ),
  destination_function_long = file.path(
    analysis_table_dir,
    "03_destination_function_long.csv"
  ),
  destination_function_summary = file.path(
    analysis_table_dir,
    "03_destination_function_summary.csv"
  )
)

dir.create(
  analysis_table_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

is_missing_value <- function(values) {
  is.na(values) | str_trim(as.character(values)) == ""
}

normalize_category <- function(values) {
  normalized <- str_trim(as.character(values))
  if_else(is.na(normalized) | normalized == "", "Missing", normalized)
}

display_table <- function(data, digits = 6) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    knitr::kable(data, digits = digits, row.names = FALSE)
  }
}

make_composition_summary <- function(
  pair_categories,
  country_universe,
  country_pair_totals
) {
  observed_categories <- pair_categories %>%
    distinct(category) %>%
    arrange(category)

  country_category_counts <- pair_categories %>%
    count(country, category, name = "n_allocations")

  country_category_grid <- crossing(
    country = country_universe$country,
    category = observed_categories$category
  ) %>%
    left_join(
      country_category_counts,
      by = c("country", "category")
    ) %>%
    mutate(n_allocations = replace_na(n_allocations, 0L)) %>%
    left_join(country_pair_totals, by = "country") %>%
    mutate(within_country_share = if_else(
      total_allocations > 0L,
      n_allocations / total_allocations,
      NA_real_
    ))

  summary_table <- country_category_grid %>%
    group_by(category) %>%
    summarise(
      pooled_n_allocations = sum(n_allocations),
      n_countries_using = sum(n_allocations > 0L),
      median_allocations_among_users = median(
        n_allocations[n_allocations > 0L]
      ),
      mean_within_country_share = mean(within_country_share),
      .groups = "drop"
    ) %>%
    arrange(desc(pooled_n_allocations), category)

  list(
    summary = summary_table,
    country_category_grid = country_category_grid
  )
}
```

# 1. Load and validate the four declared inputs

The COFOG/GFSM mapping is semicolon-delimited. Inspection identifies `cofog_code` as its code column and `cofog_name` as its function-label column.

```{r load-inputs}
missing_input_files <- names(input_paths)[!file.exists(input_paths)]
if (length(missing_input_files) > 0L) {
  stop(
    "Missing required input files: ",
    paste(missing_input_files, collapse = "; ")
  )
}

SSA_CENSUS <- read_csv(
  input_paths[["SSA_CENSUS"]],
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

SSA_EARMARK_ALLOCATIONS <- read_csv(
  input_paths[["SSA_EARMARK_ALLOCATIONS"]],
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

SSA_EARMARK_ALLOCATIONS_EXPANDED <- read_csv(
  input_paths[["SSA_EARMARK_ALLOCATIONS_EXPANDED"]],
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

COFOG_GFSM2014_mapping <- read_delim(
  input_paths[["COFOG_GFSM2014_mapping"]],
  delim = ";",
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE,
  trim_ws = FALSE
)

required_fields <- list(
  SSA_CENSUS = c("country", "instrument_id"),
  SSA_EARMARK_ALLOCATIONS = c(
    "country",
    "pair_id",
    "row_type",
    "share_level",
    "assignment_type",
    "allocation_nature",
    "is_purpose_restricted",
    "beneficiary_type",
    "destination_function",
    "destination_function_detail",
    if (is_partial_exclusive) "earmark_status"
  ),
  SSA_EARMARK_ALLOCATIONS_EXPANDED = c(
    "country", "instrument_id", "pair_id", "census_instrument_id",
    if (is_partial_exclusive) "earmark_status"
  ),
  COFOG_GFSM2014_mapping = c("cofog_code", "cofog_name")
)

input_objects <- list(
  SSA_CENSUS = SSA_CENSUS,
  SSA_EARMARK_ALLOCATIONS = SSA_EARMARK_ALLOCATIONS,
  SSA_EARMARK_ALLOCATIONS_EXPANDED = SSA_EARMARK_ALLOCATIONS_EXPANDED,
  COFOG_GFSM2014_mapping = COFOG_GFSM2014_mapping
)

input_inventory <- imap_dfr(input_objects, function(data, dataset) {
  missing_fields <- setdiff(required_fields[[dataset]], names(data))
  has_country <- "country" %in% names(data)

  tibble(
    dataset = dataset,
    file = basename(input_paths[[dataset]]),
    n_rows = nrow(data),
    n_columns = ncol(data),
    n_countries = if (has_country) {
      n_distinct(data$country[!is_missing_value(data$country)])
    } else {
      NA_integer_
    },
    missing_required_fields = if (length(missing_fields) == 0L) {
      NA_character_
    } else {
      paste(missing_fields, collapse = "; ")
    },
    schema_status = if_else(
      length(missing_fields) == 0L, "valid", "invalid"
    )
  )
})

if (any(input_inventory$schema_status == "invalid")) {
  print(input_inventory)
  stop("At least one input is missing required fields.")
}

mapping_column_audit <- tibble(
  role = c("code", "label"),
  identified_column = c("cofog_code", "cofog_name"),
  present = c(
    "cofog_code" %in% names(COFOG_GFSM2014_mapping),
    "cofog_name" %in% names(COFOG_GFSM2014_mapping)
  )
)

SSA_EARMARK_ALLOCATIONS_UNFILTERED <- SSA_EARMARK_ALLOCATIONS
SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED <-
  SSA_EARMARK_ALLOCATIONS_EXPANDED

if (is_partial_exclusive) {
  allowed_earmark_statuses <- c(
    "assigned_unrestricted", "cost_recovery_only",
    "partial_earmark", "exclusive_earmark"
  )
  allocation_status_problems <- bind_rows(
    SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
      filter(
        is_missing_value(earmark_status) |
          !earmark_status %in% allowed_earmark_statuses
      ) %>%
      transmute(dataset = "SSA_EARMARK_ALLOCATIONS", country, instrument_id,
                earmark_status),
    SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
      filter(
        is_missing_value(earmark_status) |
          !earmark_status %in% allowed_earmark_statuses
      ) %>%
      transmute(
        dataset = "SSA_EARMARK_ALLOCATIONS_EXPANDED",
        country, instrument_id, earmark_status
      )
  )
  allocation_status_map <- SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
    distinct(country, instrument_id, earmark_status)
  expanded_status_map <- SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
    distinct(country, instrument_id, earmark_status)
  status_key_conflicts <- bind_rows(
    allocation_status_map %>%
      count(country, instrument_id, name = "n_statuses") %>%
      filter(n_statuses != 1L) %>%
      mutate(dataset = "SSA_EARMARK_ALLOCATIONS", .before = 1),
    expanded_status_map %>%
      count(country, instrument_id, name = "n_statuses") %>%
      filter(n_statuses != 1L) %>%
      mutate(dataset = "SSA_EARMARK_ALLOCATIONS_EXPANDED", .before = 1)
  )
  expanded_status_mismatches <- expanded_status_map %>%
    anti_join(
      allocation_status_map,
      by = c("country", "instrument_id", "earmark_status")
    )

  if (nrow(allocation_status_problems) > 0L ||
      nrow(status_key_conflicts) > 0L ||
      nrow(expanded_status_mismatches) > 0L) {
    print(allocation_status_problems)
    print(status_key_conflicts)
    print(expanded_status_mismatches)
    stop("Allocation earmark_status validation failed before filtering.")
  }

  SSA_EARMARK_ALLOCATIONS <- SSA_EARMARK_ALLOCATIONS_UNFILTERED %>%
    filter(earmark_status %in% partial_exclusive_statuses)
  SSA_EARMARK_ALLOCATIONS_EXPANDED <-
    SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED %>%
    filter(earmark_status %in% partial_exclusive_statuses)
}

analysis_filter_audit <- tribble(
  ~dataset, ~measure, ~before, ~after,
  "SSA_EARMARK_ALLOCATIONS", "rows",
  nrow(SSA_EARMARK_ALLOCATIONS_UNFILTERED),
  nrow(SSA_EARMARK_ALLOCATIONS),
  "SSA_EARMARK_ALLOCATIONS", "distinct country + pair_id",
  nrow(distinct(
    SSA_EARMARK_ALLOCATIONS_UNFILTERED, country, pair_id
  )),
  nrow(distinct(SSA_EARMARK_ALLOCATIONS, country, pair_id)),
  "SSA_EARMARK_ALLOCATIONS_EXPANDED", "expanded links",
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED_UNFILTERED),
  nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED)
)

filtered_status_distribution <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country, instrument_id, earmark_status) %>%
  count(earmark_status, name = "n_allocation_source_instruments") %>%
  arrange(earmark_status)
filtered_countries <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country) %>%
  arrange(country)

if (is_partial_exclusive &&
    (!all(SSA_EARMARK_ALLOCATIONS$earmark_status %in%
          partial_exclusive_statuses) ||
     !all(SSA_EARMARK_ALLOCATIONS_EXPANDED$earmark_status %in%
          partial_exclusive_statuses))) {
  stop("A non-partial/non-exclusive status remains after filtering.")
}
```

## Input inventory

```{r display-input-inventory}
display_table(input_inventory)
```

## Mapping columns

```{r display-mapping-columns}
display_table(mapping_column_audit)
```

## Allocation-status filter audit

```{r display-analysis-filter-audit}
display_table(analysis_filter_audit)
display_table(filtered_status_distribution)
cat(
  "Countries remaining in the allocation sample (", nrow(filtered_countries),
  "): ", paste(filtered_countries$country, collapse = ", "), ".\n",
  "Only partial_earmark and exclusive_earmark present: ",
  if_else(
    all(SSA_EARMARK_ALLOCATIONS$earmark_status %in%
        partial_exclusive_statuses) &&
      all(SSA_EARMARK_ALLOCATIONS_EXPANDED$earmark_status %in%
          partial_exclusive_statuses),
    "yes", "no"
  ), ".\n",
  sep = ""
)
```

# 2. Allocation-key and scope checks

No row is filtered to obtain the declared scope. The allocation table must already contain only canonical level-1 allocation rows.

```{r validate-allocation-keys-and-scope}
allocation_scope_audit <- tibble(
  check = c(
    "missing_or_blank_country",
    "missing_or_blank_pair_id",
    "share_level_not_1_or_missing",
    "row_type_not_allocation_or_missing"
  ),
  n_problem_rows = c(
    sum(is_missing_value(SSA_EARMARK_ALLOCATIONS$country)),
    sum(is_missing_value(SSA_EARMARK_ALLOCATIONS$pair_id)),
    sum(
      is_missing_value(SSA_EARMARK_ALLOCATIONS$share_level) |
        SSA_EARMARK_ALLOCATIONS$share_level != "1"
    ),
    sum(
      is_missing_value(SSA_EARMARK_ALLOCATIONS$row_type) |
        SSA_EARMARK_ALLOCATIONS$row_type != "allocation"
    )
  )
) %>%
  mutate(status = if_else(n_problem_rows == 0L, "passed", "failed"))

if (any(allocation_scope_audit$n_problem_rows > 0L)) {
  print(allocation_scope_audit)
  stop(
    "Allocation key or scope validation failed. No summary output was written."
  )
}

country_universe <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country) %>%
  arrange(country)

allocation_pairs <- SSA_EARMARK_ALLOCATIONS %>%
  distinct(country, pair_id) %>%
  arrange(country, pair_id)

country_pair_totals <- allocation_pairs %>%
  count(country, name = "total_allocations") %>%
  arrange(country)

duplicate_pair_audit <- SSA_EARMARK_ALLOCATIONS %>%
  count(country, pair_id, name = "n_source_rows") %>%
  filter(n_source_rows > 1L) %>%
  mutate(excess_rows = n_source_rows - 1L) %>%
  arrange(country, pair_id)

allocation_grain_audit <- tibble(
  measure = c(
    "allocation_table_rows",
    "distinct_country_pair_allocations",
    "excess_rows_above_pair_grain",
    "countries_in_allocation_universe"
  ),
  value = c(
    nrow(SSA_EARMARK_ALLOCATIONS),
    nrow(allocation_pairs),
    nrow(SSA_EARMARK_ALLOCATIONS) - nrow(allocation_pairs),
    nrow(country_universe)
  )
)
```

```{r display-allocation-scope-audit}
display_table(allocation_scope_audit)
```

```{r display-allocation-grain-audit}
display_table(allocation_grain_audit)
```

```{r display-country-allocation-counts}
display_table(country_pair_totals)
```

```{r display-duplicate-pair-audit}
display_table(duplicate_pair_audit)
```

## Single-valued category consistency

`assignment_type`, `allocation_nature`, `is_purpose_restricted`, and `beneficiary_type` are single-valued at the allocation-pair grain. Missing values are treated as the explicit category `Missing` before consistency is tested, so a missing/non-missing disagreement is also a conflict.

Destination function is not included in this stop gate because the task explicitly permits one allocation to finance several functions and defines a separate long pair-function grain.

```{r validate-single-valued-categories}
single_valued_fields <- c(
  "assignment_type",
  "allocation_nature",
  "is_purpose_restricted",
  "beneficiary_type"
)

pair_category_consistency <- SSA_EARMARK_ALLOCATIONS %>%
  select(country, pair_id, evidence_id, all_of(single_valued_fields)) %>%
  mutate(across(all_of(single_valued_fields), normalize_category)) %>%
  pivot_longer(
    cols = all_of(single_valued_fields),
    names_to = "variable",
    values_to = "category"
  ) %>%
  group_by(country, pair_id, variable) %>%
  summarise(
    n_source_rows = n(),
    n_distinct_categories = n_distinct(category),
    observed_categories = paste(sort(unique(category)), collapse = " | "),
    evidence_ids = paste(evidence_id, collapse = " | "),
    .groups = "drop"
  ) %>%
  arrange(country, pair_id, variable)

single_value_conflicts <- pair_category_consistency %>%
  filter(n_distinct_categories > 1L)

if (nrow(single_value_conflicts) > 0L) {
  print(single_value_conflicts)
  stop(
    "At least one allocation pair has conflicting values for a single-valued ",
    "category. No summary output was written."
  )
}
```

```{r display-single-value-conflicts}
display_table(single_value_conflicts)
```

# 3. Pair-level category preparation and audits

Each of the four single-valued category datasets now contains exactly one row per `country + pair_id`.

```{r prepare-pair-categories}
assignment_type_pairs <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    country,
    pair_id,
    category = normalize_category(assignment_type)
  ) %>%
  distinct(country, pair_id, category)

allocation_nature_pairs <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    country,
    pair_id,
    category = normalize_category(allocation_nature)
  ) %>%
  distinct(country, pair_id, category)

purpose_restriction_pairs <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    country,
    pair_id,
    category = normalize_category(is_purpose_restricted)
  ) %>%
  distinct(country, pair_id, category)

beneficiary_type_pairs <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    country,
    pair_id,
    category = normalize_category(beneficiary_type)
  ) %>%
  distinct(country, pair_id, category)

pair_category_tables <- list(
  assignment_type = assignment_type_pairs,
  allocation_nature = allocation_nature_pairs,
  is_purpose_restricted = purpose_restriction_pairs,
  beneficiary_type = beneficiary_type_pairs
)

stopifnot(
  all(map_int(pair_category_tables, nrow) == nrow(allocation_pairs)),
  all(map_int(
    pair_category_tables,
    ~ nrow(distinct(.x, country, pair_id))
  ) == nrow(allocation_pairs))
)

missing_category_audit <- imap_dfr(
  pair_category_tables,
  function(data, variable) {
    data %>%
      summarise(
        variable = variable,
        n_missing_allocations = sum(category == "Missing"),
        n_countries_with_missing = n_distinct(country[category == "Missing"]),
        missing_pairs = {
          values <- pair_id[category == "Missing"]
          if (length(values) == 0L) {
            NA_character_
          } else {
            paste(values, collapse = "; ")
          }
        }
      )
  }
)

expected_categories <- list(
  assignment_type = c(
    "earmark_functional",
    "hybrid_devolved_earmark",
    "equalization_transfer",
    "community_levy_external",
    "tax_sharing_specific",
    "tax_sharing_general",
    "derivation_transfer"
  ),
  allocation_nature = c(
    "proceeds_share", "cost_recovery_component"
  ),
  is_purpose_restricted = c("0", "1"),
  beneficiary_type = c(
    "fund",
    "agency",
    "ministry",
    "collectivite_territoriale",
    "sector",
    "public_employees",
    "program",
    "supranational",
    "general_budget"
  )
)

category_value_audit <- imap_dfr(
  pair_category_tables,
  function(data, variable) {
    expected_values <- expected_categories[[variable]]

    data %>%
      count(category, name = "n_allocations") %>%
      mutate(
        variable = variable,
        status = case_when(
          category == "Missing" ~ "Missing",
          category %in% expected_values ~ "expected",
          TRUE ~ "unexpected_preserved"
        ),
        .before = 1
      )
  }
) %>%
  arrange(variable, status, category)

unexpected_category_values <- category_value_audit %>%
  filter(status == "unexpected_preserved")
```

```{r display-missing-category-audit}
display_table(missing_category_audit)
```

```{r display-category-value-audit}
display_table(category_value_audit)
```

```{r display-unexpected-category-values}
display_table(unexpected_category_values)
```

# 4. Assignment type

```{r compute-assignment-type-summary}
assignment_type_result <- make_composition_summary(
  assignment_type_pairs,
  country_universe,
  country_pair_totals
)
assignment_type_summary <- assignment_type_result$summary
```

```{r display-assignment-type-summary}
display_table(assignment_type_summary)
```

# 5. Allocation nature

```{r compute-allocation-nature-summary}
allocation_nature_result <- make_composition_summary(
  allocation_nature_pairs,
  country_universe,
  country_pair_totals
)
allocation_nature_summary <- allocation_nature_result$summary
```

```{r display-allocation-nature-summary}
display_table(allocation_nature_summary)
```

# 6. Purpose restriction

```{r compute-purpose-restriction-summary}
purpose_restriction_result <- make_composition_summary(
  purpose_restriction_pairs,
  country_universe,
  country_pair_totals
)
purpose_restriction_summary <- purpose_restriction_result$summary
```

```{r display-purpose-restriction-summary}
display_table(purpose_restriction_summary)
```

# 7. Beneficiary type

```{r compute-beneficiary-type-summary}
beneficiary_type_result <- make_composition_summary(
  beneficiary_type_pairs,
  country_universe,
  country_pair_totals
)
beneficiary_type_summary <- beneficiary_type_result$summary
```

```{r display-beneficiary-type-summary}
display_table(beneficiary_type_summary)
```

# 8. Destination-function long dataset

Destination function is multi-valued. Primary values and semicolon-separated detail values are pooled across all source rows for a pair, trimmed, emptied values removed, and then deduplicated at `country + pair_id + destination_function`. `Missing` is assigned only if the pair has no usable value in either function field.

```{r prepare-destination-functions}
primary_function_values <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    country,
    pair_id,
    destination_function = str_trim(destination_function)
  ) %>%
  filter(!is_missing_value(destination_function))

detail_function_values <- SSA_EARMARK_ALLOCATIONS %>%
  select(country, pair_id, destination_function_detail) %>%
  separate_rows(destination_function_detail, sep = ";", convert = FALSE) %>%
  transmute(
    country,
    pair_id,
    destination_function = str_trim(destination_function_detail)
  ) %>%
  filter(!is_missing_value(destination_function))

observed_pair_functions <- bind_rows(
  primary_function_values,
  detail_function_values
) %>%
  distinct(country, pair_id, destination_function)

pairs_without_function <- allocation_pairs %>%
  anti_join(
    observed_pair_functions %>% distinct(country, pair_id),
    by = c("country", "pair_id")
  ) %>%
  mutate(destination_function = "Missing")

pair_function_values <- bind_rows(
  observed_pair_functions,
  pairs_without_function
) %>%
  distinct(country, pair_id, destination_function) %>%
  arrange(country, pair_id, destination_function)

multi_function_pair_audit <- pair_function_values %>%
  count(country, pair_id, name = "n_destination_functions") %>%
  filter(n_destination_functions > 1L) %>%
  left_join(
    pair_function_values %>%
      group_by(country, pair_id) %>%
      summarise(
        destination_functions = paste(
          destination_function,
          collapse = "; "
        ),
        .groups = "drop"
      ),
    by = c("country", "pair_id")
  ) %>%
  arrange(country, pair_id)

primary_function_multiplicity <- primary_function_values %>%
  distinct(country, pair_id, destination_function) %>%
  group_by(country, pair_id) %>%
  summarise(
    n_primary_functions = n(),
    primary_functions = paste(destination_function, collapse = "; "),
    .groups = "drop"
  ) %>%
  filter(n_primary_functions > 1L) %>%
  arrange(country, pair_id)

mapping_lookup <- COFOG_GFSM2014_mapping %>%
  transmute(
    destination_function = str_trim(cofog_code),
    destination_function_label = cofog_name,
    mapping_status = "mapped_gfsm"
  )

mapping_key_audit <- mapping_lookup %>%
  group_by(destination_function) %>%
  summarise(
    n_rows = n(),
    n_labels = n_distinct(destination_function_label),
    labels = paste(
      sort(unique(destination_function_label)),
      collapse = " | "
    ),
    .groups = "drop"
  )

if (
  any(is_missing_value(mapping_lookup$destination_function)) ||
    any(is_missing_value(mapping_lookup$destination_function_label)) ||
    any(mapping_key_audit$n_rows > 1L) ||
    any(mapping_key_audit$n_labels > 1L)
) {
  print(mapping_key_audit %>% filter(n_rows > 1L | n_labels > 1L))
  stop(
    "COFOG/GFSM mapping keys or labels are missing, duplicated, or conflicting."
  )
}

special_function_lookup <- tribble(
  ~destination_function, ~destination_function_label, ~mapping_status,
  "general_budget", "General budget", "preserved_non_gfsm",
  "local_budget", "Local budget", "preserved_non_gfsm",
  "unallocated", "Unallocated", "preserved_non_gfsm",
  "Missing", "Missing", "preserved_non_gfsm"
)

function_label_lookup <- bind_rows(
  mapping_lookup,
  special_function_lookup
)

destination_function_long <- pair_function_values %>%
  left_join(function_label_lookup, by = "destination_function") %>%
  mutate(
    mapping_status = replace_na(mapping_status, "unmapped"),
    destination_function_label = if_else(
      mapping_status == "unmapped",
      paste0("Unmapped code: ", destination_function),
      destination_function_label
    )
  ) %>%
  select(
    country,
    pair_id,
    destination_function,
    destination_function_label,
    mapping_status
  ) %>%
  arrange(country, pair_id, destination_function)

unmapped_function_codes <- destination_function_long %>%
  filter(mapping_status == "unmapped") %>%
  group_by(destination_function, destination_function_label) %>%
  summarise(
    n_allocation_links = n(),
    n_countries = n_distinct(country),
    countries = paste(sort(unique(country)), collapse = "; "),
    .groups = "drop"
  ) %>%
  arrange(destination_function)

destination_function_long_audit <- tibble(
  measure = c(
    "distinct_allocation_pairs",
    "distinct_pair_function_links",
    "pairs_with_multiple_functions",
    "pairs_assigned_Missing",
    "unmapped_function_codes"
  ),
  value = c(
    nrow(allocation_pairs),
    nrow(destination_function_long),
    nrow(multi_function_pair_audit),
    sum(destination_function_long$destination_function == "Missing"),
    nrow(unmapped_function_codes)
  )
)

stopifnot(
  !any(duplicated(
    destination_function_long[c(
      "country", "pair_id", "destination_function"
    )]
  )),
  nrow(distinct(destination_function_long, country, pair_id)) ==
    nrow(allocation_pairs)
)
```

## Primary-function multiplicity

This audit is non-fatal because destination function is explicitly multi-valued. Every observed function is retained in the long table.

```{r display-primary-function-multiplicity}
display_table(primary_function_multiplicity)
```

## Long-table audit

```{r display-destination-long-audit}
display_table(destination_function_long_audit)
```

```{r display-multi-function-pairs}
display_table(multi_function_pair_audit)
```

## Unmapped function codes

```{r display-unmapped-function-codes}
display_table(unmapped_function_codes)
```

## Destination-function long table

```{r display-destination-function-long}
display_table(destination_function_long)
```

# 9. Destination-function composition

The within-country denominator remains the total number of distinct allocation pairs in each country. Because one allocation may finance several functions, destination-function counts and shares can sum to more than the total number or share of allocations.

```{r compute-destination-function-summary}
destination_function_result <- make_composition_summary(
  destination_function_long %>%
    transmute(
      country,
      pair_id,
      category = destination_function
    ),
  country_universe,
  country_pair_totals
)

destination_function_summary <- destination_function_result$summary %>%
  left_join(
    destination_function_long %>%
      distinct(
        category = destination_function,
        category_label = destination_function_label,
        mapping_status
      ),
    by = "category"
  ) %>%
  select(
    category,
    category_label,
    pooled_n_allocations,
    n_countries_using,
    median_allocations_among_users,
    mean_within_country_share,
    mapping_status
  )
```

```{r display-destination-function-summary}
display_table(destination_function_summary)
```

# 10. Final validation

For each single-valued composition, the pooled category count must equal the 761 distinct allocations and each country's category shares must sum to one. These identities do not apply to the multi-valued destination-function table.

```{r final-validation}
ordinary_results <- list(
  assignment_type = assignment_type_result,
  allocation_nature = allocation_nature_result,
  is_purpose_restricted = purpose_restriction_result,
  beneficiary_type = beneficiary_type_result
)

ordinary_summary_validation <- imap_dfr(
  ordinary_results,
  function(result, variable) {
    country_share_sums <- result$country_category_grid %>%
      group_by(country) %>%
      summarise(
        share_sum = sum(within_country_share),
        .groups = "drop"
      )

    tibble(
      variable = variable,
      pooled_count = sum(result$summary$pooled_n_allocations),
      expected_allocation_count = nrow(allocation_pairs),
      pooled_count_matches = pooled_count == expected_allocation_count,
      n_country_share_sum_problems = sum(
        abs(country_share_sums$share_sum - 1) > 1e-12
      ),
      status = if_else(
        pooled_count_matches & n_country_share_sum_problems == 0L,
        "passed",
        "failed"
      )
    )
  }
)

destination_denominator_validation <-
  destination_function_result$country_category_grid %>%
  group_by(country) %>%
  summarise(
    denominator_values = n_distinct(total_allocations),
    denominator_used = first(total_allocations),
    .groups = "drop"
  ) %>%
  left_join(
    country_pair_totals %>%
      rename(expected_total_allocations = total_allocations),
    by = "country"
  ) %>%
  mutate(
    denominator_matches =
      denominator_values == 1L &
      denominator_used == expected_total_allocations
  )

fatal_validation_problems <- bind_rows(
  ordinary_summary_validation %>%
    filter(status == "failed") %>%
    transmute(
      source = variable,
      problem = "ordinary composition identity failed"
    ),
  destination_denominator_validation %>%
    filter(!denominator_matches) %>%
    transmute(
      source = country,
      problem = "destination denominator differs from allocation-pair total"
    )
)

if (nrow(fatal_validation_problems) > 0L) {
  print(fatal_validation_problems)
  stop("Final validation failed. No summary output was written.")
}
```

```{r display-ordinary-summary-validation}
display_table(ordinary_summary_validation)
```

```{r display-destination-denominator-validation}
display_table(destination_denominator_validation)
```

```{r display-fatal-validation-problems}
display_table(fatal_validation_problems)
```

# 11. Export and verify the six requested tables

```{r export-results}
output_tables <- list(
  assignment_type = assignment_type_summary,
  allocation_nature = allocation_nature_summary,
  purpose_restriction = purpose_restriction_summary,
  beneficiary_type = beneficiary_type_summary,
  destination_function_long = destination_function_long,
  destination_function_summary = destination_function_summary
)

walk2(
  output_tables,
  output_paths[names(output_tables)],
  ~ write_csv(.x, .y, na = "")
)

output_verification <- imap_dfr(output_tables, function(data, output_name) {
  output_column_types <- map(data, function(column) {
    if (is.character(column)) {
      col_character()
    } else if (is.integer(column)) {
      col_integer()
    } else if (is.double(column)) {
      col_double()
    } else if (is.logical(column)) {
      col_logical()
    } else {
      col_guess()
    }
  })

  saved_data <- read_csv(
    output_paths[[output_name]],
    col_types = do.call(cols, output_column_types),
    show_col_types = FALSE,
    progress = FALSE
  )

  tibble(
    output = output_name,
    file = basename(output_paths[[output_name]]),
    n_rows = nrow(saved_data),
    n_columns = ncol(saved_data),
    row_count_matches = nrow(saved_data) == nrow(data),
    column_names_match = identical(names(saved_data), names(data)),
    values_match = isTRUE(all.equal(
      saved_data,
      data,
      check.attributes = FALSE,
      tolerance = 1e-12
    )),
    file_exists = file.exists(output_paths[[output_name]])
  )
})

if (
  any(!output_verification$row_count_matches) ||
    any(!output_verification$column_names_match) ||
    any(!output_verification$values_match) ||
    any(!output_verification$file_exists)
) {
  print(output_verification)
  stop("At least one exported CSV failed read-back verification.")
}
```

```{r display-output-verification}
display_table(output_verification)
```

# 12. Run summary

```{r run-summary, results='asis'}
missing_summary_text <- missing_category_audit %>%
  transmute(text = paste0(variable, "=", n_missing_allocations)) %>%
  pull(text) %>%
  paste(collapse = "; ")

unexpected_summary_text <- if (nrow(unexpected_category_values) == 0L) {
  "none"
} else {
  unexpected_category_values %>%
    transmute(text = paste0(variable, ":", category, " (", n_allocations, ")")) %>%
    pull(text) %>%
    paste(collapse = "; ")
}

cat(
  "- Analysis variant: ", analysis_variant, ".\n",
  "- Countries included (", nrow(country_universe), "): ",
  paste(country_universe$country, collapse = ", "), ".\n",
  "- Allocation-table rows: ", nrow(SSA_EARMARK_ALLOCATIONS),
  "; distinct country + pair_id allocations: ", nrow(allocation_pairs), ".\n",
  "- Missing pair-level categories: ", missing_summary_text,
  "; destination_function=",
  sum(destination_function_long$destination_function == "Missing"), ".\n",
  "- Distinct destination-function links: ",
  nrow(destination_function_long),
  "; allocations with multiple functions: ",
  nrow(multi_function_pair_audit), ".\n",
  "- Unmapped destination-function codes: ",
  nrow(unmapped_function_codes), ".\n",
  "- Fatal validation problems: ", nrow(fatal_validation_problems),
  "; unexpected category values preserved: ",
  unexpected_summary_text, ".\n",
  sep = ""
)
```
