---
title: "Allocation-source earmark status classification"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook classifies each allocation-source instrument from the original, non-expanded allocation table. It appends only `earmark_status` to the two derived allocation files. Existing fields, missing values, rows, and row order are preserved exactly. The expanded table is never used to count allocation channels.

```{r setup, include=FALSE}
input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

target_paths <- c(
  allocation = file.path(
    project_root, "data", "derived", "SSA_EARMARK_ALLOCATIONS.csv"
  ),
  expanded = file.path(
    project_root,
    "data",
    "derived",
    "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  )
)

required_allocation_columns <- c(
  "country", "instrument_id", "pair_id", "allocation_nature",
  "is_purpose_restricted", "beneficiary_type"
)

read_all_character <- function(path) {
  read_csv(
    path,
    col_types = cols(.default = col_character()),
    na = "",
    trim_ws = FALSE,
    name_repair = "minimal",
    show_col_types = FALSE,
    progress = FALSE
  )
}

display_table <- function(data) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    knitr::kable(data, row.names = FALSE)
  }
}

is_missing_key <- function(values) {
  is.na(values) | str_trim(values) == ""
}

remove_previous_status <- function(data, dataset_name) {
  status_positions <- which(names(data) == "earmark_status")

  if (length(status_positions) > 1L) {
    stop(dataset_name, " contains more than one earmark_status column.")
  }
  if (length(status_positions) == 1L && status_positions != ncol(data)) {
    stop(dataset_name, " has earmark_status in a non-final column position.")
  }

  data %>% select(-any_of("earmark_status"))
}

same_existing_values_and_order <- function(before, after) {
  old_columns <- names(before)
  identical(names(after)[seq_along(old_columns)], old_columns) &&
    nrow(before) == nrow(after) &&
    all(vapply(
      old_columns,
      function(column) identical(before[[column]], after[[column]]),
      logical(1)
    ))
}

missing_by_column <- function(data) {
  vapply(data, function(column) sum(is.na(column)), integer(1))
}

duplicate_row_count <- function(data) {
  sum(duplicated(as.data.frame(data, stringsAsFactors = FALSE)))
}

make_table_checks <- function(dataset_name, before_data, after_data) {
  old_columns <- names(before_data)
  old_values_same <- same_existing_values_and_order(before_data, after_data)
  old_missing_same <- identical(
    missing_by_column(before_data),
    missing_by_column(after_data %>% select(all_of(old_columns)))
  )
  status_by_key <- after_data %>%
    filter(!is_missing_key(country), !is_missing_key(instrument_id)) %>%
    group_by(country, instrument_id) %>%
    summarise(
      n_status_values = n_distinct(earmark_status, na.rm = TRUE),
      .groups = "drop"
    )

  tibble(
    dataset = dataset_name,
    check = c(
      "rows before = rows after",
      "existing columns unchanged and in the same order",
      "existing values and row order unchanged",
      "no duplicate rows introduced by the merge",
      "existing-column missingness unchanged",
      "exactly one status per non-missing country + instrument_id",
      "earmark_status has no missing values",
      "output is original columns plus one appended column"
    ),
    before = c(
      as.character(nrow(before_data)),
      paste0(length(old_columns), " original columns"),
      "original field vectors",
      as.character(duplicate_row_count(before_data)),
      "per-column missing counts",
      as.character(n_distinct(paste(
        before_data$country, before_data$instrument_id
      ))),
      "not applicable",
      paste0(length(old_columns), " columns")
    ),
    after = c(
      as.character(nrow(after_data)),
      paste0(length(old_columns), " original columns"),
      if_else(old_values_same, "identical", "changed"),
      as.character(duplicate_row_count(
        after_data %>% select(all_of(old_columns))
      )),
      if_else(old_missing_same, "identical", "changed"),
      paste0(
        nrow(status_by_key), " keys; max distinct status = ",
        ifelse(nrow(status_by_key) == 0L, NA_integer_, max(status_by_key$n_status_values))
      ),
      as.character(sum(is.na(after_data$earmark_status))),
      paste0(ncol(after_data), " columns; earmark_status last")
    ),
    passed = c(
      nrow(before_data) == nrow(after_data),
      identical(names(after_data)[seq_along(old_columns)], old_columns),
      old_values_same,
      duplicate_row_count(before_data) == duplicate_row_count(
        after_data %>% select(all_of(old_columns))
      ),
      old_missing_same,
      nrow(status_by_key) == n_distinct(paste(
        before_data$country, before_data$instrument_id
      )) &&
        all(status_by_key$n_status_values == 1L),
      !any(is.na(after_data$earmark_status)),
      identical(names(after_data), c(old_columns, "earmark_status"))
    )
  )
}
```

# 1. Load and validate the two allocation tables

All columns are imported as character values so identifiers, numeric-looking legal text, accents, and existing missing cells can be written back without type conversion.

```{r load-and-validate-inputs}
if (!all(file.exists(target_paths))) {
  stop(
    "Missing required input file(s): ",
    paste(target_paths[!file.exists(target_paths)], collapse = "; ")
  )
}

allocation_loaded <- read_all_character(target_paths[["allocation"]])
expanded_loaded <- read_all_character(target_paths[["expanded"]])

SSA_EARMARK_ALLOCATIONS_before <- remove_previous_status(
  allocation_loaded, "SSA_EARMARK_ALLOCATIONS"
)
SSA_EARMARK_ALLOCATIONS_EXPANDED_before <- remove_previous_status(
  expanded_loaded, "SSA_EARMARK_ALLOCATIONS_EXPANDED"
)

missing_required <- setdiff(
  required_allocation_columns,
  names(SSA_EARMARK_ALLOCATIONS_before)
)
missing_expanded_required <- setdiff(
  c("country", "instrument_id", "pair_id"),
  names(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
)

if (length(missing_required) > 0L || length(missing_expanded_required) > 0L) {
  stop(
    "Required columns are missing. Allocation: ",
    paste(missing_required, collapse = "; "),
    "; Expanded: ",
    paste(missing_expanded_required, collapse = "; ")
  )
}

key_missingness <- bind_rows(
  tibble(
    dataset = "SSA_EARMARK_ALLOCATIONS",
    field = c("country", "instrument_id", "pair_id"),
    n_missing = c(
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_before$country)),
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_before$instrument_id)),
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_before$pair_id))
    )
  ),
  tibble(
    dataset = "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    field = c("country", "instrument_id", "pair_id"),
    n_missing = c(
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_EXPANDED_before$country)),
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_EXPANDED_before$instrument_id)),
      sum(is_missing_key(SSA_EARMARK_ALLOCATIONS_EXPANDED_before$pair_id))
    )
  )
)

if (any(key_missingness$n_missing > 0L)) {
  print(key_missingness %>% filter(n_missing > 0L))
  stop("Missing merge or counting keys prevent a complete classification.")
}

pair_instrument_conflicts <- SSA_EARMARK_ALLOCATIONS_before %>%
  distinct(country, pair_id, instrument_id) %>%
  count(country, pair_id, name = "n_instruments") %>%
  filter(n_instruments > 1L)

if (nrow(pair_instrument_conflicts) > 0L) {
  print(pair_instrument_conflicts)
  stop("At least one allocation pair maps to several source instruments.")
}

expanded_keys_missing_from_allocation <-
  SSA_EARMARK_ALLOCATIONS_EXPANDED_before %>%
  distinct(country, instrument_id) %>%
  anti_join(
    SSA_EARMARK_ALLOCATIONS_before %>% distinct(country, instrument_id),
    by = c("country", "instrument_id")
  )

if (nrow(expanded_keys_missing_from_allocation) > 0L) {
  print(expanded_keys_missing_from_allocation)
  stop("Expanded allocation-source keys are absent from the original table.")
}

classification_input_missingness <- tibble(
  field = c(
    "allocation_nature", "is_purpose_restricted", "beneficiary_type"
  ),
  n_missing_rows = c(
    sum(is.na(SSA_EARMARK_ALLOCATIONS_before$allocation_nature)),
    sum(is.na(SSA_EARMARK_ALLOCATIONS_before$is_purpose_restricted)),
    sum(is.na(SSA_EARMARK_ALLOCATIONS_before$beneficiary_type))
  )
)
```

```{r display-input-audit}
input_inventory <- tibble(
  dataset = c(
    "SSA_EARMARK_ALLOCATIONS", "SSA_EARMARK_ALLOCATIONS_EXPANDED"
  ),
  rows = c(
    nrow(SSA_EARMARK_ALLOCATIONS_before),
    nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
  ),
  original_columns = c(
    ncol(SSA_EARMARK_ALLOCATIONS_before),
    ncol(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
  ),
  distinct_source_instruments = c(
    n_distinct(paste(
      SSA_EARMARK_ALLOCATIONS_before$country,
      SSA_EARMARK_ALLOCATIONS_before$instrument_id
    )),
    n_distinct(paste(
      SSA_EARMARK_ALLOCATIONS_EXPANDED_before$country,
      SSA_EARMARK_ALLOCATIONS_EXPANDED_before$instrument_id
    ))
  )
)

display_table(input_inventory)
display_table(key_missingness)
display_table(classification_input_missingness)
```

# 2. Classify allocation-source instruments

The two channel flags exist only inside this computation. Only flags that are explicitly `TRUE` enter a channel count; pre-existing missing source values remain missing and are not imputed.

```{r compute-earmark-status}
allocation_classification_rows <- SSA_EARMARK_ALLOCATIONS_before %>%
  mutate(
    restricted_channel =
      allocation_nature == "proceeds_share" &
      is_purpose_restricted == "1" &
      beneficiary_type != "general_budget",
    unrestricted_channel =
      allocation_nature == "proceeds_share" &
      (
        is_purpose_restricted == "0" |
        beneficiary_type == "general_budget"
      )
  )

instrument_classification_audit <- allocation_classification_rows %>%
  group_by(country, instrument_id) %>%
  summarise(
    n_restricted_channels = n_distinct(
      pair_id[restricted_channel %in% TRUE], na.rm = TRUE
    ),
    n_unrestricted_channels = n_distinct(
      pair_id[unrestricted_channel %in% TRUE], na.rm = TRUE
    ),
    n_proceeds_channels = n_distinct(
      pair_id[allocation_nature == "proceeds_share"], na.rm = TRUE
    ),
    .groups = "drop"
  ) %>%
  mutate(
    earmark_status = case_when(
      n_proceeds_channels == 0L ~ "cost_recovery_only",
      n_restricted_channels == 0L ~ "assigned_unrestricted",
      n_unrestricted_channels > 0L ~ "partial_earmark",
      TRUE ~ "exclusive_earmark"
    )
  )

instrument_earmark_status <- instrument_classification_audit %>%
  select(country, instrument_id, earmark_status)

allowed_statuses <- c(
  "cost_recovery_only", "assigned_unrestricted",
  "partial_earmark", "exclusive_earmark"
)

classification_map_checks <- tibble(
  check = c(
    "one classification row per country + instrument_id",
    "all source instruments classified",
    "no missing earmark_status",
    "only requested status categories used",
    "all expanded source-instrument keys covered"
  ),
  passed = c(
    nrow(instrument_earmark_status) ==
      n_distinct(paste(
        instrument_earmark_status$country,
        instrument_earmark_status$instrument_id
      )),
    nrow(instrument_earmark_status) ==
      n_distinct(paste(
        SSA_EARMARK_ALLOCATIONS_before$country,
        SSA_EARMARK_ALLOCATIONS_before$instrument_id
      )),
    !any(is.na(instrument_earmark_status$earmark_status)),
    all(instrument_earmark_status$earmark_status %in% allowed_statuses),
    nrow(expanded_keys_missing_from_allocation) == 0L
  )
)

if (!all(classification_map_checks$passed)) {
  print(classification_map_checks %>% filter(!passed))
  stop("Instrument-level classification checks failed.")
}

incomplete_classification_inputs <- allocation_classification_rows %>%
  filter(
    is.na(allocation_nature) |
      is.na(is_purpose_restricted) |
      is.na(beneficiary_type)
  ) %>%
  distinct(country, instrument_id) %>%
  left_join(
    instrument_classification_audit,
    by = c("country", "instrument_id")
  ) %>%
  arrange(country, instrument_id)
```

```{r display-classification-audit}
display_table(classification_map_checks)
display_table(incomplete_classification_inputs)
```

# 3. Append only `earmark_status`

```{r append-earmark-status}
append_earmark_status <- function(data, status_map) {
  original_columns <- names(data)

  data %>%
    mutate(.original_row_order = row_number()) %>%
    left_join(
      status_map,
      by = c("country", "instrument_id"),
      relationship = "many-to-one"
    ) %>%
    arrange(.original_row_order) %>%
    select(all_of(original_columns), earmark_status)
}

SSA_EARMARK_ALLOCATIONS <- append_earmark_status(
  SSA_EARMARK_ALLOCATIONS_before,
  instrument_earmark_status
)
SSA_EARMARK_ALLOCATIONS_EXPANDED <- append_earmark_status(
  SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
  instrument_earmark_status
)

prewrite_checks <- bind_rows(
  make_table_checks(
    "SSA_EARMARK_ALLOCATIONS",
    SSA_EARMARK_ALLOCATIONS_before,
    SSA_EARMARK_ALLOCATIONS
  ),
  make_table_checks(
    "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
    SSA_EARMARK_ALLOCATIONS_EXPANDED
  )
)

if (!all(prewrite_checks$passed)) {
  print(prewrite_checks %>% filter(!passed))
  stop("Pre-write preservation checks failed; no output was overwritten.")
}
```

# 4. Stage, verify, and overwrite both CSV files

Both candidate files are written to temporary paths in `data/derived`, read back, and fully validated before either target is overwritten. The two-file commit keeps temporary backups and rolls both targets back if a final check fails.

```{r stage-and-commit-outputs}
stage_paths <- c(
  allocation = tempfile(
    pattern = ".SSA_EARMARK_ALLOCATIONS_",
    tmpdir = dirname(target_paths[["allocation"]]),
    fileext = ".csv"
  ),
  expanded = tempfile(
    pattern = ".SSA_EARMARK_ALLOCATIONS_EXPANDED_",
    tmpdir = dirname(target_paths[["expanded"]]),
    fileext = ".csv"
  )
)

write_csv(SSA_EARMARK_ALLOCATIONS, stage_paths[["allocation"]], na = "")
write_csv(
  SSA_EARMARK_ALLOCATIONS_EXPANDED,
  stage_paths[["expanded"]],
  na = ""
)

staged_allocation <- read_all_character(stage_paths[["allocation"]])
staged_expanded <- read_all_character(stage_paths[["expanded"]])

staged_checks <- bind_rows(
  make_table_checks(
    "SSA_EARMARK_ALLOCATIONS",
    SSA_EARMARK_ALLOCATIONS_before,
    staged_allocation
  ),
  make_table_checks(
    "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
    staged_expanded
  )
)

if (!all(staged_checks$passed)) {
  file.remove(stage_paths[file.exists(stage_paths)])
  print(staged_checks %>% filter(!passed))
  stop("Staged-file checks failed; no output was overwritten.")
}

commit_two_files <- function(stage_paths, target_paths) {
  backup_paths <- c(
    allocation = tempfile(
      pattern = ".SSA_EARMARK_ALLOCATIONS_backup_",
      tmpdir = dirname(target_paths[["allocation"]]),
      fileext = ".csv"
    ),
    expanded = tempfile(
      pattern = ".SSA_EARMARK_ALLOCATIONS_EXPANDED_backup_",
      tmpdir = dirname(target_paths[["expanded"]]),
      fileext = ".csv"
    )
  )
  committed <- FALSE

  on.exit({
    if (!committed) {
      for (name in names(target_paths)) {
        if (file.exists(backup_paths[[name]])) {
          file.copy(
            backup_paths[[name]], target_paths[[name]], overwrite = TRUE
          )
        }
      }
    }
    file.remove(stage_paths[file.exists(stage_paths)])
    file.remove(backup_paths[file.exists(backup_paths)])
  }, add = TRUE)

  backup_ok <- vapply(
    names(target_paths),
    function(name) file.copy(
      target_paths[[name]], backup_paths[[name]], overwrite = FALSE
    ),
    logical(1)
  )
  if (!all(backup_ok)) {
    stop("Could not create both temporary rollback copies.")
  }

  commit_ok <- vapply(
    names(target_paths),
    function(name) file.copy(
      stage_paths[[name]], target_paths[[name]], overwrite = TRUE
    ),
    logical(1)
  )
  if (!all(commit_ok)) {
    stop("Could not overwrite both target CSV files.")
  }

  final_allocation <- read_all_character(target_paths[["allocation"]])
  final_expanded <- read_all_character(target_paths[["expanded"]])
  final_checks <- bind_rows(
    make_table_checks(
      "SSA_EARMARK_ALLOCATIONS",
      SSA_EARMARK_ALLOCATIONS_before,
      final_allocation
    ),
    make_table_checks(
      "SSA_EARMARK_ALLOCATIONS_EXPANDED",
      SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
      final_expanded
    )
  )

  if (!all(final_checks$passed)) {
    print(final_checks %>% filter(!passed))
    stop("Final read-back checks failed; restoring both prior files.")
  }

  committed <- TRUE
  list(
    allocation = final_allocation,
    expanded = final_expanded,
    checks = final_checks
  )
}

committed_outputs <- commit_two_files(stage_paths, target_paths)
final_sanity_checks <- committed_outputs$checks
```

# 5. Inline sanity checks and status distribution

```{r display-final-sanity-checks}
display_table(final_sanity_checks)
```

The distribution below counts allocation-source instruments, not allocation rows or expanded Census links.

```{r compute-status-distribution}
earmark_status_distribution <- bind_rows(
  instrument_earmark_status %>%
    count(country, earmark_status, name = "n_instruments"),
  instrument_earmark_status %>%
    count(earmark_status, name = "n_instruments") %>%
    mutate(country = "Overall") %>%
    select(country, earmark_status, n_instruments)
) %>%
  arrange(country == "Overall", country, earmark_status)
```

```{r display-status-distribution}
display_table(earmark_status_distribution)
```

# 6. Run summary

```{r run-summary, results='asis'}
overall_distribution <- earmark_status_distribution %>%
  filter(country == "Overall")

cat(
  "- Files updated: `data/derived/SSA_EARMARK_ALLOCATIONS.csv` and ",
  "`data/derived/SSA_EARMARK_ALLOCATIONS_EXPANDED.csv`.\n",
  "- Rows preserved: ", nrow(committed_outputs$allocation),
  " Allocation rows and ", nrow(committed_outputs$expanded),
  " expanded rows.\n",
  "- Allocation-source instruments classified: ",
  nrow(instrument_earmark_status), ".\n",
  "- Missing `earmark_status` values: 0 in both files.\n",
  "- Existing columns, field values, missingness, duplicate-row counts, and ",
  "row order were unchanged.\n",
  "- Overall distribution: ",
  paste0(
    overall_distribution$earmark_status,
    " = ", overall_distribution$n_instruments,
    collapse = "; "
  ),
  ".\n",
  sep = ""
)
```
