---
title: "Whether statutory allocation amounts are determined in law"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: paged
---

This notebook creates the requested `commitment_stated` indicator in memory and
summarises it at the distinct `country + pair_id` allocation-channel grain. The
input allocation file is not overwritten, no existing field is changed, and no
input row is removed. The database describes de jure statutory rules rather
than actual collections, transfers, disbursements, or expenditure execution.

```{r setup, include=FALSE}
if (.Platform$OS.type == "windows") {
  active_utf8_locale <- Sys.setlocale(
    "LC_CTYPE",
    "English_United States.utf8"
  )
  if (identical(active_utf8_locale, "")) {
    stop("A UTF-8 locale is required to preserve accented legal text.")
  }
}

input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(readr)
library(stringr)
library(tibble)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

allocation_path <- file.path(
  project_root,
  "data",
  "derived",
  "SSA_EARMARK_ALLOCATIONS.csv"
)
general_output_path <- file.path(
  project_root,
  "analysis",
  "tables",
  "general",
  "commitment_stated_by_country.csv"
)
partial_exclusive_output_path <- file.path(
  project_root,
  "analysis",
  "tables",
  "partial_exclusive",
  "commitment_stated_by_country.csv"
)

dir.create(dirname(general_output_path), recursive = TRUE, showWarnings = FALSE)
dir.create(
  dirname(partial_exclusive_output_path),
  recursive = TRUE,
  showWarnings = FALSE
)

display_table <- function(data, digits = 6) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    print(knitr::kable(data, digits = digits, row.names = FALSE))
    cat("\n")
  }
}

is_missing_text <- function(values) {
  is.na(values) | str_trim(as.character(values)) == ""
}
```

# 1. Load and validate the allocation table

All columns are initially read as text so the classification audit can inspect
the source categories exactly as stored.

```{r load-and-validate-input}
if (!file.exists(allocation_path)) {
  stop("Missing input file: ", allocation_path)
}

input_file_before <- file.info(allocation_path)
input_md5_before <- unname(tools::md5sum(allocation_path))

SSA_EARMARK_ALLOCATIONS <- read_csv(
  allocation_path,
  col_types = cols(.default = col_character()),
  show_col_types = FALSE,
  progress = FALSE
)

required_columns <- c(
  "country",
  "pair_id",
  "share_basis",
  "earmark_status"
)
missing_columns <- setdiff(
  required_columns,
  names(SSA_EARMARK_ALLOCATIONS)
)

input_inventory <- tibble(
  input_file = "data/derived/SSA_EARMARK_ALLOCATIONS.csv",
  n_rows = nrow(SSA_EARMARK_ALLOCATIONS),
  n_columns = ncol(SSA_EARMARK_ALLOCATIONS),
  n_countries = n_distinct(SSA_EARMARK_ALLOCATIONS$country),
  n_distinct_country_pair = nrow(
    distinct(SSA_EARMARK_ALLOCATIONS, country, pair_id)
  ),
  missing_required_columns = if_else(
    length(missing_columns) == 0L,
    NA_character_,
    paste(missing_columns, collapse = "; ")
  )
)

if (length(missing_columns) > 0L) {
  print(input_inventory)
  stop(
    "Required allocation columns are missing: ",
    paste(missing_columns, collapse = "; ")
  )
}

missing_key_rows <- SSA_EARMARK_ALLOCATIONS %>%
  summarise(
    missing_country_rows = sum(is_missing_text(country)),
    missing_pair_id_rows = sum(is_missing_text(pair_id))
  )

if (any(unlist(missing_key_rows) > 0L)) {
  print(missing_key_rows)
  stop("Missing country or pair_id prevents complete allocation-channel counts.")
}
```

```{r display-input-inventory}
display_table(input_inventory)
display_table(missing_key_rows)
```

# 2. Audit `share_basis` before classification

The only expected non-missing values are `stated`,
`whole_proceeds_implied`, and `deferred_arrete`. Missing values are reported
explicitly because the requested rule classifies them as zero.

```{r audit-share-basis}
expected_share_basis <- c(
  "stated",
  "whole_proceeds_implied",
  "deferred_arrete"
)

share_basis_audit <- SSA_EARMARK_ALLOCATIONS %>%
  transmute(
    share_basis_value = if_else(
      is_missing_text(share_basis),
      "Missing",
      str_trim(share_basis)
    )
  ) %>%
  count(share_basis_value, name = "n_rows") %>%
  mutate(
    classification_rule = case_when(
      share_basis_value %in% c(
        "stated",
        "whole_proceeds_implied"
      ) ~ "commitment_stated = 1",
      share_basis_value %in% c(
        "deferred_arrete",
        "Missing"
      ) ~ "commitment_stated = 0",
      TRUE ~ "unexpected; not classified"
    ),
    category_status = if_else(
      share_basis_value == "Missing" |
        share_basis_value %in% expected_share_basis,
      "expected",
      "unexpected"
    )
  ) %>%
  arrange(category_status, share_basis_value)

unexpected_share_basis <- share_basis_audit %>%
  filter(category_status == "unexpected")
```

```{r display-share-basis-audit}
display_table(share_basis_audit)
display_table(unexpected_share_basis)
```

```{r stop-on-unexpected-share-basis}
if (nrow(unexpected_share_basis) > 0L) {
  stop(
    "Unexpected share_basis values were reported above. ",
    "No classification or output was created."
  )
}
```

# 3. Create `commitment_stated` without changing the input fields

```{r create-commitment-stated}
original_column_names <- names(SSA_EARMARK_ALLOCATIONS)

SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT <-
  SSA_EARMARK_ALLOCATIONS %>%
  mutate(
    commitment_stated = case_when(
      str_trim(share_basis) %in% c(
        "stated",
        "whole_proceeds_implied"
      ) ~ 1L,
      str_trim(share_basis) == "deferred_arrete" |
        is_missing_text(share_basis) ~ 0L,
      TRUE ~ NA_integer_
    )
  )

row_level_preservation <- tibble(
  check = c(
    "Rows before equal rows after",
    "Existing columns remain first and in their original order",
    "Existing field values are unchanged",
    "commitment_stated has no missing values"
  ),
  passed = c(
    nrow(SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT) ==
      nrow(SSA_EARMARK_ALLOCATIONS),
    identical(
      names(SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT),
      c(original_column_names, "commitment_stated")
    ),
    all(vapply(
      original_column_names,
      function(field) {
        identical(
          SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT[[field]],
          SSA_EARMARK_ALLOCATIONS[[field]]
        )
      },
      logical(1)
    )),
    !any(is.na(
      SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT$commitment_stated
    ))
  )
)

if (!all(row_level_preservation$passed)) {
  print(row_level_preservation)
  stop("Row-level preservation failed after deriving commitment_stated.")
}
```

```{r display-row-level-preservation}
display_table(row_level_preservation)
```

# 4. Establish the distinct allocation-channel grain

Repeated evidence rows must agree on the derived indicator and on
`earmark_status`. The notebook stops rather than choosing a value if one
`country + pair_id` has a conflict.

```{r prepare-pair-level-data}
pair_commitment_conflicts <-
  SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT %>%
  distinct(country, pair_id, commitment_stated) %>%
  count(country, pair_id, name = "n_commitment_values") %>%
  filter(n_commitment_values > 1L)

pair_status_conflicts <-
  SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT %>%
  mutate(
    earmark_status_audit = if_else(
      is_missing_text(earmark_status),
      "Missing",
      str_trim(earmark_status)
    )
  ) %>%
  distinct(country, pair_id, earmark_status_audit) %>%
  count(country, pair_id, name = "n_earmark_status_values") %>%
  filter(n_earmark_status_values > 1L)

if (nrow(pair_commitment_conflicts) > 0L) {
  print(pair_commitment_conflicts)
  stop("A country + pair_id has conflicting commitment_stated values.")
}
if (nrow(pair_status_conflicts) > 0L) {
  print(pair_status_conflicts)
  stop("A country + pair_id has conflicting earmark_status values.")
}

allocation_pairs_full <-
  SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT %>%
  distinct(country, pair_id, commitment_stated)

if (nrow(allocation_pairs_full) !=
    nrow(distinct(allocation_pairs_full, country, pair_id))) {
  stop("The full pair table is not unique at country + pair_id.")
}

partial_exclusive_statuses <- c(
  "partial_earmark",
  "exclusive_earmark"
)

SSA_EARMARK_ALLOCATIONS_PARTIAL_EXCLUSIVE <-
  SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT %>%
  filter(str_trim(earmark_status) %in% partial_exclusive_statuses)

allocation_pairs_partial_exclusive <-
  SSA_EARMARK_ALLOCATIONS_PARTIAL_EXCLUSIVE %>%
  distinct(country, pair_id, commitment_stated)

if (nrow(allocation_pairs_partial_exclusive) !=
    nrow(distinct(
      allocation_pairs_partial_exclusive,
      country,
      pair_id
    ))) {
  stop("The filtered pair table is not unique at country + pair_id.")
}

filter_audit <- tibble(
  analysis_scope = c(
    "All statutory allocations",
    "Partial and exclusive earmark instruments"
  ),
  filter_condition = c(
    "None",
    "earmark_status in partial_earmark or exclusive_earmark"
  ),
  n_rows = c(
    nrow(SSA_EARMARK_ALLOCATIONS_WITH_COMMITMENT),
    nrow(SSA_EARMARK_ALLOCATIONS_PARTIAL_EXCLUSIVE)
  ),
  n_distinct_country_pair = c(
    nrow(allocation_pairs_full),
    nrow(allocation_pairs_partial_exclusive)
  ),
  n_countries = c(
    n_distinct(allocation_pairs_full$country),
    n_distinct(allocation_pairs_partial_exclusive$country)
  )
)
```

```{r display-pair-and-filter-audits}
display_table(pair_commitment_conflicts)
display_table(pair_status_conflicts)
display_table(filter_audit)
```

# 5. Country summary — all statutory allocations

```{r compute-general-country-summary}
make_country_summary <- function(pair_data) {
  pair_data %>%
    group_by(country) %>%
    summarise(
      n_allocations = n(),
      n_commitment_stated_1 = sum(commitment_stated == 1L),
      n_commitment_stated_0 = sum(commitment_stated == 0L),
      share_commitment_stated = if_else(
        n_allocations > 0L,
        n_commitment_stated_1 / n_allocations,
        NA_real_
      ),
      .groups = "drop"
    ) %>%
    arrange(country)
}

commitment_stated_general <- make_country_summary(allocation_pairs_full)
```

```{r display-general-country-summary}
display_table(commitment_stated_general)
```

# 6. Country summary — partial and exclusive earmarks

```{r compute-partial-exclusive-country-summary}
commitment_stated_partial_exclusive <- make_country_summary(
  allocation_pairs_partial_exclusive
)
```

```{r display-partial-exclusive-country-summary}
display_table(commitment_stated_partial_exclusive)
```

# 7. Inline sanity checks

```{r validate-results}
make_scope_checks <- function(pair_data, country_summary, scope_name) {
  tibble(
    analysis_scope = scope_name,
    total_distinct_country_pair = nrow(pair_data),
    n_commitment_stated_1 = sum(pair_data$commitment_stated == 1L),
    n_commitment_stated_0 = sum(pair_data$commitment_stated == 0L),
    country_counts_reconcile =
      sum(country_summary$n_allocations) == nrow(pair_data),
    commitment_counts_reconcile = all(
      country_summary$n_commitment_stated_1 +
        country_summary$n_commitment_stated_0 ==
        country_summary$n_allocations
    ),
    shares_reconcile = all(
      abs(
        country_summary$share_commitment_stated -
          country_summary$n_commitment_stated_1 /
            country_summary$n_allocations
      ) < 1e-12
    ),
    pair_keys_unique =
      nrow(pair_data) == nrow(distinct(pair_data, country, pair_id))
  )
}

sanity_checks <- bind_rows(
  make_scope_checks(
    allocation_pairs_full,
    commitment_stated_general,
    "All statutory allocations"
  ),
  make_scope_checks(
    allocation_pairs_partial_exclusive,
    commitment_stated_partial_exclusive,
    "Partial and exclusive earmark instruments"
  )
) %>%
  mutate(
    no_unexpected_share_basis_classified =
      nrow(unexpected_share_basis) == 0L,
    no_pair_level_commitment_conflicts =
      nrow(pair_commitment_conflicts) == 0L
  )

fatal_check_columns <- c(
  "country_counts_reconcile",
  "commitment_counts_reconcile",
  "shares_reconcile",
  "pair_keys_unique",
  "no_unexpected_share_basis_classified",
  "no_pair_level_commitment_conflicts"
)

if (!all(unlist(sanity_checks[fatal_check_columns]))) {
  print(sanity_checks)
  stop("One or more commitment-stated sanity checks failed.")
}
```

```{r display-sanity-checks}
display_table(sanity_checks)
```

# 8. Export and verify the requested tables

```{r export-and-verify}
expected_output_columns <- c(
  "country",
  "n_allocations",
  "n_commitment_stated_1",
  "n_commitment_stated_0",
  "share_commitment_stated"
)

stopifnot(
  identical(names(commitment_stated_general), expected_output_columns),
  identical(
    names(commitment_stated_partial_exclusive),
    expected_output_columns
  )
)

write_csv(commitment_stated_general, general_output_path, na = "")
write_csv(
  commitment_stated_partial_exclusive,
  partial_exclusive_output_path,
  na = ""
)

output_col_types <- cols(
  country = col_character(),
  n_allocations = col_integer(),
  n_commitment_stated_1 = col_integer(),
  n_commitment_stated_0 = col_integer(),
  share_commitment_stated = col_double()
)

general_saved <- read_csv(
  general_output_path,
  col_types = output_col_types,
  show_col_types = FALSE,
  progress = FALSE
)
partial_exclusive_saved <- read_csv(
  partial_exclusive_output_path,
  col_types = output_col_types,
  show_col_types = FALSE,
  progress = FALSE
)

output_verification <- tibble(
  output_file = c(
    "analysis/tables/general/commitment_stated_by_country.csv",
    paste0(
      "analysis/tables/partial_exclusive/",
      "commitment_stated_by_country.csv"
    )
  ),
  file_exists = c(
    file.exists(general_output_path),
    file.exists(partial_exclusive_output_path)
  ),
  rows_expected = c(
    nrow(commitment_stated_general),
    nrow(commitment_stated_partial_exclusive)
  ),
  rows_saved = c(
    nrow(general_saved),
    nrow(partial_exclusive_saved)
  ),
  exact_columns = c(
    identical(names(general_saved), expected_output_columns),
    identical(
      names(partial_exclusive_saved),
      expected_output_columns
    )
  ),
  exact_values = c(
    isTRUE(all.equal(
      general_saved,
      commitment_stated_general,
      check.attributes = FALSE
    )),
    isTRUE(all.equal(
      partial_exclusive_saved,
      commitment_stated_partial_exclusive,
      check.attributes = FALSE
    ))
  )
)

input_file_after <- file.info(allocation_path)
input_md5_after <- unname(tools::md5sum(allocation_path))
input_preservation <- tibble(
  check = c(
    "Input MD5 unchanged",
    "Input size unchanged",
    "Input modification time unchanged"
  ),
  passed = c(
    identical(input_md5_before, input_md5_after),
    identical(input_file_before$size, input_file_after$size),
    identical(input_file_before$mtime, input_file_after$mtime)
  )
)

if (!all(output_verification$file_exists) ||
    !all(output_verification$rows_expected ==
         output_verification$rows_saved) ||
    !all(output_verification$exact_columns) ||
    !all(output_verification$exact_values) ||
    !all(input_preservation$passed)) {
  print(output_verification)
  print(input_preservation)
  stop("Output or input-preservation verification failed.")
}
```

```{r display-output-verification}
display_table(output_verification)
display_table(input_preservation)
```

# 9. Run summary

```{r run-summary}
cat(
  "- Unexpected share_basis categories: ",
  nrow(unexpected_share_basis), ".\n",
  "- Full allocation scope: ",
  nrow(allocation_pairs_full), " distinct country + pair_id channels; ",
  sum(allocation_pairs_full$commitment_stated == 1L),
  " with commitment_stated = 1 and ",
  sum(allocation_pairs_full$commitment_stated == 0L),
  " with commitment_stated = 0.\n",
  "- Partial/exclusive scope: ",
  nrow(allocation_pairs_partial_exclusive),
  " distinct country + pair_id channels; ",
  sum(allocation_pairs_partial_exclusive$commitment_stated == 1L),
  " with commitment_stated = 1 and ",
  sum(allocation_pairs_partial_exclusive$commitment_stated == 0L),
  " with commitment_stated = 0.\n",
  "- Pair-level conflicts: ",
  nrow(pair_commitment_conflicts), ".\n",
  "- Country shares and counts reconcile: ",
  all(unlist(sanity_checks[fatal_check_columns])), ".\n",
  "- Input allocation file unchanged: ",
  all(input_preservation$passed), ".\n",
  sep = ""
)
```
