---
title: "Revenue sources underlying statutory allocation links"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook attaches the requested revenue classification to the analytical
Census and expanded allocation-link tables. It then describes the composition
of statutory allocation links by revenue class. The database is a **de jure**
legal inventory; the counts and shares below do not measure revenue amounts,
collections, transfers, disbursements, or expenditure execution.

```{r setup, include=FALSE}
if (.Platform$OS.type == "windows") {
  active_utf8_locale <- Sys.setlocale(
    "LC_CTYPE",
    "English_United States.utf8"
  )
  if (identical(active_utf8_locale, "")) {
    stop("A UTF-8 locale is required to preserve accented legal text.")
  }
}

input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(readr)
library(readxl)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

input_paths <- c(
  census = file.path(
    project_root, "data", "derived", "SSA_CENSUS.csv"
  ),
  expanded = file.path(
    project_root,
    "data",
    "derived",
    "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  ),
  mapping = file.path(
    project_root, "data", "revenue_class_mapping.xlsx"
  )
)

output_paths <- c(
  census = input_paths[["census"]],
  expanded = input_paths[["expanded"]],
  summary = file.path(
    project_root,
    "analysis",
    "tables",
    "general",
    "08_revenue_class_summary.csv"
  ),
  all_plot_data = file.path(
    project_root,
    "analysis",
    "tables",
    "general",
    "08_revenue_class_all.csv"
  ),
  earmarked_plot_data = file.path(
    project_root,
    "analysis",
    "tables",
    "partial_exclusive",
    "08_revenue_class_partial_exclusive.csv"
  ),
  all_figure = file.path(
    project_root,
    "analysis",
    "figures",
    "general",
    "08_revenue_class_all.png"
  ),
  earmarked_figure = file.path(
    project_root,
    "analysis",
    "figures",
    "partial_exclusive",
    "08_revenue_class_partial_exclusive.png"
  )
)

for (directory in unique(dirname(output_paths))) {
  dir.create(directory, recursive = TRUE, showWarnings = FALSE)
}

requested_classification_columns <- c(
  "revenue_class",
  "revenue_subclass"
)
partial_exclusive_statuses <- c(
  "partial_earmark",
  "exclusive_earmark"
)
main_blue <- "#1F4E79"

read_all_character <- function(path) {
  read_csv(
    path,
    col_types = cols(.default = col_character()),
    na = "",
    trim_ws = FALSE,
    name_repair = "minimal",
    show_col_types = FALSE,
    progress = FALSE
  )
}

display_table <- function(data, digits = 6) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    print(knitr::kable(data, digits = digits, row.names = FALSE))
    cat("\n")
  }
}

is_missing_text <- function(values) {
  is.na(values) | str_trim(as.character(values)) == ""
}

remove_previous_classification <- function(data, dataset_name) {
  positions <- match(requested_classification_columns, names(data))
  present <- !is.na(positions)

  if (any(present) && !all(present)) {
    stop(
      dataset_name,
      " contains only one of revenue_class and revenue_subclass."
    )
  }

  if (all(present)) {
    expected_positions <- (ncol(data) - 1L):ncol(data)
    if (!identical(as.integer(positions), as.integer(expected_positions))) {
      stop(
        dataset_name,
        " has prior revenue classifications outside the final two columns."
      )
    }
    return(data %>% select(-all_of(requested_classification_columns)))
  }

  data
}

same_existing_values_and_order <- function(before, after) {
  old_columns <- names(before)
  identical(names(after)[seq_along(old_columns)], old_columns) &&
    nrow(before) == nrow(after) &&
    all(vapply(
      old_columns,
      function(column) identical(before[[column]], after[[column]]),
      logical(1)
    ))
}

missing_by_column <- function(data) {
  vapply(data, function(column) sum(is.na(column)), integer(1))
}

duplicate_row_count <- function(data) {
  sum(duplicated(as.data.frame(data, stringsAsFactors = FALSE)))
}

make_preservation_checks <- function(dataset_name, before_data, after_data) {
  old_columns <- names(before_data)
  old_values_same <- same_existing_values_and_order(before_data, after_data)
  old_missing_same <- identical(
    missing_by_column(before_data),
    missing_by_column(after_data %>% select(all_of(old_columns)))
  )

  tibble(
    dataset = dataset_name,
    check = c(
      "rows before = rows after",
      "existing columns unchanged and in the same order",
      "existing values and row order unchanged",
      "existing-column missingness unchanged",
      "no duplicate rows introduced in existing fields",
      "output is original columns plus two appended classifications"
    ),
    before = c(
      as.character(nrow(before_data)),
      paste0(length(old_columns), " existing columns"),
      "existing field vectors",
      "per-column missing counts",
      as.character(duplicate_row_count(before_data)),
      paste(old_columns, collapse = "; ")
    ),
    after = c(
      as.character(nrow(after_data)),
      paste0(length(old_columns), " existing columns"),
      if_else(old_values_same, "identical", "changed"),
      if_else(old_missing_same, "identical", "changed"),
      as.character(duplicate_row_count(
        after_data %>% select(all_of(old_columns))
      )),
      paste(names(after_data), collapse = "; ")
    ),
    passed = c(
      nrow(before_data) == nrow(after_data),
      identical(names(after_data)[seq_along(old_columns)], old_columns),
      old_values_same,
      old_missing_same,
      duplicate_row_count(before_data) == duplicate_row_count(
        after_data %>% select(all_of(old_columns))
      ),
      identical(
        names(after_data),
        c(old_columns, requested_classification_columns)
      )
    )
  )
}

theme_revenue_class <- theme_minimal(base_size = 14) +
  theme(
    plot.title.position = "plot",
    plot.title = element_text(
      face = "bold",
      colour = "#222222",
      size = 18
    ),
    plot.subtitle = element_text(
      colour = "#4D4D4D",
      size = 13,
      margin = margin(b = 10)
    ),
    plot.caption = element_text(
      colour = "#666666",
      hjust = 0,
      size = 10,
      margin = margin(t = 10)
    ),
    axis.title = element_text(
      face = "bold",
      colour = "#333333",
      size = 13
    ),
    axis.text.x = element_text(colour = "#333333", size = 12),
    axis.text.y = element_text(colour = "#333333", size = 12),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    plot.margin = margin(12, 24, 12, 12)
  )
```

# 1. Load and validate inputs

The two CSV files are read entirely as character data so their existing
identifiers, legal text, missing cells, and row order can be preserved exactly.

```{r load-inputs}
if (!all(file.exists(input_paths))) {
  stop(
    "Missing required input file(s): ",
    paste(input_paths[!file.exists(input_paths)], collapse = "; ")
  )
}

census_loaded <- read_all_character(input_paths[["census"]])
expanded_loaded <- read_all_character(input_paths[["expanded"]])

SSA_CENSUS_before <- remove_previous_classification(
  census_loaded,
  "SSA_CENSUS"
)
SSA_EARMARK_ALLOCATIONS_EXPANDED_before <- remove_previous_classification(
  expanded_loaded,
  "SSA_EARMARK_ALLOCATIONS_EXPANDED"
)

required_census_columns <- c(
  "country",
  "evidence_id",
  "instrument_id",
  "instrument_nature"
)
required_expanded_columns <- c(
  "country",
  "pair_id",
  "census_evidence_id",
  "census_instrument_id",
  "earmark_status"
)

missing_census_columns <- setdiff(
  required_census_columns,
  names(SSA_CENSUS_before)
)
missing_expanded_columns <- setdiff(
  required_expanded_columns,
  names(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
)

input_inventory <- tibble(
  dataset = c("SSA_CENSUS", "SSA_EARMARK_ALLOCATIONS_EXPANDED"),
  n_rows = c(
    nrow(SSA_CENSUS_before),
    nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
  ),
  n_columns_before_classification = c(
    ncol(SSA_CENSUS_before),
    ncol(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)
  ),
  missing_required_columns = c(
    if_else(
      length(missing_census_columns) == 0L,
      NA_character_,
      paste(missing_census_columns, collapse = "; ")
    ),
    if_else(
      length(missing_expanded_columns) == 0L,
      NA_character_,
      paste(missing_expanded_columns, collapse = "; ")
    )
  )
)

if (length(missing_census_columns) > 0L ||
    length(missing_expanded_columns) > 0L) {
  print(input_inventory)
  stop("One or more required input columns are missing.")
}
```

```{r display-input-inventory}
display_table(input_inventory)
```

# 2. Inspect and validate the revenue mapping

The workbook is expected to provide one exact mapping for each stored
`instrument_nature`. Matching is exact; no fuzzy, translated, or manually
inferred classifications are introduced.

```{r load-and-audit-mapping}
mapping_sheets <- excel_sheets(input_paths[["mapping"]])
if (length(mapping_sheets) != 1L) {
  stop(
    "The revenue mapping workbook must contain exactly one sheet; found: ",
    paste(mapping_sheets, collapse = "; ")
  )
}

revenue_mapping_loaded <- read_excel(
  input_paths[["mapping"]],
  sheet = mapping_sheets[[1]],
  col_types = "text",
  .name_repair = "minimal"
)

required_mapping_columns <- c(
  "instrument_nature",
  "revenue_class",
  "revenue_subclass"
)
missing_mapping_columns <- setdiff(
  required_mapping_columns,
  names(revenue_mapping_loaded)
)
additional_mapping_columns <- setdiff(
  names(revenue_mapping_loaded),
  required_mapping_columns
)

mapping_schema_audit <- tibble(
  workbook = "data/revenue_class_mapping.xlsx",
  sheet = mapping_sheets[[1]],
  n_rows = nrow(revenue_mapping_loaded),
  n_columns = ncol(revenue_mapping_loaded),
  missing_columns = if_else(
    length(missing_mapping_columns) == 0L,
    NA_character_,
    paste(missing_mapping_columns, collapse = "; ")
  ),
  additional_columns = if_else(
    length(additional_mapping_columns) == 0L,
    NA_character_,
    paste(additional_mapping_columns, collapse = "; ")
  )
)

if (length(missing_mapping_columns) > 0L) {
  print(mapping_schema_audit)
  stop("The mapping workbook is missing required columns.")
}

mapping_missing_fields <- revenue_mapping_loaded %>%
  transmute(
    instrument_nature,
    revenue_class,
    revenue_subclass,
    missing_instrument_nature = is_missing_text(instrument_nature),
    missing_revenue_class = is_missing_text(revenue_class),
    missing_revenue_subclass = is_missing_text(revenue_subclass)
  ) %>%
  filter(
    missing_instrument_nature |
      missing_revenue_class |
      missing_revenue_subclass
  )

mapping_duplicate_keys <- revenue_mapping_loaded %>%
  filter(!is_missing_text(instrument_nature)) %>%
  count(instrument_nature, name = "n_mapping_rows") %>%
  filter(n_mapping_rows > 1L) %>%
  arrange(desc(n_mapping_rows), instrument_nature)

mapping_conflicts <- revenue_mapping_loaded %>%
  filter(!is_missing_text(instrument_nature)) %>%
  distinct(instrument_nature, revenue_class, revenue_subclass) %>%
  count(instrument_nature, name = "n_distinct_mappings") %>%
  filter(n_distinct_mappings > 1L) %>%
  arrange(desc(n_distinct_mappings), instrument_nature)

if (any(is_missing_text(revenue_mapping_loaded$instrument_nature)) ||
    nrow(mapping_duplicate_keys) > 0L ||
    nrow(mapping_conflicts) > 0L) {
  print(mapping_missing_fields)
  print(mapping_duplicate_keys)
  print(mapping_conflicts)
  stop(
    "The mapping key is missing, duplicated, or ambiguous. ",
    "No derived file was changed."
  )
}

revenue_mapping <- revenue_mapping_loaded %>%
  select(all_of(required_mapping_columns))
```

```{r display-mapping-audit}
display_table(mapping_schema_audit)
display_table(mapping_missing_fields)
display_table(mapping_duplicate_keys)
display_table(mapping_conflicts)
```

# 3. Append revenue classifications to Census

Unmatched `instrument_nature` values are reported and left unclassified. The
original Census fields and row order remain unchanged.

```{r audit-unmatched-instrument-natures}
unmatched_instrument_nature <- SSA_CENSUS_before %>%
  anti_join(
    revenue_mapping %>% select(instrument_nature),
    by = "instrument_nature"
  ) %>%
  mutate(
    instrument_nature_display = if_else(
      is_missing_text(instrument_nature),
      "Missing",
      instrument_nature
    )
  ) %>%
  count(
    country,
    instrument_nature = instrument_nature_display,
    name = "n_census_rows"
  ) %>%
  arrange(country, instrument_nature)

instrument_nature_mapping_audit <- tibble(
  check = c(
    "Census rows",
    "distinct non-missing Census instrument_nature values",
    "matched Census rows",
    "unmatched Census rows",
    "unmatched distinct instrument_nature values"
  ),
  value = c(
    nrow(SSA_CENSUS_before),
    n_distinct(
      SSA_CENSUS_before$instrument_nature[
        !is_missing_text(SSA_CENSUS_before$instrument_nature)
      ]
    ),
    nrow(SSA_CENSUS_before) - sum(unmatched_instrument_nature$n_census_rows),
    sum(unmatched_instrument_nature$n_census_rows),
    nrow(unmatched_instrument_nature)
  )
)
```

```{r display-unmatched-instrument-natures}
display_table(instrument_nature_mapping_audit)
display_table(unmatched_instrument_nature)
```

```{r classify-census}
SSA_CENSUS <- SSA_CENSUS_before %>%
  mutate(.original_row_order = row_number()) %>%
  left_join(
    revenue_mapping,
    by = "instrument_nature",
    relationship = "many-to-one"
  ) %>%
  arrange(.original_row_order) %>%
  select(
    all_of(names(SSA_CENSUS_before)),
    all_of(requested_classification_columns)
  )

census_prewrite_checks <- make_preservation_checks(
  "SSA_CENSUS",
  SSA_CENSUS_before,
  SSA_CENSUS
)

if (!all(census_prewrite_checks$passed)) {
  print(census_prewrite_checks %>% filter(!passed))
  stop("Census preservation checks failed; no output was written.")
}
```

# 4. Append Census classifications to expanded allocation links

The join uses `country + census_evidence_id = evidence_id`. A matched Census
row must also carry the same instrument identifier as the expanded link.

```{r validate-census-join-key}
census_lookup_key_audit <- SSA_CENSUS %>%
  summarise(
    n_rows = n(),
    missing_country = sum(is_missing_text(country)),
    missing_evidence_id = sum(is_missing_text(evidence_id)),
    n_duplicate_country_evidence_keys = nrow(
      SSA_CENSUS %>%
        count(country, evidence_id, name = "n_rows") %>%
        filter(n_rows > 1L)
    )
  )

duplicate_census_lookup_keys <- SSA_CENSUS %>%
  count(country, evidence_id, name = "n_rows") %>%
  filter(n_rows > 1L) %>%
  arrange(desc(n_rows), country, evidence_id)

if (census_lookup_key_audit$missing_country > 0L ||
    census_lookup_key_audit$missing_evidence_id > 0L ||
    nrow(duplicate_census_lookup_keys) > 0L) {
  print(census_lookup_key_audit)
  print(duplicate_census_lookup_keys)
  stop(
    "The Census join key is incomplete or ambiguous. ",
    "No expanded file was changed."
  )
}
```

```{r join-expanded-classification}
census_classification_lookup <- SSA_CENSUS %>%
  transmute(
    country,
    census_evidence_id = evidence_id,
    census_instrument_id_from_census = instrument_id,
    revenue_class,
    revenue_subclass,
    census_match_found = TRUE
  )

expanded_joined <- SSA_EARMARK_ALLOCATIONS_EXPANDED_before %>%
  mutate(.original_row_order = row_number()) %>%
  left_join(
    census_classification_lookup,
    by = c("country", "census_evidence_id"),
    relationship = "many-to-one"
  ) %>%
  arrange(.original_row_order)

instrument_link_mismatches <- expanded_joined %>%
  filter(
    !is.na(census_match_found),
    census_instrument_id != census_instrument_id_from_census
  ) %>%
  select(
    country,
    pair_id,
    census_evidence_id,
    expanded_census_instrument_id = census_instrument_id,
    census_instrument_id = census_instrument_id_from_census
  )

if (nrow(instrument_link_mismatches) > 0L) {
  print(instrument_link_mismatches)
  stop(
    "Matched Census evidence IDs disagree with expanded census instrument IDs. ",
    "No output was written."
  )
}

unmatched_census_joins <- expanded_joined %>%
  filter(is.na(census_match_found)) %>%
  count(country, name = "n_unmatched_expanded_rows") %>%
  left_join(
    expanded_joined %>%
      filter(is.na(census_match_found)) %>%
      distinct(country, census_evidence_id) %>%
      count(country, name = "n_distinct_unmatched_census_evidence_ids"),
    by = "country"
  ) %>%
  arrange(desc(n_unmatched_expanded_rows), country)

join_audit <- tibble(
  check = c(
    "expanded rows before join",
    "expanded rows after join",
    "matched Census rows",
    "unmatched Census joins",
    "ambiguous Census joins",
    "matched instrument-ID disagreements"
  ),
  value = c(
    nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED_before),
    nrow(expanded_joined),
    sum(!is.na(expanded_joined$census_match_found)),
    sum(is.na(expanded_joined$census_match_found)),
    nrow(duplicate_census_lookup_keys),
    nrow(instrument_link_mismatches)
  )
)

SSA_EARMARK_ALLOCATIONS_EXPANDED <- expanded_joined %>%
  select(
    all_of(names(SSA_EARMARK_ALLOCATIONS_EXPANDED_before)),
    all_of(requested_classification_columns)
  )

expanded_prewrite_checks <- make_preservation_checks(
  "SSA_EARMARK_ALLOCATIONS_EXPANDED",
  SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
  SSA_EARMARK_ALLOCATIONS_EXPANDED
)

if (!all(expanded_prewrite_checks$passed)) {
  print(expanded_prewrite_checks %>% filter(!passed))
  stop("Expanded-table preservation checks failed; no output was written.")
}
```

```{r display-expanded-join-audit}
display_table(census_lookup_key_audit)
display_table(duplicate_census_lookup_keys)
display_table(join_audit)
display_table(unmatched_census_joins)
display_table(instrument_link_mismatches)
```

# 5. Verify the atomic allocation-link unit

One atomic allocation link is one distinct
`country + pair_id + census_instrument_id` combination. Counts are constructed
only after confirming this key is complete and unique.

```{r audit-atomic-link-key}
atomic_key_missing_rows <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  filter(
    is_missing_text(country) |
      is_missing_text(pair_id) |
      is_missing_text(census_instrument_id)
  ) %>%
  select(
    country,
    pair_id,
    census_instrument_id,
    census_evidence_id
  )

duplicate_atomic_link_keys <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  count(
    country,
    pair_id,
    census_instrument_id,
    name = "n_rows"
  ) %>%
  filter(n_rows > 1L) %>%
  arrange(desc(n_rows), country, pair_id, census_instrument_id)

atomic_key_audit <- tibble(
  check = c(
    "expanded rows",
    "distinct complete atomic-link keys",
    "rows with an incomplete atomic-link key",
    "duplicate atomic-link keys",
    "duplicate-row excess over the atomic key"
  ),
  value = c(
    nrow(SSA_EARMARK_ALLOCATIONS_EXPANDED),
    nrow(
      SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
        filter(
          !is_missing_text(country),
          !is_missing_text(pair_id),
          !is_missing_text(census_instrument_id)
        ) %>%
        distinct(country, pair_id, census_instrument_id)
    ),
    nrow(atomic_key_missing_rows),
    nrow(duplicate_atomic_link_keys),
    sum(pmax(duplicate_atomic_link_keys$n_rows - 1L, 0L))
  )
)

if (nrow(atomic_key_missing_rows) > 0L ||
    nrow(duplicate_atomic_link_keys) > 0L) {
  print(atomic_key_audit)
  print(atomic_key_missing_rows)
  print(duplicate_atomic_link_keys)
  stop(
    "The atomic allocation-link key is incomplete or duplicated. ",
    "No summary or output was created."
  )
}

atomic_links_all <- SSA_EARMARK_ALLOCATIONS_EXPANDED %>%
  distinct(country, pair_id, census_instrument_id, .keep_all = TRUE) %>%
  mutate(
    revenue_class_display = if_else(
      is_missing_text(revenue_class),
      "Missing",
      revenue_class
    )
  )

atomic_links_earmarked <- atomic_links_all %>%
  filter(earmark_status %in% partial_exclusive_statuses)

earmark_status_filter_audit <- tibble(
  sample = c(
    "All statutory allocations",
    "Partial and exclusive earmarks"
  ),
  n_atomic_links = c(
    nrow(atomic_links_all),
    nrow(atomic_links_earmarked)
  ),
  statuses_present = c(
    paste(
      sort(unique(atomic_links_all$earmark_status)),
      collapse = "; "
    ),
    paste(
      sort(unique(atomic_links_earmarked$earmark_status)),
      collapse = "; "
    )
  )
)
```

```{r display-atomic-link-audit}
display_table(atomic_key_audit)
display_table(atomic_key_missing_rows)
display_table(duplicate_atomic_link_keys)
display_table(earmark_status_filter_audit)
```

# 6. Revenue-class summary

Missing Census matches and unmatched `instrument_nature` values remain visible
as `Missing`. No link is removed from either denominator.

```{r compute-revenue-class-summary}
total_links_all <- nrow(atomic_links_all)
total_links_earmarked <- nrow(atomic_links_earmarked)

if (total_links_all == 0L || total_links_earmarked == 0L) {
  stop("Both the full and filtered samples must contain at least one link.")
}

revenue_class_counts_all <- atomic_links_all %>%
  count(revenue_class = revenue_class_display, name = "n_links_all")

revenue_class_counts_earmarked <- atomic_links_earmarked %>%
  count(
    revenue_class = revenue_class_display,
    name = "n_links_earmarked"
  )

revenue_class_summary <- full_join(
  revenue_class_counts_all,
  revenue_class_counts_earmarked,
  by = "revenue_class"
) %>%
  mutate(
    n_links_all = replace_na(n_links_all, 0L),
    n_links_earmarked = replace_na(n_links_earmarked, 0L),
    share_links_all = n_links_all / total_links_all,
    share_links_earmarked = n_links_earmarked / total_links_earmarked
  ) %>%
  select(
    revenue_class,
    n_links_all,
    share_links_all,
    n_links_earmarked,
    share_links_earmarked
  ) %>%
  arrange(desc(n_links_all), revenue_class)

share_and_total_checks <- tibble(
  sample = c(
    "All statutory allocations",
    "Partial and exclusive earmarks"
  ),
  total_atomic_links = c(total_links_all, total_links_earmarked),
  sum_class_link_counts = c(
    sum(revenue_class_summary$n_links_all),
    sum(revenue_class_summary$n_links_earmarked)
  ),
  sum_class_shares = c(
    sum(revenue_class_summary$share_links_all),
    sum(revenue_class_summary$share_links_earmarked)
  ),
  passed = c(
    sum(revenue_class_summary$n_links_all) == total_links_all &&
      isTRUE(all.equal(
        sum(revenue_class_summary$share_links_all),
        1,
        tolerance = 1e-12
      )),
    sum(revenue_class_summary$n_links_earmarked) == total_links_earmarked &&
      isTRUE(all.equal(
        sum(revenue_class_summary$share_links_earmarked),
        1,
        tolerance = 1e-12
      ))
  )
)

if (!all(share_and_total_checks$passed)) {
  print(share_and_total_checks)
  stop("Revenue-class counts or shares do not reconcile to their sample total.")
}
```

```{r display-revenue-class-summary}
display_table(revenue_class_summary)
display_table(share_and_total_checks)
```

# 7. Exact plot data

The two CSV tables below contain exactly the rows and values displayed in their
matching figures. The readable labels are derived only for presentation; the
stored `revenue_class` codes remain unchanged.

```{r prepare-plot-data}
make_revenue_class_label <- function(values) {
  values %>%
    str_replace_all("_", " ") %>%
    str_to_sentence() %>%
    str_wrap(width = 34)
}

revenue_class_all_plot_data <- revenue_class_summary %>%
  transmute(
    revenue_class,
    revenue_class_label = make_revenue_class_label(revenue_class),
    n_links_all,
    share_links_all
  ) %>%
  arrange(desc(share_links_all), revenue_class)

revenue_class_earmarked_plot_data <- revenue_class_summary %>%
  transmute(
    revenue_class,
    revenue_class_label = make_revenue_class_label(revenue_class),
    n_links_earmarked,
    share_links_earmarked
  ) %>%
  arrange(desc(share_links_earmarked), revenue_class)
```

```{r display-plot-data}
display_table(revenue_class_all_plot_data)
display_table(revenue_class_earmarked_plot_data)
```

# 8. All statutory allocations

```{r build-allocation-figure}
all_plot_frame <- revenue_class_all_plot_data %>%
  mutate(
    revenue_class_label = factor(
      revenue_class_label,
      levels = rev(revenue_class_label)
    )
  )

revenue_class_all_plot <- ggplot(
  all_plot_frame,
  aes(x = share_links_all, y = revenue_class_label)
) +
  geom_col(width = 0.68, fill = main_blue) +
  geom_text(
    aes(label = label_percent(accuracy = 0.1)(share_links_all)),
    hjust = -0.12,
    colour = "#333333",
    size = 4.2
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title = "Which revenue sources underpin statutory allocations?",
    subtitle = paste0(
      "Share of ", comma(total_links_all),
      " source-to-destination links in the observed legal corpora"
    ),
    x = "Share of statutory allocation links",
    y = NULL,
    caption = str_wrap(
      paste0(
        "Note: One link is one country × allocation channel × linked Census ",
        "instrument. Missing includes unmatched classifications and links ",
        "without a match in the analytical Census."
      ),
      width = 105
    )
  ) +
  theme_revenue_class
```

```{r display-allocation-figure, echo=FALSE}
revenue_class_all_plot
```

# 9. Partial and exclusive earmarks

```{r build-earmarked-figure}
earmarked_plot_frame <- revenue_class_earmarked_plot_data %>%
  mutate(
    revenue_class_label = factor(
      revenue_class_label,
      levels = rev(revenue_class_label)
    )
  )

revenue_class_earmarked_plot <- ggplot(
  earmarked_plot_frame,
  aes(x = share_links_earmarked, y = revenue_class_label)
) +
  geom_col(width = 0.68, fill = main_blue) +
  geom_text(
    aes(label = label_percent(accuracy = 0.1)(share_links_earmarked)),
    hjust = -0.12,
    colour = "#333333",
    size = 4.2
  ) +
  scale_x_continuous(
    labels = label_percent(accuracy = 1),
    limits = c(0, NA),
    expand = expansion(mult = c(0, 0.18))
  ) +
  labs(
    title = "Which revenue sources underpin functional earmarks?",
    subtitle = paste0(
      "Share of ", comma(total_links_earmarked),
      " links from partial and exclusive earmark instruments"
    ),
    x = "Share of earmark links",
    y = NULL,
    caption = str_wrap(
      paste0(
        "Note: One link is one country × allocation channel × linked Census ",
        "instrument. The classification does not measure the share of ",
        "revenue amounts that is earmarked."
      ),
      width = 105
    )
  ) +
  theme_revenue_class
```

```{r display-earmarked-figure, echo=FALSE}
revenue_class_earmarked_plot
```

# 10. Stage, verify, and commit all outputs

All candidate files are written to temporary paths and checked before any final
path is overwritten. Existing final files are backed up and restored if a
commit or final read-back check fails.

```{r stage-and-verify-outputs}
stage_dir <- tempfile(pattern = "08_revenue_class_stage_")
dir.create(stage_dir, recursive = TRUE, showWarnings = FALSE)

stage_paths <- file.path(stage_dir, basename(output_paths))
names(stage_paths) <- names(output_paths)

write_csv(SSA_CENSUS, stage_paths[["census"]], na = "")
write_csv(
  SSA_EARMARK_ALLOCATIONS_EXPANDED,
  stage_paths[["expanded"]],
  na = ""
)
write_csv(
  revenue_class_summary,
  stage_paths[["summary"]],
  na = ""
)
write_csv(
  revenue_class_all_plot_data,
  stage_paths[["all_plot_data"]],
  na = ""
)
write_csv(
  revenue_class_earmarked_plot_data,
  stage_paths[["earmarked_plot_data"]],
  na = ""
)

ggsave(
  filename = stage_paths[["all_figure"]],
  plot = revenue_class_all_plot,
  width = 10.5,
  height = 7.2,
  units = "in",
  dpi = 300,
  bg = "white"
)
ggsave(
  filename = stage_paths[["earmarked_figure"]],
  plot = revenue_class_earmarked_plot,
  width = 10.5,
  height = 7.2,
  units = "in",
  dpi = 300,
  bg = "white"
)

staged_census <- read_all_character(stage_paths[["census"]])
staged_expanded <- read_all_character(stage_paths[["expanded"]])

staged_preservation_checks <- bind_rows(
  make_preservation_checks(
    "SSA_CENSUS",
    SSA_CENSUS_before,
    staged_census
  ),
  make_preservation_checks(
    "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
    staged_expanded
  )
)

read_analysis_csv <- function(path) {
  read_csv(path, show_col_types = FALSE, progress = FALSE)
}

same_table_values <- function(saved, expected) {
  identical(names(saved), names(expected)) &&
    nrow(saved) == nrow(expected) &&
    isTRUE(all.equal(
      as.data.frame(saved),
      as.data.frame(expected),
      check.attributes = FALSE,
      tolerance = 1e-12
    ))
}

staged_analysis_checks <- tibble(
  output = c(
    "08_revenue_class_summary.csv",
    "08_revenue_class_all.csv",
    "08_revenue_class_partial_exclusive.csv",
    "08_revenue_class_all.png",
    "08_revenue_class_partial_exclusive.png"
  ),
  passed = c(
    same_table_values(
      read_analysis_csv(stage_paths[["summary"]]),
      revenue_class_summary
    ),
    same_table_values(
      read_analysis_csv(stage_paths[["all_plot_data"]]),
      revenue_class_all_plot_data
    ),
    same_table_values(
      read_analysis_csv(stage_paths[["earmarked_plot_data"]]),
      revenue_class_earmarked_plot_data
    ),
    file.info(stage_paths[["all_figure"]])$size > 10000,
    file.info(stage_paths[["earmarked_figure"]])$size > 10000
  )
)

if (!all(staged_preservation_checks$passed) ||
    !all(staged_analysis_checks$passed)) {
  print(staged_preservation_checks %>% filter(!passed))
  print(staged_analysis_checks %>% filter(!passed))
  stop("Staged output checks failed; no final path was overwritten.")
}
```

```{r commit-output-bundle}
commit_output_bundle <- function(stage_paths, output_paths) {
  existed_before <- file.exists(output_paths)
  backup_dir <- tempfile(pattern = "08_revenue_class_backup_")
  dir.create(backup_dir, recursive = TRUE, showWarnings = FALSE)
  backup_paths <- file.path(backup_dir, basename(output_paths))
  names(backup_paths) <- names(output_paths)
  committed <- FALSE

  on.exit({
    if (!committed) {
      for (name in names(output_paths)) {
        if (existed_before[[name]] && file.exists(backup_paths[[name]])) {
          file.copy(
            backup_paths[[name]],
            output_paths[[name]],
            overwrite = TRUE
          )
        } else if (!existed_before[[name]] &&
                   file.exists(output_paths[[name]])) {
          file.remove(output_paths[[name]])
        }
      }
    }
    unlink(stage_dir, recursive = TRUE, force = TRUE)
    unlink(backup_dir, recursive = TRUE, force = TRUE)
  }, add = TRUE)

  for (name in names(output_paths)[existed_before]) {
    if (!file.copy(
      output_paths[[name]],
      backup_paths[[name]],
      overwrite = FALSE
    )) {
      stop("Could not create a rollback copy for ", name, ".")
    }
  }

  commit_ok <- vapply(
    names(output_paths),
    function(name) file.copy(
      stage_paths[[name]],
      output_paths[[name]],
      overwrite = TRUE
    ),
    logical(1)
  )
  if (!all(commit_ok)) {
    stop("Could not commit every requested output file.")
  }

  final_census <- read_all_character(output_paths[["census"]])
  final_expanded <- read_all_character(output_paths[["expanded"]])

  final_preservation_checks <- bind_rows(
    make_preservation_checks(
      "SSA_CENSUS",
      SSA_CENSUS_before,
      final_census
    ),
    make_preservation_checks(
      "SSA_EARMARK_ALLOCATIONS_EXPANDED",
      SSA_EARMARK_ALLOCATIONS_EXPANDED_before,
      final_expanded
    )
  )

  final_analysis_checks <- tibble(
    output = c(
      "08_revenue_class_summary.csv",
      "08_revenue_class_all.csv",
      "08_revenue_class_partial_exclusive.csv",
      "08_revenue_class_all.png",
      "08_revenue_class_partial_exclusive.png"
    ),
    passed = c(
      same_table_values(
        read_analysis_csv(output_paths[["summary"]]),
        revenue_class_summary
      ),
      same_table_values(
        read_analysis_csv(output_paths[["all_plot_data"]]),
        revenue_class_all_plot_data
      ),
      same_table_values(
        read_analysis_csv(output_paths[["earmarked_plot_data"]]),
        revenue_class_earmarked_plot_data
      ),
      file.info(output_paths[["all_figure"]])$size > 10000,
      file.info(output_paths[["earmarked_figure"]])$size > 10000
    )
  )

  if (!all(final_preservation_checks$passed) ||
      !all(final_analysis_checks$passed)) {
    print(final_preservation_checks %>% filter(!passed))
    print(final_analysis_checks %>% filter(!passed))
    stop("Final read-back checks failed; restoring prior files.")
  }

  committed <- TRUE
  list(
    census = final_census,
    expanded = final_expanded,
    preservation_checks = final_preservation_checks,
    analysis_checks = final_analysis_checks
  )
}

committed_outputs <- commit_output_bundle(stage_paths, output_paths)
```

```{r display-final-output-checks}
display_table(committed_outputs$preservation_checks)
display_table(committed_outputs$analysis_checks)
```

# 11. Run summary

```{r run-summary, results='asis'}
cat(
  "- Census rows preserved: ", nrow(committed_outputs$census), ".\n",
  "- Expanded allocation-link rows preserved: ",
  nrow(committed_outputs$expanded), ".\n",
  "- Atomic links: ", comma(total_links_all),
  " in the full sample and ", comma(total_links_earmarked),
  " among partial and exclusive earmarks.\n",
  "- Unmatched Census instrument natures: ",
  nrow(unmatched_instrument_nature), " values across ",
  sum(unmatched_instrument_nature$n_census_rows), " Census rows.\n",
  "- Unmatched expanded-to-Census joins: ",
  sum(is.na(expanded_joined$census_match_found)), " rows.\n",
  "- Duplicate atomic-link keys: ", nrow(duplicate_atomic_link_keys), ".\n",
  "- Revenue-class shares sum to ",
  percent(sum(revenue_class_summary$share_links_all), accuracy = 0.1),
  " in the full sample and ",
  percent(
    sum(revenue_class_summary$share_links_earmarked),
    accuracy = 0.1
  ),
  " in the filtered sample.\n",
  sep = ""
)
```
