---
title: "Revenue sources and statutory allocation destinations across countries"
author: "IMF research project"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    df_print: paged
---

This notebook creates one source-to-destination Sankey plot for every country
with eligible partial or exclusive earmark links. The database is a **de jure**
legal inventory; the figures do not measure collections, transfers,
disbursements, or expenditure execution.

```{r setup, include=FALSE}
if (.Platform$OS.type == "windows") {
  active_utf8_locale <- Sys.setlocale(
    "LC_CTYPE",
    "English_United States.utf8"
  )
  if (identical(active_utf8_locale, "")) {
    stop("A UTF-8 locale is required to preserve accented text.")
  }
}

input_path <- knitr::current_input(dir = TRUE)
project_root <- normalizePath(
  file.path(dirname(input_path), "..", ".."),
  winslash = "/",
  mustWork = TRUE
)

knitr::opts_knit$set(root.dir = project_root)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggalluvial)
library(knitr)

options(stringsAsFactors = FALSE, scipen = 999)

input_paths <- c(
  expanded = file.path(
    project_root,
    "data",
    "derived",
    "SSA_EARMARK_ALLOCATIONS_EXPANDED.csv"
  ),
  cofog_mapping = file.path(
    project_root,
    "data",
    "COFOG_GFSM2014_mapping.csv"
  )
)

table_dir <- file.path(
  project_root,
  "analysis",
  "tables",
  "partial_exclusive",
  "country_focus"
)
figure_dir <- file.path(
  project_root,
  "analysis",
  "figures",
  "partial_exclusive",
  "country_focus"
)
dir.create(table_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figure_dir, recursive = TRUE, showWarnings = FALSE)

partial_exclusive_statuses <- c(
  "partial_earmark",
  "exclusive_earmark"
)
excluded_revenue_classes <- c(
  "Missing",
  "unclassified_review"
)
excluded_destinations <- c(
  "local_budget",
  "unallocated",
  "general_budget"
)

# This lookup changes labels only; country inclusion is always data-driven.
country_names <- tribble(
  ~country, ~country_name,
  "BEN", "Benin",
  "BFA", "Burkina Faso",
  "CAF", "Central African Republic",
  "CIV", "C\u00f4te d'Ivoire",
  "CMR", "Cameroon",
  "COD", "Democratic Republic of the Congo",
  "COG", "Republic of the Congo",
  "COM", "Comoros",
  "GAB", "Gabon",
  "GHA", "Ghana",
  "GIN", "Guinea",
  "MDG", "Madagascar",
  "MLI", "Mali",
  "NER", "Niger",
  "NGA", "Nigeria",
  "SEN", "Senegal",
  "TCD", "Chad",
  "TGO", "Togo",
  "ZAF", "South Africa"
)

revenue_class_colours <- c(
  "administrative_regulatory_and_user_charges" = "#4C6F91",
  "environmental_and_pollution" = "#557A59",
  "excises_and_specific_consumption" = "#8A763D",
  "financial_and_transaction_taxes" = "#4E7C7D",
  "general_consumption_and_turnover" = "#466B70",
  "income_profits_and_capital_gains" = "#5A6992",
  "international_trade_and_customs" = "#7E6753",
  "natural_resources_and_extractives" = "#2F6F73",
  "payroll_and_social_contributions" = "#8B6F47",
  "penalties_interest_and_other_receipts" = "#887379",
  "property_and_wealth" = "#6B6B6B",
  "registration_stamp_and_legal_duties" = "#7A4F68",
  "sector_specific_and_business_levies" = "#1F4E79"
)

read_all_character <- function(path) {
  read_csv(
    path,
    col_types = cols(.default = col_character()),
    na = "",
    trim_ws = FALSE,
    name_repair = "minimal",
    show_col_types = FALSE,
    progress = FALSE
  )
}

display_table <- function(data, digits = 6) {
  if (nrow(data) == 0L) {
    cat("No rows to report.\n")
  } else {
    print(knitr::kable(data, digits = digits, row.names = FALSE))
    cat("\n")
  }
}

is_missing_text <- function(values) {
  is.na(values) | str_trim(as.character(values)) == ""
}

make_revenue_class_label <- function(values) {
  values %>%
    str_replace_all("_", " ") %>%
    str_to_sentence()
}

make_destination_label <- function(code, name) {
  if_else(
    str_detect(code, "^[0-9]+$"),
    paste0(code, " \u2014 ", name),
    name
  )
}

destination_order_key <- function(code) {
  case_when(
    str_detect(code, "^[0-9]+$") ~ as.numeric(code),
    code == "Missing" ~ 1001,
    TRUE ~ 2000
  )
}
```

# 1. Load and validate the inputs

The expanded allocation table is the analytical input. The COFOG mapping is
used only to attach readable names to the destination codes already present in
that table.

```{r load-and-validate-inputs}
if (!all(file.exists(input_paths))) {
  stop(
    "Missing input files: ",
    paste(names(input_paths)[!file.exists(input_paths)], collapse = "; ")
  )
}

input_md5_before <- unname(tools::md5sum(input_paths))

expanded_loaded <- read_all_character(input_paths[["expanded"]])
cofog_mapping_loaded <- read_delim(
  input_paths[["cofog_mapping"]],
  delim = ";",
  col_types = cols(.default = col_character()),
  trim_ws = TRUE,
  name_repair = "minimal",
  show_col_types = FALSE,
  progress = FALSE
)

required_expanded_columns <- c(
  "country",
  "pair_id",
  "census_instrument_id",
  "earmark_status",
  "destination_function",
  "destination_function_detail",
  "revenue_class"
)
required_mapping_columns <- c(
  "cofog_code",
  "level",
  "cofog_name"
)

missing_expanded_columns <- setdiff(
  required_expanded_columns,
  names(expanded_loaded)
)
missing_mapping_columns <- setdiff(
  required_mapping_columns,
  names(cofog_mapping_loaded)
)

schema_audit <- tibble(
  input = c(
    "SSA_EARMARK_ALLOCATIONS_EXPANDED",
    "COFOG_GFSM2014_mapping"
  ),
  rows = c(nrow(expanded_loaded), nrow(cofog_mapping_loaded)),
  columns = c(ncol(expanded_loaded), ncol(cofog_mapping_loaded)),
  missing_required_columns = c(
    if_else(
      length(missing_expanded_columns) == 0L,
      NA_character_,
      paste(missing_expanded_columns, collapse = "; ")
    ),
    if_else(
      length(missing_mapping_columns) == 0L,
      NA_character_,
      paste(missing_mapping_columns, collapse = "; ")
    )
  )
)

if (length(missing_expanded_columns) > 0L ||
    length(missing_mapping_columns) > 0L) {
  print(schema_audit)
  stop("One or more required columns are missing.")
}

missing_country_names <- setdiff(
  sort(unique(expanded_loaded$country)),
  country_names$country
)
if (length(missing_country_names) > 0L) {
  stop(
    "Country-name labels are missing for: ",
    paste(missing_country_names, collapse = "; ")
  )
}
```

```{r display-schema-audit}
display_table(schema_audit)
```

# 2. Apply the earmark-status and revenue-class filters

Blank revenue classifications are made visible as `Missing` before applying
the requested revenue-class exclusion. No other revenue classification is
changed.

```{r filter-status-and-revenue-class}
expanded_prepared <- expanded_loaded %>%
  mutate(
    country = str_trim(country),
    revenue_class_display = if_else(
      is_missing_text(revenue_class),
      "Missing",
      str_trim(revenue_class)
    )
  )

status_eligible <- expanded_prepared %>%
  filter(earmark_status %in% partial_exclusive_statuses)

revenue_eligible <- status_eligible %>%
  filter(!revenue_class_display %in% excluded_revenue_classes)

missing_status_keys <- status_eligible %>%
  filter(
    is_missing_text(country) |
      is_missing_text(pair_id) |
      is_missing_text(census_instrument_id)
  ) %>%
  select(country, pair_id, census_instrument_id)

if (nrow(missing_status_keys) > 0L) {
  print(missing_status_keys)
  stop("The allocation-link key is incomplete after the status filter.")
}
if (!all(status_eligible$earmark_status %in% partial_exclusive_statuses)) {
  stop("A non-partial/non-exclusive status remains after filtering.")
}

revenue_class_exclusion_audit <- status_eligible %>%
  mutate(
    exclusion = case_when(
      revenue_class_display == "Missing" ~ "Missing",
      revenue_class_display == "unclassified_review" ~
        "unclassified_review",
      TRUE ~ "Retained"
    )
  ) %>%
  count(country, exclusion, name = "n_expanded_rows") %>%
  complete(
    country = sort(unique(expanded_prepared$country)),
    exclusion = c("Missing", "unclassified_review", "Retained"),
    fill = list(n_expanded_rows = 0L)
  )

retained_revenue_classes <- sort(unique(
  revenue_eligible$revenue_class_display
))
unexpected_revenue_classes <- setdiff(
  retained_revenue_classes,
  names(revenue_class_colours)
)
if (length(unexpected_revenue_classes) > 0L) {
  revenue_class_colours <- c(
    revenue_class_colours,
    set_names(
      rep("#6B6B6B", length(unexpected_revenue_classes)),
      unexpected_revenue_classes
    )
  )
}
```

```{r display-revenue-filter-audit}
display_table(
  revenue_class_exclusion_audit %>%
    filter(exclusion != "Retained", n_expanded_rows > 0L)
)
if (length(unexpected_revenue_classes) == 0L) {
  cat("No retained revenue classes lack a specified plot colour.\n")
} else {
  cat(
    "Retained revenue classes using the documented neutral fallback colour: ",
    paste(unexpected_revenue_classes, collapse = "; "),
    "\n"
  )
}
```

# 3. Expand and deduplicate destinations

The primary destination is combined with every semicolon-separated detail
destination. Empty tokens are removed. A row with no destination in either
field is retained as `Missing`. The final counting unit is exactly one distinct
`country + pair_id + census_instrument_id + destination_function` link.

```{r expand-and-deduplicate-destinations}
primary_destination_candidates <- revenue_eligible %>%
  transmute(
    country,
    pair_id,
    census_instrument_id,
    revenue_class = revenue_class_display,
    destination_function = str_trim(destination_function),
    token_source = "primary"
  ) %>%
  filter(!is_missing_text(destination_function))

detail_destination_candidates <- revenue_eligible %>%
  select(
    country,
    pair_id,
    census_instrument_id,
    revenue_class = revenue_class_display,
    destination_function_detail
  ) %>%
  separate_longer_delim(destination_function_detail, delim = ";") %>%
  transmute(
    country,
    pair_id,
    census_instrument_id,
    revenue_class,
    destination_function = str_trim(destination_function_detail),
    token_source = "detail"
  ) %>%
  filter(!is_missing_text(destination_function))

nonmissing_destination_candidates <- bind_rows(
  primary_destination_candidates,
  detail_destination_candidates
)

links_without_destination <- revenue_eligible %>%
  distinct(
    country,
    pair_id,
    census_instrument_id,
    revenue_class = revenue_class_display
  ) %>%
  anti_join(
    nonmissing_destination_candidates %>%
      distinct(country, pair_id, census_instrument_id),
    by = c("country", "pair_id", "census_instrument_id")
  ) %>%
  mutate(
    destination_function = "Missing",
    token_source = "missing"
  )

destination_candidates <- bind_rows(
  nonmissing_destination_candidates,
  links_without_destination
)

candidate_key_audit <- destination_candidates %>%
  group_by(
    country,
    pair_id,
    census_instrument_id,
    destination_function
  ) %>%
  summarise(
    n_candidate_rows = n(),
    n_revenue_classes = n_distinct(revenue_class),
    token_sources = paste(sort(unique(token_source)), collapse = "; "),
    .groups = "drop"
  )

conflicting_link_sources <- candidate_key_audit %>%
  filter(n_revenue_classes > 1L)

duplicate_link_candidates <- candidate_key_audit %>%
  filter(n_candidate_rows > 1L) %>%
  arrange(
    country,
    desc(n_candidate_rows),
    pair_id,
    census_instrument_id,
    destination_function
  )

if (nrow(conflicting_link_sources) > 0L) {
  print(conflicting_link_sources)
  stop("An allocation link has conflicting revenue classifications.")
}

destination_links_deduplicated <- destination_candidates %>%
  arrange(country, pair_id, census_instrument_id, token_source) %>%
  distinct(
    country,
    pair_id,
    census_instrument_id,
    destination_function,
    .keep_all = TRUE
  )

post_dedup_duplicate_keys <- destination_links_deduplicated %>%
  count(
    country,
    pair_id,
    census_instrument_id,
    destination_function,
    name = "n_rows"
  ) %>%
  filter(n_rows > 1L)

if (nrow(post_dedup_duplicate_keys) > 0L) {
  print(post_dedup_duplicate_keys)
  stop("Allocation-link keys remain duplicated after deduplication.")
}

destination_exclusion_audit <- destination_links_deduplicated %>%
  mutate(
    destination_outcome = case_when(
      destination_function == "general_budget" ~ "general_budget",
      destination_function == "local_budget" ~ "local_budget",
      destination_function == "unallocated" ~ "unallocated",
      TRUE ~ "Retained"
    )
  ) %>%
  count(country, destination_outcome, name = "n_allocation_links") %>%
  complete(
    country = sort(unique(expanded_prepared$country)),
    destination_outcome = c(
      "general_budget",
      "local_budget",
      "unallocated",
      "Retained"
    ),
    fill = list(n_allocation_links = 0L)
  )

destination_links_filtered <- destination_links_deduplicated %>%
  filter(!destination_function %in% excluded_destinations)
```

```{r display-expansion-audit}
display_table(duplicate_link_candidates)
display_table(conflicting_link_sources)
display_table(post_dedup_duplicate_keys)
display_table(
  destination_exclusion_audit %>%
    filter(destination_outcome != "Retained", n_allocation_links > 0L)
)
```

# 4. Attach readable destination labels

Only four-digit COFOG groups are used for the readable destination names. An
unexpected destination code would be retained and explicitly marked as
unmapped rather than silently dropped.

```{r map-destination-labels}
cofog_group_mapping <- cofog_mapping_loaded %>%
  mutate(
    cofog_code = str_trim(cofog_code),
    level = str_to_lower(str_trim(level)),
    cofog_name = str_trim(cofog_name)
  ) %>%
  filter(level == "group") %>%
  select(cofog_code, cofog_name)

duplicate_cofog_group_codes <- cofog_group_mapping %>%
  count(cofog_code, name = "n_mapping_rows") %>%
  filter(n_mapping_rows > 1L)

if (nrow(duplicate_cofog_group_codes) > 0L ||
    any(is_missing_text(cofog_group_mapping$cofog_code))) {
  print(duplicate_cofog_group_codes)
  stop("The COFOG group mapping key is missing or duplicated.")
}

destination_links_labelled <- destination_links_filtered %>%
  left_join(
    cofog_group_mapping,
    by = c("destination_function" = "cofog_code"),
    relationship = "many-to-one"
  ) %>%
  mutate(
    destination_mapping_status = case_when(
      !is.na(cofog_name) ~ "mapped_gfsm_group",
      destination_function == "Missing" ~ "missing",
      TRUE ~ "unmapped"
    ),
    destination_function_name = case_when(
      !is.na(cofog_name) ~ cofog_name,
      destination_function == "Missing" ~ "Missing destination",
      TRUE ~ paste0("Unmapped code: ", destination_function)
    ),
    destination_label = make_destination_label(
      destination_function,
      destination_function_name
    ),
    revenue_class_label = make_revenue_class_label(revenue_class)
  ) %>%
  left_join(country_names, by = "country", relationship = "many-to-one")

unmapped_destination_functions <- destination_links_labelled %>%
  filter(destination_mapping_status == "unmapped") %>%
  count(country, destination_function, name = "n_allocation_links")

missing_retained_categories <- destination_links_labelled %>%
  summarise(
    missing_revenue_class = sum(is_missing_text(revenue_class)),
    missing_destination_function = sum(is_missing_text(destination_function)),
    missing_country = sum(is_missing_text(country)),
    missing_pair_id = sum(is_missing_text(pair_id)),
    missing_census_instrument_id = sum(
      is_missing_text(census_instrument_id)
    )
  )

if (any(unlist(missing_retained_categories) > 0L)) {
  print(missing_retained_categories)
  stop("A required retained plot category or key is missing.")
}
```

```{r display-mapping-audit}
display_table(duplicate_cofog_group_codes)
display_table(unmapped_destination_functions)
display_table(missing_retained_categories)
```

# 5. Country-level filter and link audit

Status and revenue-class exclusions below are counts of expanded input rows.
Destination exclusions and plotted totals are counts after destination
expansion and allocation-link deduplication, so the columns are intentionally
reported as separate stages rather than as a single row waterfall.

```{r build-country-audit}
country_universe <- expanded_prepared %>%
  distinct(country) %>%
  left_join(country_names, by = "country", relationship = "many-to-one")

country_audit <- country_universe %>%
  left_join(
    expanded_prepared %>%
      count(country, name = "input_expanded_rows"),
    by = "country"
  ) %>%
  left_join(
    status_eligible %>%
      count(country, name = "status_eligible_rows"),
    by = "country"
  ) %>%
  left_join(
    revenue_class_exclusion_audit %>%
      filter(exclusion == "Missing") %>%
      transmute(
        country,
        rows_excluded_missing_revenue = n_expanded_rows
      ),
    by = "country"
  ) %>%
  left_join(
    revenue_class_exclusion_audit %>%
      filter(exclusion == "unclassified_review") %>%
      transmute(
        country,
        rows_excluded_unclassified_review = n_expanded_rows
      ),
    by = "country"
  ) %>%
  left_join(
    revenue_eligible %>%
      count(country, name = "revenue_eligible_rows"),
    by = "country"
  ) %>%
  left_join(
    destination_candidates %>%
      count(country, name = "destination_candidates"),
    by = "country"
  ) %>%
  left_join(
    duplicate_link_candidates %>%
      group_by(country) %>%
      summarise(
        duplicate_link_candidates_removed = sum(n_candidate_rows - 1L),
        .groups = "drop"
      ),
    by = "country"
  ) %>%
  left_join(
    destination_exclusion_audit %>%
      filter(destination_outcome == "general_budget") %>%
      transmute(
        country,
        links_excluded_general_budget = n_allocation_links
      ),
    by = "country"
  ) %>%
  left_join(
    destination_exclusion_audit %>%
      filter(destination_outcome == "local_budget") %>%
      transmute(
        country,
        links_excluded_local_budget = n_allocation_links
      ),
    by = "country"
  ) %>%
  left_join(
    destination_exclusion_audit %>%
      filter(destination_outcome == "unallocated") %>%
      transmute(
        country,
        links_excluded_unallocated = n_allocation_links
      ),
    by = "country"
  ) %>%
  left_join(
    destination_links_labelled %>%
      group_by(country) %>%
      summarise(
        links_plotted = n(),
        pairs_plotted = n_distinct(pair_id),
        revenue_classes_plotted = n_distinct(revenue_class),
        destination_functions_plotted = n_distinct(destination_function),
        .groups = "drop"
      ),
    by = "country"
  ) %>%
  mutate(
    across(
      where(is.numeric),
      ~replace_na(.x, 0L)
    ),
    rows_excluded_status = input_expanded_rows - status_eligible_rows,
    plot_status = case_when(
      status_eligible_rows == 0L ~
        "Not plotted: no partial or exclusive earmark rows",
      revenue_eligible_rows == 0L ~
        "Not plotted: all eligible rows have excluded revenue classes",
      links_plotted == 0L ~
        "Not plotted: all links have excluded destinations",
      TRUE ~ "Plotted"
    )
  ) %>%
  arrange(country)

country_audit <- country_audit %>%
  select(
    country,
    country_name,
    input_expanded_rows,
    status_eligible_rows,
    rows_excluded_status,
    rows_excluded_missing_revenue,
    rows_excluded_unclassified_review,
    revenue_eligible_rows,
    destination_candidates,
    duplicate_link_candidates_removed,
    links_excluded_general_budget,
    links_excluded_local_budget,
    links_excluded_unallocated,
    links_plotted,
    pairs_plotted,
    revenue_classes_plotted,
    destination_functions_plotted,
    plot_status
  )

countries_to_plot <- country_audit %>%
  filter(plot_status == "Plotted") %>%
  pull(country)

if (length(countries_to_plot) == 0L) {
  stop("No country has an eligible source-to-destination link.")
}
```

```{r display-country-audit}
display_table(country_audit)
```

# 6. Aggregate the exact plot data

Each CSV row is one country-level source-to-destination edge. The edge count is
the number of distinct allocation links represented by that flow.

```{r aggregate-plot-data}
flow_counts <- destination_links_labelled %>%
  count(
    country,
    country_name,
    revenue_class,
    revenue_class_label,
    destination_function,
    destination_function_name,
    destination_label,
    destination_mapping_status,
    name = "n_allocation_links"
  )

source_order <- flow_counts %>%
  group_by(country, revenue_class, revenue_class_label) %>%
  summarise(
    source_total_links = sum(n_allocation_links),
    .groups = "drop"
  ) %>%
  group_by(country) %>%
  arrange(desc(source_total_links), revenue_class_label, .by_group = TRUE) %>%
  mutate(
    source_order = row_number(),
    source_plot_label = str_wrap(revenue_class_label, width = 28),
    source_colour = unname(revenue_class_colours[revenue_class])
  ) %>%
  ungroup()

destination_order <- flow_counts %>%
  group_by(
    country,
    destination_function,
    destination_function_name,
    destination_label,
    destination_mapping_status
  ) %>%
  summarise(
    destination_total_links = sum(n_allocation_links),
    .groups = "drop"
  ) %>%
  mutate(order_key = destination_order_key(destination_function)) %>%
  group_by(country) %>%
  arrange(order_key, destination_function, destination_label, .by_group = TRUE) %>%
  mutate(
    destination_order = row_number(),
    destination_plot_label = str_wrap(destination_label, width = 30)
  ) %>%
  ungroup() %>%
  select(-order_key)

country_plot_settings <- flow_counts %>%
  group_by(country) %>%
  summarise(
    total_allocation_links = sum(n_allocation_links),
    n_source_nodes = n_distinct(revenue_class),
    n_destination_nodes = n_distinct(destination_function),
    .groups = "drop"
  ) %>%
  mutate(
    destination_label_minimum = case_when(
      total_allocation_links <= 20L ~ 1L,
      total_allocation_links <= 40L ~ 2L,
      total_allocation_links <= 150L ~ 3L,
      TRUE ~ 5L
    ),
    source_label_minimum = pmax(
      destination_label_minimum,
      ceiling(total_allocation_links * 0.08)
    ),
    plot_width_inches = if_else(n_destination_nodes > 16L, 18, 16),
    plot_height_inches = pmin(
      20,
      pmax(9, 7 + 0.35 * pmax(n_source_nodes, n_destination_nodes))
    )
  )

all_country_flow_data <- flow_counts %>%
  left_join(
    source_order,
    by = c("country", "revenue_class", "revenue_class_label")
  ) %>%
  left_join(
    destination_order,
    by = c(
      "country",
      "destination_function",
      "destination_function_name",
      "destination_label",
      "destination_mapping_status"
    )
  ) %>%
  left_join(country_plot_settings, by = "country") %>%
  mutate(
    source_label_shown = source_total_links >= source_label_minimum,
    destination_label_shown =
      destination_total_links >= destination_label_minimum
  ) %>%
  arrange(country, source_order, destination_order) %>%
  select(
    country,
    country_name,
    revenue_class,
    revenue_class_label,
    source_plot_label,
    source_order,
    source_total_links,
    source_colour,
    source_label_minimum,
    source_label_shown,
    destination_function,
    destination_function_name,
    destination_label,
    destination_plot_label,
    destination_order,
    destination_total_links,
    destination_mapping_status,
    destination_label_minimum,
    destination_label_shown,
    n_allocation_links
  )

plot_data_by_country <- all_country_flow_data %>%
  split(.$country) %>%
  map(~arrange(.x, source_order, destination_order))

aggregation_checks <- imap_dfr(
  plot_data_by_country,
  function(data, country_code) {
    expected_links <- country_audit %>%
      filter(country == country_code) %>%
      pull(links_plotted)
    tibble(
      country = country_code,
      csv_rows = nrow(data),
      sum_edge_counts = sum(data$n_allocation_links),
      expected_links = expected_links,
      duplicate_edges = sum(duplicated(
        data[c("country", "revenue_class", "destination_function")]
      )),
      passed = sum_edge_counts == expected_links & duplicate_edges == 0L
    )
  }
)

if (!all(aggregation_checks$passed)) {
  print(aggregation_checks %>% filter(!passed))
  stop("One or more country flow tables failed reconciliation.")
}
```

```{r display-aggregation-checks}
display_table(aggregation_checks)
```

# 7. Create the Sankey plots

Flow colours identify source revenue classes and remain stable across all
countries. Node and axis-side text are deliberately enlarged for presentation
use. Very small destination nodes are left unlabeled when necessary; their
codes and names remain in the matching CSV.

```{r define-plot-function}
make_sankey_plot <- function(flow_data) {
  country_name <- unique(flow_data$country_name)
  total_links <- sum(flow_data$n_allocation_links)
  destination_label_minimum <- unique(
    flow_data$destination_label_minimum
  )
  source_label_minimum <- unique(flow_data$source_label_minimum)

  source_levels <- flow_data %>%
    distinct(source_plot_label, source_order) %>%
    arrange(source_order) %>%
    pull(source_plot_label)
  destination_levels <- flow_data %>%
    distinct(destination_plot_label, destination_order) %>%
    arrange(destination_order) %>%
    pull(destination_plot_label)
  legend_values <- flow_data %>%
    distinct(revenue_class, revenue_class_label, source_order) %>%
    arrange(source_order)

  plot_data <- flow_data %>%
    mutate(
      source_plot_label = factor(
        source_plot_label,
        levels = source_levels
      ),
      destination_plot_label = factor(
        destination_plot_label,
        levels = destination_levels
      )
    )

  ggplot(
    plot_data,
    aes(
      axis1 = source_plot_label,
      axis2 = destination_plot_label,
      y = n_allocation_links
    )
  ) +
    geom_alluvium(
      aes(fill = revenue_class),
      width = 0.18,
      alpha = 0.82,
      knot.pos = 0.45,
      colour = "#FFFFFF",
      linewidth = 0.16
    ) +
    geom_stratum(
      width = 0.20,
      fill = "#D7E0E7",
      colour = "#FFFFFF",
      linewidth = 0.55
    ) +
    geom_text(
      stat = "stratum",
      aes(
        label = after_stat(ifelse(
          (x == 1 & count >= source_label_minimum) |
            (x == 2 & count >= destination_label_minimum),
          as.character(stratum),
          ""
        ))
      ),
      colour = "#253746",
      size = 5.0,
      lineheight = 0.95,
      fontface = "plain"
    ) +
    scale_x_discrete(
      limits = c("Revenue source", "Earmarked destination"),
      expand = c(0.14, 0.14)
    ) +
    scale_fill_manual(
      values = revenue_class_colours,
      breaks = legend_values$revenue_class,
      labels = legend_values$revenue_class_label,
      drop = TRUE,
      name = "Revenue source"
    ) +
    guides(
      fill = guide_legend(
        ncol = 3,
        byrow = TRUE,
        title.position = "top",
        override.aes = list(alpha = 0.95)
      )
    ) +
    labs(
      title = paste0(
        "From revenue source to earmarked destination \u2014 ",
        country_name
      ),
      subtitle = paste0(
        "Flow width represents ",
        comma(total_links),
        " distinct statutory allocation links"
      ),
      x = NULL,
      y = NULL,
      caption = str_wrap(
        paste0(
          "Note: Partial and exclusive earmark instruments only. General ",
          "budget, local budget, and unallocated destinations are excluded. ",
          "Each link is a distinct allocation pair, linked census instrument, ",
          "and destination combination. Source nodes with fewer than ",
          source_label_minimum,
          " links and destination nodes with fewer than ",
          destination_label_minimum,
          " links are left unlabeled when needed; every label is retained in ",
          "the matching CSV."
        ),
        width = 135
      )
    ) +
    theme_minimal(base_size = 15) +
    theme(
      plot.title.position = "plot",
      plot.title = element_text(
        face = "bold",
        colour = "#222222",
        size = 18
      ),
      plot.subtitle = element_text(
        colour = "#4D4D4D",
        size = 13,
        margin = margin(b = 12)
      ),
      plot.caption = element_text(
        colour = "#666666",
        hjust = 0,
        size = 10,
        margin = margin(t = 11)
      ),
      axis.text.x = element_text(
        face = "bold",
        colour = "#333333",
        size = 15
      ),
      axis.text.y = element_blank(),
      axis.ticks = element_blank(),
      panel.grid = element_blank(),
      legend.position = "bottom",
      legend.text = element_text(size = 11.5),
      legend.title = element_text(face = "bold", size = 12.5),
      legend.key.height = unit(0.55, "cm"),
      legend.key.width = unit(0.85, "cm"),
      plot.margin = margin(14, 26, 14, 26)
    )
}

plot_objects_by_country <- map(
  plot_data_by_country,
  make_sankey_plot
)
```

# 8. Save and verify matching PNG and CSV outputs

Files are first written to a temporary staging directory. The complete bundle
is copied into the existing country-focus folders only after every CSV and PNG
passes its read-back checks.

```{r define-output-paths}
output_stems <- set_names(
  paste0(
    "11_",
    str_to_lower(names(plot_data_by_country)),
    "_revenue_to_destination_sankey"
  ),
  names(plot_data_by_country)
)

csv_output_paths <- set_names(
  file.path(table_dir, paste0(output_stems, ".csv")),
  names(output_stems)
)
png_output_paths <- set_names(
  file.path(figure_dir, paste0(output_stems, ".png")),
  names(output_stems)
)
```

```{r stage-and-verify-output-bundle}
stage_dir <- tempfile(pattern = "11_all_country_sankey_stage_")
dir.create(stage_dir, recursive = TRUE, showWarnings = FALSE)

stage_csv_paths <- set_names(
  file.path(stage_dir, paste0(output_stems, ".csv")),
  names(output_stems)
)
stage_png_paths <- set_names(
  file.path(stage_dir, paste0(output_stems, ".png")),
  names(output_stems)
)

iwalk(plot_data_by_country, function(data, country_code) {
  write_csv(data, stage_csv_paths[[country_code]], na = "")
})

iwalk(plot_objects_by_country, function(plot_object, country_code) {
  dimensions <- country_plot_settings %>%
    filter(country == country_code)
  ggsave(
    filename = stage_png_paths[[country_code]],
    plot = plot_object,
    width = dimensions$plot_width_inches,
    height = dimensions$plot_height_inches,
    units = "in",
    dpi = 300,
    bg = "white",
    limitsize = FALSE
  )
})

same_table_values <- function(saved, expected) {
  expected_as_text <- expected %>%
    mutate(across(everything(), as.character))
  identical(names(saved), names(expected)) &&
    nrow(saved) == nrow(expected) &&
    identical(as.data.frame(saved), as.data.frame(expected_as_text))
}

staged_csv_checks <- imap_dfr(
  plot_data_by_country,
  function(expected, country_code) {
    saved <- read_csv(
      stage_csv_paths[[country_code]],
      col_types = cols(.default = col_character()),
      show_col_types = FALSE,
      progress = FALSE
    )
    tibble(
      country = country_code,
      csv_rows = nrow(saved),
      passed = same_table_values(saved, expected)
    )
  }
)

staged_png_checks <- imap_dfr(
  plot_objects_by_country,
  function(plot_object, country_code) {
    tibble(
      country = country_code,
      png_bytes = file.info(stage_png_paths[[country_code]])$size,
      passed = file.exists(stage_png_paths[[country_code]]) &&
        file.info(stage_png_paths[[country_code]])$size > 10000
    )
  }
)

if (!all(staged_csv_checks$passed) || !all(staged_png_checks$passed)) {
  print(staged_csv_checks %>% filter(!passed))
  print(staged_png_checks %>% filter(!passed))
  stop("Staged CSV or PNG checks failed; final outputs were not changed.")
}
```

```{r commit-output-bundle}
final_paths <- c(
  set_names(
    unname(csv_output_paths),
    paste0(names(csv_output_paths), "_csv")
  ),
  set_names(
    unname(png_output_paths),
    paste0(names(png_output_paths), "_png")
  )
)
stage_paths <- c(
  set_names(
    unname(stage_csv_paths),
    paste0(names(stage_csv_paths), "_csv")
  ),
  set_names(
    unname(stage_png_paths),
    paste0(names(stage_png_paths), "_png")
  )
)

commit_output_bundle <- function(stage_paths, final_paths) {
  existed_before <- set_names(file.exists(final_paths), names(final_paths))
  backup_dir <- tempfile(pattern = "11_all_country_sankey_backup_")
  dir.create(backup_dir, recursive = TRUE, showWarnings = FALSE)
  backup_paths <- file.path(backup_dir, basename(final_paths))
  names(backup_paths) <- names(final_paths)
  committed <- FALSE

  on.exit({
    if (!committed) {
      for (name in names(final_paths)) {
        if (existed_before[[name]] && file.exists(backup_paths[[name]])) {
          file.copy(
            backup_paths[[name]],
            final_paths[[name]],
            overwrite = TRUE
          )
        } else if (!existed_before[[name]] &&
                   file.exists(final_paths[[name]])) {
          file.remove(final_paths[[name]])
        }
      }
    }
    unlink(stage_dir, recursive = TRUE, force = TRUE)
    unlink(backup_dir, recursive = TRUE, force = TRUE)
  }, add = TRUE)

  for (name in names(final_paths)[existed_before]) {
    if (!file.copy(
      final_paths[[name]],
      backup_paths[[name]],
      overwrite = FALSE
    )) {
      stop("Could not create a rollback copy for ", name, ".")
    }
  }

  commit_ok <- vapply(
    names(final_paths),
    function(name) file.copy(
      stage_paths[[name]],
      final_paths[[name]],
      overwrite = TRUE
    ),
    logical(1)
  )

  if (!all(commit_ok)) {
    stop("Could not commit the complete output bundle.")
  }

  committed <- TRUE
  invisible(TRUE)
}

commit_output_bundle(stage_paths, final_paths)
```

```{r verify-final-outputs}
final_csv_checks <- imap_dfr(
  plot_data_by_country,
  function(expected, country_code) {
    saved <- read_csv(
      csv_output_paths[[country_code]],
      col_types = cols(.default = col_character()),
      show_col_types = FALSE,
      progress = FALSE
    )
    tibble(
      country = country_code,
      csv_path = csv_output_paths[[country_code]],
      csv_rows = nrow(saved),
      csv_matches = same_table_values(saved, expected)
    )
  }
)

final_png_checks <- imap_dfr(
  plot_objects_by_country,
  function(plot_object, country_code) {
    tibble(
      country = country_code,
      png_path = png_output_paths[[country_code]],
      png_bytes = file.info(png_output_paths[[country_code]])$size,
      png_valid = file.exists(png_output_paths[[country_code]]) &&
        file.info(png_output_paths[[country_code]])$size > 10000
    )
  }
)

input_md5_after <- unname(tools::md5sum(input_paths))
input_preservation_check <- tibble(
  input = names(input_paths),
  md5_before = input_md5_before,
  md5_after = input_md5_after,
  unchanged = input_md5_before == input_md5_after
)

if (!all(final_csv_checks$csv_matches) ||
    !all(final_png_checks$png_valid) ||
    !all(input_preservation_check$unchanged)) {
  print(final_csv_checks %>% filter(!csv_matches))
  print(final_png_checks %>% filter(!png_valid))
  print(input_preservation_check %>% filter(!unchanged))
  stop("Final output or input-preservation checks failed.")
}

output_manifest <- final_csv_checks %>%
  select(country, csv_path, csv_rows, csv_matches) %>%
  left_join(
    final_png_checks %>%
      select(country, png_path, png_bytes, png_valid),
    by = "country"
  ) %>%
  left_join(
    country_audit %>%
      select(country, country_name, links_plotted),
    by = "country"
  ) %>%
  arrange(country) %>%
  select(
    country,
    country_name,
    links_plotted,
    csv_rows,
    csv_matches,
    png_bytes,
    png_valid,
    csv_path,
    png_path
  )
```

```{r display-output-checks}
display_table(input_preservation_check)
display_table(output_manifest)
```

# 9. Saved figures

```{r display-saved-figures, echo=FALSE, results='asis'}
for (country_code in names(png_output_paths)) {
  country_name <- country_names %>%
    filter(country == country_code) %>%
    pull(country_name)
  cat("\n\n## ", country_name, "\n\n", sep = "")
  print(knitr::include_graphics(
    png_output_paths[[country_code]],
    rel_path = FALSE
  ))
  cat("\n\n")
}
```

# 10. Concise completion summary

```{r completion-summary, results='asis'}
not_plotted <- country_audit %>%
  filter(plot_status != "Plotted")

cat(
  "- Countries plotted: ",
  length(countries_to_plot),
  " (",
  paste(countries_to_plot, collapse = ", "),
  ").\n",
  "- Distinct allocation links plotted: ",
  comma(sum(country_audit$links_plotted)),
  ".\n",
  "- Expanded rows excluded by earmark status: ",
  comma(sum(country_audit$rows_excluded_status)),
  ".\n",
  "- Status-eligible rows excluded for missing revenue class: ",
  comma(sum(country_audit$rows_excluded_missing_revenue)),
  ".\n",
  "- Status-eligible rows excluded as unclassified_review: ",
  comma(sum(country_audit$rows_excluded_unclassified_review)),
  ".\n",
  "- Expanded destination links excluded as general budget, local budget, ",
  "or unallocated: ",
  comma(
    sum(country_audit$links_excluded_general_budget) +
      sum(country_audit$links_excluded_local_budget) +
      sum(country_audit$links_excluded_unallocated)
  ),
  ".\n",
  "- Duplicate allocation-link candidates removed: ",
  comma(sum(country_audit$duplicate_link_candidates_removed)),
  ".\n",
  "- Unmapped retained destination codes: ",
  nrow(unmapped_destination_functions),
  ".\n",
  "- Countries without an eligible plot: ",
  paste0(
    not_plotted$country,
    " (",
    str_remove(not_plotted$plot_status, "^Not plotted: "),
    ")",
    collapse = "; "
  ),
  ".\n",
  sep = ""
)
```
